# Appendices {-}

```{=latex}
\appendix
```


```{=latex}
\input{/Users/ebbamark/Documents/Documents - Nuff-Malham/GitHub/pub_fin_schools/code/si_tex_docs/si.tex}
```

# Descriptive Regression Results

In the following set of results, I report descriptive regressions to establish relationships between property taxes, education expenditure, GDP (total, private industry, O&G&mining), etc.
All regression models that follow include TWFE (CZ- and year- fixed effects) and standard errors clustered by commuting zone.
All functional forms in the feols() functions below are of the form Y ~ X In the cases in which multiple estimations are included via sw(Xa, Xb, Xc + Xd), the function will return results for Y~Xa, Y~Xb, Y ~ Xc + Xd.

```

## Property Tax ~ GDP

GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) leads to a 0.38% (0.32%) increase in property taxes collected (per capita).

```

## Education Expenditure ~ Revenue Sources

The below regressions are included to establish the relationship between education expenditure and its component parts. These regressions simply corroborate what is displayed in the section on Key Relationships in [LINK](https://ebbam.github.io/pub_fin_schools/code/cz_results.html) (ie. that the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).

```

## Education Expenditure ~ GDP

A 1% increase in GDP pc is associated with a 0.19% increase in education expenditure per pupil, dominated by the effect of GDP from private industry (0.16%). I include here also the GDP generated from the oil, gas, mining, and quarrying sector. The effect is small and statistically insignificant.

```

# Key Relationships between Economic Variables

Below we display key relationships between several of the economic variables in our study.

```

## Baseline Regressions with State Fixed Effects

Regressions establishing baseline relationships between local economic variables and elementary education expenditure using state-fixed effects rather than commuting-zone level effects.

```

## SS Construction

Plots of the data inputs to the shift-share instrument.

```

## Panel VAR Specification

$$
Y_{it} = \alpha_i +  \sum_{k = 1}^{4} \gamma_{k}A_{i,t-k} + \beta X_{it} + \varepsilon_{it}
$$

Where we approach a level and per capita value expression of the
relationship between total education expenditure, intergovernmental
revenue, house prices conditioned on GDP and wage levels.

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Total Educ. Exp.})_{it} \\
\log(\text{real Total IG Revenue})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP})_{it} \\
\log(\text{wage})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $3 \times 3$ coefficient matrices
- $\beta$ is a $3 \times 2$ matrix of coefficients on the exogenous variables
- $\alpha_i$ is a vector of unit fixed effects
- $\varepsilon_{it}$ is the error term

Where

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Own Source Rev. per person})_{it} \\
\log(\text{real IG Revenue per person})_{it} \\
\log(\text{wage})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP per capita})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $4 \times 4$ coefficient matrices
- $B$ is a $4 \times 1$ coefficient matrix
- $\alpha_i$ unit fixed effects
- $\varepsilon_{it}$ error term

```{r, ref.label='appendix_pvar', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Panel VAR Specification

$$
Y_{it} = \alpha_i +  \sum_{k = 1}^{4} \gamma_{k}A_{i,t-k} + \beta X_{it} + \varepsilon_{it}
$$

Where we approach a level and per capita value expression of the
relationship between total education expenditure, intergovernmental
revenue, house prices conditioned on GDP and wage levels.

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Total Educ. Exp.})_{it} \\
\log(\text{real Total IG Revenue})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP})_{it} \\
\log(\text{wage})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $3 \times 3$ coefficient matrices
- $\beta$ is a $3 \times 2$ matrix of coefficients on the exogenous variables
- $\alpha_i$ is a vector of unit fixed effects
- $\varepsilon_{it}$ is the error term

Where

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Own Source Rev. per person})_{it} \\
\log(\text{real IG Revenue per person})_{it} \\
\log(\text{wage})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP per capita})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $4 \times 4$ coefficient matrices
- $B$ is a $4 \times 1$ coefficient matrix
- $\alpha_i$ unit fixed effects
- $\varepsilon_{it}$ error term

```{r, ref.label='unnamed-chunk-81', echo = FALSE, eval=TRUE, results = 'asis'}
```

