# Appendices {-}

```{=latex}
\appendix
```


```{=latex}
\input{/Users/ebbamark/Documents/Documents - Nuff-Malham/GitHub/pub_fin_schools/code/si_tex_docs/si.tex}
```

# Descriptive Regression Results

In the following set of results, I report descriptive regressions to establish relationships between property taxes, education expenditure, GDP (total, private industry, O&G&mining), etc.
All regression models that follow include TWFE (CZ- and year- fixed effects) and standard errors clustered by commuting zone.
All functional forms in the feols() functions below are of the form Y ~ X In the cases in which multiple estimations are included via sw(Xa, Xb, Xc + Xd), the function will return results for Y~Xa, Y~Xb, Y ~ Xc + Xd.

```{r, ref.label='appendix1', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Property Tax ~ GDP

GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) leads to a 0.38% (0.32%) increase in property taxes collected (per capita).

```{r, ref.label='appendix2', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ Revenue Sources

The below regressions are included to establish the relationship between education expenditure and its component parts (ie. the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).

```{r, ref.label='appendix3', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ GDP

A 1% increase in GDP pc is associated with a 0.19% increase in education expenditure per pupil, dominated by the effect of GDP from private industry (0.16%). I include here also the GDP generated from the oil, gas, mining, and quarrying sector. The effect is small and statistically insignificant.

```{r, ref.label='appendix4', echo = FALSE, eval=TRUE, results = 'asis'}
```

# Key Relationships between Economic Variables

Below we display key relationships between several of the economic variables in our study.

```{r, ref.label='KRs', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Baseline Regressions with State Fixed Effects

Regressions establishing baseline relationships between local economic variables and elementary education expenditure using state-fixed effects rather than commuting-zone level effects.

```{r, ref.label='baseline_regs_state_fe', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Funding Share Interaction

Furthermore, given the heterogeneity in reliance on intergovernmental
transfers (largely coming from the state), we interact all economic
predictors above with a variable that represents the share of total
elementary education expenditure (as a continuous variable) coming from
state-level funding.


\textcolor{red}{1. heatmap of effect vs. share when varying the share in a linear combinantion of interactive and stand-alone beta. 2. Make sure you can reconcile table 4 and table 2 together by showing that the linear combination of the coefficients using the mean share is the same as in table 2.}

```{r, ref.label='funding_share_interaction', echo = FALSE, eval=TRUE, results = 'asis'}
```

## SS Construction

Plots of the data inputs to the shift-share instrument.

```{r, ref.label='appendix_industry_shares_cleaning', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Wage-based Shift Share Instrument Results

First, we display the results for a 2SLS estimation using our wage-based
shift-share instrument.


In Table 5, the instrumental variable estimates provide evidence of a
robust causal relationship between local wages and public education
expenditure. Utilising our wage-based shift-share instrument we see
highly significant and relevant first-stage relationships when the
shift-share instrument is imposed in levels (except in column 5). In
each case except columns 5-6 (l1 SS, CZ FE), the first-stage regression
yields a statistically significant
	extcolor{violet}{and economically large coefficient}. Varying the
time-lag and inclusion of state or commuting zone fixed effects, we see
that a 1% increase in the shift-share measure (which can be interpreted
as a natural logarithm) is associated with a 0.02-0.06% increase in
average weekly wages (p \< 0.01), with an F-statistic between 11-103
(near or above conventional weak instrument thresholds) confirming
instrument relevance. The Wu-Hausman tests reject the null of
exogeneity, confirming that OLS estimates are biased and IV estimation
is appropriate. Wald tests of joint significance further support the
strength of the instruments.


Using wage shocks in levels yields strong instruments, high first-stage
F-statistics, and stable second-stage estimates: higher local wages
robustly increase education spending. Furthermore, given the dependent
variable measures per pupil expenditure, this result implies direct
effects in experience per student. In contrast, when shocks are measured
in growth rates (Table 6), the instruments lose predictive power
(first-stage F-statistics \~1–2), resulting in weak identification. The
second-stage coefficients become unstable and often insignificant, while
Hausman tests fail to reject exogeneity. This suggests that the
growth-rate specification is poorly identified and cannot provide
reliable causal inference, whereas the level specification produces
credible and consistent results.


The specification that uses wage shocks in levels provides the most
credible identification strategy. Since levels capture the
cross-sectional fiscal variation that drives differences in property
values and school spending, the level specification is more consistent
with the economic mechanisms of interest and delivers more reliable
causal estimates. At the same time, the weakness of the growth-rate
specification does raise concerns about the robustness of the results.
If the relationship between wages, house prices, and education
expenditure is driven by common non-stationary trends, then regressions
in levels risk spurious correlation. The fact that the IV design loses
power when variables are differenced into growth rates may suggest that
part of the strong level results reflect long-run trends rather than
short-run causal shocks. While the large first-stage F-statistics and
Hausman tests in the level specification support its validity, the weak
performance of the growth-rate version cautions that the results could
be sensitive to issues of persistence and trending in the data. Taken
together, these results suggest that while the level specification
provides strong identification and compelling evidence of a positive
causal effect of local wages on education spending, the weak performance
of the growth-rate specification highlights the need for caution, as the
strength of the findings may partly reflect long-run trending
relationships rather than purely exogenous shocks. However, examining
the structure of the growth rate shock, the instability of the variable
is likely causing the poor identification in the growth rate
regressions.

```{r, ref.label='wage_ss_res', echo = FALSE, eval=TRUE, results = 'asis'}
```

### Baseline Models for Growth Rate Sample Partitioning



```{r, ref.label='baseline_growth_rate_partitioning', echo = FALSE, eval=TRUE, results = 'asis'}
```

### Wage-based SS Instrument for Growth Rate Sample Partitioning

When applying our instrumental variable design we find that majority of
our relevant signal is picked up in regions across both trend metrics. Though the effects are more consistent across wage-trending sub-samples.

```{r, ref.label='wage_ss_growth_rate_partitioning', echo = FALSE, eval=TRUE, results = 'asis'}
```

### Wage-based SS State by State Estimation



```{r, ref.label='wage_ss_state_by_state', echo = FALSE, eval=TRUE, results = 'asis'}
```

#### Wage-based SS in Industry-by-Industry Estimation

Using our wage-based shift share instrument, the below demonstrates the
overall treatment effect of local wage changes instrumented via our
wage-based shift-share shock.

```{r, ref.label='wage_ss_ind_by_ind', echo = FALSE, eval=TRUE, results = 'asis'}
```

### High-Income Outliers

There is a somewhat non-linear relationship between property taxes and
elementary expenditure as property taxes collected rise as represented
in Figure 22 below. This happens largely as a result of very high-income
commuting zones. Therefore, we exclude any commuting zone that spends
more than 22k per pupil to avoid any distorting effects. This removes 12
CZs (\~2% of the sample). \textcolor{red}{This could benefit from
more robust outlier detection.}
To assess whether the main results are driven by a small number of very
high-income jurisdictions, I re-estimate the baseline and IV
specifications excluding such outliers. The findings are fully
consistent with the baseline analysis: house prices remain a strong
predictor of local education spending, and the IV estimates continue to
imply that a 10% increase in house prices raises per-pupil expenditure
by roughly 4–6%. The wage-based shift-share instrument yields somewhat
larger point estimates, though with wider standard errors, while the
VA-based instrument produces effects in line with earlier results.
Overall, this robustness exercise confirms that the causal relationship
between housing wealth and education spending is not confined to
affluent areas but reflects a broader, generalizable pattern. These
findings confirm that the main result is not driven solely by affluent
jurisdictions, but reflects a more general relationship between local
housing wealth and education spending.

```{r, ref.label='high_income_outliers', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Panel VAR Specification

$$
Y_{it} = \alpha_i +  \sum_{k = 1}^{4} \gamma_{k}A_{i,t-k} + \beta X_{it} + \varepsilon_{it}
$$

Where we approach a level and per capita value expression of the
relationship between total education expenditure, intergovernmental
revenue, house prices conditioned on GDP and wage levels.

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Total Educ. Exp.})_{it} \\
\log(\text{real Total IG Revenue})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP})_{it} \\
\log(\text{wage})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $3 \times 3$ coefficient matrices
- $\beta$ is a $3 \times 2$ matrix of coefficients on the exogenous variables
- $\alpha_i$ is a vector of unit fixed effects
- $\varepsilon_{it}$ is the error term

Where

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Own Source Rev. per person})_{it} \\
\log(\text{real IG Revenue per person})_{it} \\
\log(\text{wage})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP per capita})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $4 \times 4$ coefficient matrices
- $B$ is a $4 \times 1$ coefficient matrix
- $\alpha_i$ unit fixed effects
- $\varepsilon_{it}$ error term

```{r, ref.label='appendix_pvar', echo = FALSE, eval=TRUE, results = 'asis'}
```

# Descriptive Regression Results

In the following set of results, I report descriptive regressions to establish relationships between property taxes, education expenditure, GDP (total, private industry, O&G&mining), etc.
All regression models that follow include TWFE (CZ- and year- fixed effects) and standard errors clustered by commuting zone.
All functional forms in the feols() functions below are of the form Y ~ X In the cases in which multiple estimations are included via sw(Xa, Xb, Xc + Xd), the function will return results for Y~Xa, Y~Xb, Y ~ Xc + Xd.

```{r, ref.label='unnamed-chunk-85', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Property Tax ~ GDP

GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) leads to a 0.38% (0.32%) increase in property taxes collected (per capita).

```{r, ref.label='unnamed-chunk-86', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ Revenue Sources

The below regressions are included to establish the relationship between education expenditure and its component parts (ie. the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).

```{r, ref.label='unnamed-chunk-87', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ GDP

A 1% increase in GDP pc is associated with a 0.19% increase in education expenditure per pupil, dominated by the effect of GDP from private industry (0.16%). I include here also the GDP generated from the oil, gas, mining, and quarrying sector. The effect is small and statistically insignificant.

```{r, ref.label='unnamed-chunk-88', echo = FALSE, eval=TRUE, results = 'asis'}
```

# Key Relationships between Economic Variables

Below we display key relationships between several of the economic variables in our study.

```{r, ref.label='unnamed-chunk-89', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Baseline Regressions with State Fixed Effects

Regressions establishing baseline relationships between local economic variables and elementary education expenditure using state-fixed effects rather than commuting-zone level effects.

```{r, ref.label='unnamed-chunk-90', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Funding Share Interaction

Furthermore, given the heterogeneity in reliance on intergovernmental
transfers (largely coming from the state), we interact all economic
predictors above with a variable that represents the share of total
elementary education expenditure (as a continuous variable) coming from
state-level funding.


\textcolor{red}{1. heatmap of effect vs. share when varying the share in a linear combinantion of interactive and stand-alone beta. 2. Make sure you can reconcile table 4 and table 2 together by showing that the linear combination of the coefficients using the mean share is the same as in table 2.}

```{r, ref.label='unnamed-chunk-91', echo = FALSE, eval=TRUE, results = 'asis'}
```

## SS Construction

Plots of the data inputs to the shift-share instrument.

```{r, ref.label='unnamed-chunk-92', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Wage-based Shift Share Instrument Results

First, we display the results for a 2SLS estimation using our wage-based
shift-share instrument.


In Table 5, the instrumental variable estimates provide evidence of a
robust causal relationship between local wages and public education
expenditure. Utilising our wage-based shift-share instrument we see
highly significant and relevant first-stage relationships when the
shift-share instrument is imposed in levels (except in column 5). In
each case except columns 5-6 (l1 SS, CZ FE), the first-stage regression
yields a statistically significant
	extcolor{violet}{and economically large coefficient}. Varying the
time-lag and inclusion of state or commuting zone fixed effects, we see
that a 1% increase in the shift-share measure (which can be interpreted
as a natural logarithm) is associated with a 0.02-0.06% increase in
average weekly wages (p \< 0.01), with an F-statistic between 11-103
(near or above conventional weak instrument thresholds) confirming
instrument relevance. The Wu-Hausman tests reject the null of
exogeneity, confirming that OLS estimates are biased and IV estimation
is appropriate. Wald tests of joint significance further support the
strength of the instruments.


Using wage shocks in levels yields strong instruments, high first-stage
F-statistics, and stable second-stage estimates: higher local wages
robustly increase education spending. Furthermore, given the dependent
variable measures per pupil expenditure, this result implies direct
effects in experience per student. In contrast, when shocks are measured
in growth rates (Table 6), the instruments lose predictive power
(first-stage F-statistics \~1–2), resulting in weak identification. The
second-stage coefficients become unstable and often insignificant, while
Hausman tests fail to reject exogeneity. This suggests that the
growth-rate specification is poorly identified and cannot provide
reliable causal inference, whereas the level specification produces
credible and consistent results.


The specification that uses wage shocks in levels provides the most
credible identification strategy. Since levels capture the
cross-sectional fiscal variation that drives differences in property
values and school spending, the level specification is more consistent
with the economic mechanisms of interest and delivers more reliable
causal estimates. At the same time, the weakness of the growth-rate
specification does raise concerns about the robustness of the results.
If the relationship between wages, house prices, and education
expenditure is driven by common non-stationary trends, then regressions
in levels risk spurious correlation. The fact that the IV design loses
power when variables are differenced into growth rates may suggest that
part of the strong level results reflect long-run trends rather than
short-run causal shocks. While the large first-stage F-statistics and
Hausman tests in the level specification support its validity, the weak
performance of the growth-rate version cautions that the results could
be sensitive to issues of persistence and trending in the data. Taken
together, these results suggest that while the level specification
provides strong identification and compelling evidence of a positive
causal effect of local wages on education spending, the weak performance
of the growth-rate specification highlights the need for caution, as the
strength of the findings may partly reflect long-run trending
relationships rather than purely exogenous shocks. However, examining
the structure of the growth rate shock, the instability of the variable
is likely causing the poor identification in the growth rate
regressions.

```{r, ref.label='unnamed-chunk-93', echo = FALSE, eval=TRUE, results = 'asis'}
```

### Baseline Models for Growth Rate Sample Partitioning



```{r, ref.label='unnamed-chunk-94', echo = FALSE, eval=TRUE, results = 'asis'}
```

### Wage-based SS Instrument for Growth Rate Sample Partitioning

When applying our instrumental variable design we find that majority of
our relevant signal is picked up in regions across both trend metrics. Though the effects are more consistent across wage-trending sub-samples.

```{r, ref.label='unnamed-chunk-95', echo = FALSE, eval=TRUE, results = 'asis'}
```

### Wage-based SS State by State Estimation



```{r, ref.label='unnamed-chunk-96', echo = FALSE, eval=TRUE, results = 'asis'}
```

#### Wage-based SS in Industry-by-Industry Estimation

Using our wage-based shift share instrument, the below demonstrates the
overall treatment effect of local wage changes instrumented via our
wage-based shift-share shock.

```{r, ref.label='unnamed-chunk-97', echo = FALSE, eval=TRUE, results = 'asis'}
```

### High-Income Outliers

There is a somewhat non-linear relationship between property taxes and
elementary expenditure as property taxes collected rise as represented
in Figure 22 below. This happens largely as a result of very high-income
commuting zones. Therefore, we exclude any commuting zone that spends
more than 22k per pupil to avoid any distorting effects. This removes 12
CZs (\~2% of the sample). \textcolor{red}{This could benefit from
more robust outlier detection.}
To assess whether the main results are driven by a small number of very
high-income jurisdictions, I re-estimate the baseline and IV
specifications excluding such outliers. The findings are fully
consistent with the baseline analysis: house prices remain a strong
predictor of local education spending, and the IV estimates continue to
imply that a 10% increase in house prices raises per-pupil expenditure
by roughly 4–6%. The wage-based shift-share instrument yields somewhat
larger point estimates, though with wider standard errors, while the
VA-based instrument produces effects in line with earlier results.
Overall, this robustness exercise confirms that the causal relationship
between housing wealth and education spending is not confined to
affluent areas but reflects a broader, generalizable pattern. These
findings confirm that the main result is not driven solely by affluent
jurisdictions, but reflects a more general relationship between local
housing wealth and education spending.

```{r, ref.label='unnamed-chunk-98', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Panel VAR Specification

$$
Y_{it} = \alpha_i +  \sum_{k = 1}^{4} \gamma_{k}A_{i,t-k} + \beta X_{it} + \varepsilon_{it}
$$

Where we approach a level and per capita value expression of the
relationship between total education expenditure, intergovernmental
revenue, house prices conditioned on GDP and wage levels.

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Total Educ. Exp.})_{it} \\
\log(\text{real Total IG Revenue})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP})_{it} \\
\log(\text{wage})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $3 \times 3$ coefficient matrices
- $\beta$ is a $3 \times 2$ matrix of coefficients on the exogenous variables
- $\alpha_i$ is a vector of unit fixed effects
- $\varepsilon_{it}$ is the error term

Where

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Own Source Rev. per person})_{it} \\
\log(\text{real IG Revenue per person})_{it} \\
\log(\text{wage})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP per capita})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $4 \times 4$ coefficient matrices
- $B$ is a $4 \times 1$ coefficient matrix
- $\alpha_i$ unit fixed effects
- $\varepsilon_{it}$ error term

```{r, ref.label='unnamed-chunk-99', echo = FALSE, eval=TRUE, results = 'asis'}
```

