# Appendices {-}

```{=latex}
\appendix
```


```{=latex}
\input{/Users/ebbamark/Documents/Documents - Nuff-Malham/GitHub/pub_fin_schools/code/si_tex_docs/si.tex}
```

# Key Relationships between Economic Variables

Below we display key relationships between several of the economic variables in our study. All economic indicators (house prices, private industry GDP, and wages are positively associated with education expenditure, whereas enrollment is negatively associated with education expenditure per pupil.


We report descriptive regressions in this section to establish relationships between property taxes, education expenditure, private industry GDP, and house prices. All regression models are estimated using two-way fixed effects and standard errors clustered by commuting zone.

```{r, ref.label='KRs', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Property Tax ~ GDP

GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) is assocaited with a 0.38% (0.32%) increase in property taxes collected (per capita) (\autoref{tbl_proptax_gdp}).

```{r, ref.label='appendix2', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Education Expenditure ~ Revenue Sources

The below regressions are included to establish the relationship between education expenditure and its component parts (ie. the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).

```{r, ref.label='appendix3', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

# Baseline Descriptive Estimates

We descriptively establish foundational relationships between
local economic conditions that seem to reasonably generalise across the
country using an AR(1) two-way fixed effects ordinary least-squares
panel model with standard errors clustered by commuting zone. We outline
the model specification immediately below:


$$
Y_{it} = \beta_{0} + \beta_{x}X_{it} + \theta Y_{it-1} + \delta_1 Enrollment_{it} + \delta_2 IGR_{it} + \delta_3 Black_{it} + \delta_3 Hispanic_{it} + \alpha_i + \gamma_t + \varepsilon_{it}
$$ {#eq-desc_function}


$Y_{it}$ is the natural logarithm of elementary (serving ages 6-12)
education expenditure per pupil for CZ $i$ in year $t$.
$\alpha_i$ represents a CZ fixed effect and $\gamma_t$ represents
year-fixed effects, respectively. $\varepsilon_{it}$ represents the
error term. We control for enrollment to account for scaling factors in
education expenditure and intergovernmental transfers to account for the
significant role of such transfers in funding education expenditure.
$X_{it}$ takes three forms represented by @eq-gdp-function,
@eq-wage-function, @eq-hpi-function where $h$ represents $h$-year time
lags. We estimate all equations in levels and growth rates.


$$
X_{it}^{GDP} = \sum_{h=0}^{2} \beta_h^{\text{GDP}} \log(\text{GDP}_{i,t-h})
$$ {#eq-gdp-function}


$$
X_{it}^{Wage} = \sum_{h=0}^{2} \beta_h^{\text{Wage}} \log(\text{Wage}_{i,t-h})
$$ {#eq-wage-function}


$$
X_{it}^{HPI} = \sum_{h=0}^{2} \beta_h^{\text{HPI}} \log(\text{HPI}_{i,t-h})
$$ {#eq-hpi-function}


\autoref{tbl_desc_res_lev} reports the results from regressions of log
elementary education expenditure per pupil on contemporaneous and lagged
measures of local economic activity. \autoref{tbl_desc_res_gr} presents
the analogous specifications using annualized growth rates to capture
short-run dynamics. The estimates in \autoref{tbl_desc_res_lev} show
that per-pupil education spending is systematically higher in commuting
zones with 'stronger' local economies measured in local wages,
industrial GDP, and house prices.


Lagged economic indicators, particularly private industry GDP and
average weekly wages, are positively and significantly associated with
education spending. In the baseline specification (Column 1), the
elasticity of education spending with respect to local GDP per capita
(private industry) is positive and statistically significant once lagged
values are included.


A ten-percent increase in local GDP per capita two years prior is
associated with roughly a 1% increase in current education expenditure
per pupil, suggesting that fiscal capacity effects unfold gradually over
time. In the case of industry GDP, the magnitude of the coefficients
increases with the number of lags, suggesting a gradual adjustment
process by which local economic growth translates into higher public
investment in education over time. For example, a 10% increase in lagged
(t–2) real private GDP per capita is associated with a 0.6% increase in
per-pupil spending. The house price index also enters positively and
significantly but only contemporaneously reflecting the immediate link
between house prices and the property taxes that fund public education.
This points to the fundamental relationship between community asset
wealth and public education expenditure.


Intergovernmental revenue per pupil emerges as the strongest and most
consistent predictor of education expenditure after the auto-regressive
coefficient. A 10% increase in intergovernmental transfers is associated
with approximately a 2% increase in per-pupil education spending,
controlling for CZ and year fixed effects. This finding highlights the
importance of state and federal aid in sustaining local education
budgets.


The growth rate regressions, while explaining less variance overall,
largely confirm the patterns observed in the level specifications.
Intergovernmental revenue growth remains a strong and highly significant
determinant of education expenditure growth. Lagged wage and GDP growth also emerge as important
predictors, particularly at longer lags. Notably, wage growth two years
prior is associated with a 0.31% increase in education spending growth,
suggesting that labor market improvements take at least a year to
materialize in local education budgets hinting at the relevance of our
primary identifying relationship.


Taken together, these results offer three key insights. First, public
education investment is strongly mediated by external fiscal flows,
reaffirming the role of intergovernmental transfers in equalizing local
education finance. Second, local labor market conditions, captured
through wages and GDP, exert lagged effects on education spending
consistent with lagged effects of local economic conditions to
industrial change. Third, local housing markets play a significant role
shaping education budgets, reflecting the link between property values
and tax revenues which respond contemporaneously as a result of the
direct mechanical link between property values and local public revenue
generation.


Additionally, in both levels and growth rates, the consistently negative
coefficient on enrollment indicates a scaling relationship in which
expenditure per pupil declines as enrollment sizes grow.



```{r, ref.label='baseline', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Including State Fixed Effects

Regressions establishing baseline relationships between local economic variables and elementary education expenditure using state-fixed effects rather than commuting-zone level effects.

```{r, ref.label='baseline_regs_state_fe', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Including State Funding Share Interaction Term

Furthermore, given the heterogeneity in reliance on intergovernmental
transfers (largely coming from the state), we interact all economic
predictors above with a variable that represents the share of total
elementary education expenditure (as a continuous variable) coming from
state-level funding.



```{r, ref.label='funding_share_interaction', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

# Shift-Share Instrument

The following section provides further detail on the shift-share instrument as well as supplementary regressions estimated using a wage-based shift share instrument rather than the value-added based shift-share instrument in the main text. We use the abbreviation `SS` to refer to the shift-share instrument.

```{r, ref.label='appendix_industry_shares_display_header', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Data Inputs to the SS Instrument

Plots of the data inputs to the shift-share instrument. This inputs include national and county-level employment and average weekly wage levels by 2-digit NAICS industry code.

```{r, ref.label='appendix_industry_shares_display', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Results Using the Wage-based Shift-Share Instrument

First, we display the results for a 2SLS estimation using our wage-based
shift-share instrument.


The instrumental variable estimates provide similar evidence of a
robust causal relationship between local wages and public education
expenditure. Utilising our wage-based shift-share instrument we see
highly significant and relevant first-stage relationships when the
shift-share instrument is imposed in levels (except in column 5). In
each case except columns 5-6 (l1 SS, CZ FE), the first-stage regression
yields a statistically significant and economically large coefficient. Varying the
time-lag and inclusion of state or commuting zone fixed effects, we see
that a 10% increase in the shift-share measure (which can be interpreted
as a natural logarithm) is associated with a 2-6% increase in
average weekly wages (p \< 0.01), with an F-statistic well above conventional weak instrument thresholds confirming
instrument relevance. The Wu-Hausman tests reject the null of
exogeneity, confirming that OLS estimates are biased and IV estimation
is appropriate. Wald tests of joint significance further support the
strength of the instruments.


Using wage shocks in levels yields strong instruments, high first-stage
F-statistics, and stable second-stage estimates: higher local wages
robustly increase education spending. Furthermore, given the dependent
variable measures per pupil expenditure, this result implies direct
effects in experience per student. In contrast, when shocks are measured
in growth rates, though the F-statistic remains strong, the
second-stage coefficients become statistically insignificant. This suggests that the
growth-rate specification is poorly identified and cannot provide
reliable causal inference, whereas the level specification produces
credible and consistent results.


The specification that uses wage shocks in levels provides the most
credible identification strategy. Since levels capture the
cross-sectional fiscal variation that drives differences in property
values and school spending, the level specification is more consistent
with the economic mechanisms of interest and delivers more reliable
causal estimates. Simultaneously, the weakness of the growth-rate
specification does raise concerns about the robustness of the results.
If the relationship between wages, house prices, and education
expenditure is driven by common non-stationary trends, then regressions
in levels risk spurious correlation. The fact that the IV design loses
power when variables are differenced into growth rates may suggest that
part of the strong level results reflect long-run trends rather than
short-run causal shocks. However, examining
the structure of the growth rate shock, the instability of the variable
is likely causing the poor identification in the growth rate
regressions.

```{r, ref.label='wage_ss_res', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

# Growth Rate Sample Partitioning

In the following section, we provide supplementary information and econometric estimation details and results not included in the main text pertaining to the sample partitioning exercise by commuting zone level growth rates.

```{r, ref.label='growth_rate_partitioning_extra', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Baseline Models for Growth Rate Sample Partitioning



```{r, ref.label='baseline_growth_rate_partitioning', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Wage-based SS Instrument for Growth Rate Sample Partitioning

When applying our instrumental variable design we find that the signal is consistent across various GDP growth trajectories, but the relationship weakens as communities are defined by higher rates of wage growth.

```{r, ref.label='wage_ss_growth_rate_partitioning', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

# State-by-State Estimation

In the following section, we provide additional detail on the estimated state-by-state models outlined in the main text. We also provide descriptive regression results as benchmarks for comparison with the instrumental variable approach.

```{r, ref.label='state_by_state_descriptive_sm', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Wage Effect Using Descriptive Estimation Strategy

We report the effect of a 1% increase in Wage on education expenditure per pupil as defined in @eq-desc_function where $X_{i,t}$ takes on the values in @eq-wage-function. First, we re-employ our baseline regression in which we regress
our outcome variable on wage levels at time $t-h$ where $h = [0,2]$. The
plot below displays the cumulative effect of a 1% increase in wage
levels on education expenditure per pupil in purple. The X marks
represent the individual coefficients on each time lag of the treatment
variable, the linear combination of which form the total dynamic effect
represented by the purple dots. Majority of states see positive
elasticities in relation to wage changes in this descriptive
specification, but some (South Carolina, Missouri, Kentucky, Washington) see
negative elasticities.

```{r, ref.label='state_by_state_descriptive', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Instrumental Variable versus Descriptive Relationship

The below plots displays the descriptive regression estimate compared to the IV fitted regression coefficient, describing the effect of a 1% increase in local wages on education expenditure per pupil. Majority of states find at least a coefficient of equal sign. Interestingly though, the valid and statistically significant instrumental variable coefficients only ever identify positive elasticities.

```{r, ref.label='desc_v_iv', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

# Industry-by-Industry Estimation

In the following section, we provide additional detail about the estimated industry-by-industry models in the main text.

```{r, ref.label='wage_ss_ind_by_ind_sm', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Wage-based SS in Industry-by-Industry Estimation

Using our wage-based shift share instrument, the below demonstrates the
overall treatment effect of local wage changes instrumented via our
wage-based shift-share shock.

```{r, ref.label='wage_ss_ind_by_ind', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

# Robustness Checks and Additional Analyses

In the following section, we report various robustness checks and additional analyses conducted to illuminate patterns in the data and additional heterogeneity. These include additional panel econometric estimations via quantile regressions, vector auto regressions, and some outlier testing.

```{r, ref.label='sm_inventory', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Quantile Regression

We perform a comparative quantile regression, where commuting zones are grouped by their levels of per pupil expenditure. We group them into the 10th, 25th, median, 75th, and 90th percentiles. We find that all coefficients are relatively stable apart from the relationship between per pupil expenditure and race, where the regression coefficient notably switches sign from negative to positive as the level of expenditure increases. Additionally, examining the central relationship of interest in this work, we find that the effect of wages on expenditure is significant contemporaneously in those commuting zones near the median level of expenditure, whereas those spending a relatively high amount of resources on local education, see a delayed effect.

```{r, ref.label='quantile_reg', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## High-Income Outliers

There is a somewhat non-linear relationship between property taxes and
elementary expenditure as property taxes collected rise as represented
in the figure below. This happens largely as a result of very high-income
commuting zones. Therefore, we exclude any commuting zone that spends
more than 22k per pupil to avoid any distorting effects. This removes 12
CZs (\~2% of the sample).
To assess whether the main results are driven by a small number of very
high-income jurisdictions, I re-estimate the baseline and IV
specifications excluding such outliers. The findings are fully
consistent with the baseline analysis: house prices remain a strong
predictor of local education spending, and the IV estimates continue to
imply that a 10% increase in house prices raises per-pupil expenditure
by roughly 4–6%. The wage-based shift-share instrument yields somewhat
larger point estimates, though with wider standard errors, while the
VA-based instrument produces effects in line with earlier results.
Overall, this robustness exercise confirms that the causal relationship
between housing wealth and education spending is not confined to
affluent areas but reflects a broader, generalizable pattern. These
findings confirm that the main result is not driven solely by affluent
jurisdictions, but reflects a more general relationship between local
housing wealth and education spending.

```{r, ref.label='high_income_outliers', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```

## Panel VAR Specification

$$
Y_{it} = \alpha_i +  \sum_{k = 1}^{4} \gamma_{k}A_{i,t-k} + \beta X_{it} + \varepsilon_{it}
$$

Where we approach a level and per capita value expression of the
relationship between total education expenditure, intergovernmental
revenue, house prices conditioned on GDP and wage levels.

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Total Educ. Exp.})_{it} \\
\log(\text{real Total IG Revenue})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP})_{it} \\
\log(\text{wage})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $3 \times 3$ coefficient matrices
- $\beta$ is a $3 \times 2$ matrix of coefficients on the exogenous variables
- $\alpha_i$ is a vector of unit fixed effects
- $\varepsilon_{it}$ is the error term

Where

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Own Source Rev. per person})_{it} \\
\log(\text{real IG Revenue per person})_{it} \\
\log(\text{wage})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP per capita})_{it}
\end{bmatrix}
$$

- $A_1, A_2, A_3, A_4$ are $4 \times 4$ coefficient matrices
- $B$ is a $4 \times 1$ coefficient matrix
- $\alpha_i$ unit fixed effects
- $\varepsilon_{it}$ error term

```{r, ref.label='appendix_pvar', echo = FALSE, eval=TRUE, results = 'asis'}
```

```{=latex}
\FloatBarrier
```


```{=latex}
\clearpage
```

