# Appendix {.unnumbered}

```{=latex}
\appendix
```

# Descriptive Regression Results

In the following set of results, I report descriptive regressions to establish relationships between property taxes, education expenditure, GDP (total, private industry, O&G&mining), etc.
All regression models that follow include TWFE (CZ- and year- fixed effects) and standard errors clustered by commuting zone.
All functional forms in the feols() functions below are of the form Y ~ X In the cases in which multiple estimations are included via sw(Xa, Xb, Xc + Xd), the function will return results for Y~Xa, Y~Xb, Y ~ Xc + Xd.

```{r, ref.label='appendix1', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Property Tax ~ GDP

GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) leads to a 0.38% (0.32%) increase in property taxes collected (per capita).

```{r, ref.label='appendix2', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ Revenue Sources

The below regressions are included to establish the relationship between education expenditure and its component parts. These regressions simply corroborate what is displayed in the section on Key Relationships in [LINK](https://ebbam.github.io/pub_fin_schools/code/cz_results.html) (ie. that the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).

```{r, ref.label='appendix3', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ GDP

A 1% increase in GDP pc is associated with a 0.19% increase in education expenditure per pupil, dominated by the effect of GDP from private industry (0.16%). I include here also the GDP generated from the oil, gas, mining, and quarrying sector. The effect is small and statistically insignificant.

```{r, ref.label='appendix4', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Groups

Industry-level shift-share_instrument

```{r, ref.label='appendix_ss', echo = FALSE, eval=TRUE, results = 'asis'}
```

# Descriptive Regression Results

In the following set of results, I report descriptive regressions to establish relationships between property taxes, education expenditure, GDP (total, private industry, O&G&mining), etc.
All regression models that follow include TWFE (CZ- and year- fixed effects) and standard errors clustered by commuting zone.
All functional forms in the feols() functions below are of the form Y ~ X In the cases in which multiple estimations are included via sw(Xa, Xb, Xc + Xd), the function will return results for Y~Xa, Y~Xb, Y ~ Xc + Xd.

```{r, ref.label='unnamed-chunk-51', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Property Tax ~ GDP

GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) leads to a 0.38% (0.32%) increase in property taxes collected (per capita).

```{r, ref.label='unnamed-chunk-52', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ Revenue Sources

The below regressions are included to establish the relationship between education expenditure and its component parts. These regressions simply corroborate what is displayed in the section on Key Relationships in [LINK](https://ebbam.github.io/pub_fin_schools/code/cz_results.html) (ie. that the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).

```{r, ref.label='unnamed-chunk-53', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Education Expenditure ~ GDP

A 1% increase in GDP pc is associated with a 0.19% increase in education expenditure per pupil, dominated by the effect of GDP from private industry (0.16%). I include here also the GDP generated from the oil, gas, mining, and quarrying sector. The effect is small and statistically insignificant.

```{r, ref.label='unnamed-chunk-54', echo = FALSE, eval=TRUE, results = 'asis'}
```

## Groups

Industry-level shift-share_instrument

```{r, ref.label='unnamed-chunk-55', echo = FALSE, eval=TRUE, results = 'asis'}
```

