---
title: "State Analysis"
format: html
date: "2026-03-24"
bibliography: references.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 16, fig.width = 14, cache = FALSE, message = FALSE, warning = FALSE, results = 'asis')

```

## R Markdown

```{r cars, cache = FALSE}

library(here)
library(tidyverse)
library(readxl)
library(usmap)
library(fixest)
library(kableExtra)
library(purrr)
library(ggrepel)
library(readxl)
library(readxl)
source(here('code/source_code/useful_functions.R'))
source(here('code/source_code/dicts.R'))
results_df <- readRDS(here("data/temp/results_df.RDS"))
ss_temp_filled <- readRDS(here("code/ss_cache_manual/ss_temp_filled.RDS"))

results_df_cleaned_iv <- readRDS(here("data/temp/results_df_cleaned_iv.RDS"))


sig_states <- unique(results_df_cleaned_iv$state)
# Saving a backup just in case
sig_states_temp <- sig_states
results_df <- readRDS(here("data/temp/results_df.RDS"))
reg_states <- readRDS(here("data/temp/reg_states.RDS"))
reg_states_temp <- setdiff(reg_states, sig_states_temp)
df_ivs <- readRDS(here("data/temp/df_ivs.RDS"))
cz_names <- readRDS(here("data/out/cz_w_county_names.RDS"))
school_dist_info <- readRDS(here("data/raw/state_school_district_info/raw_state_school_dist_info.RDS"))
```

```{r}

# Industry codes and names
industry_mapping <- tribble(
  ~industry_code, ~industry_name, ~category, ~broad_category,
  "11", "Agriculture, Forestry, Fishing, and Hunting", "Agriculture", "Primary",
  "21", "Mining", "Natural Resources", "Primary",
  "22", "Utilities", "Utilities", "Utilities",
  "23", "Construction", "Construction", "Construction",
  "31_33", "Manufacturing", "Manufacturing", "Manufacturing",
  "42", "Wholesale Trade", "Wholesale Trade", "Trade",
  "44_45", "Retail Trade", "Retail Trade", "Trade",
  "48_49", "Transportation and Warehousing", "Transportation", "Transportation",
  "51", "Information", "Information", "Information",
  "52", "Finance and Insurance", "Finance", "Finance",
  "53", "Real Estate and Rental and Leasing", "Real Estate", "Real Estate",
  "54", "Professional, Scientific, and Technical Services", "Professional Services", "Services",
  "55", "Management of Companies and Enterprises", "Management", "Services",
  "56", "Administrative Support and Waste Management", "Administrative", "Services",
  "61", "Educational Services", "Education", "Public Services",
  "62", "Health Care and Social Assistance", "Healthcare", "Public Services",
  "71", "Arts, Entertainment, and Recreation", "Arts/Entertainment", "Services",
  "72", "Accommodation and Food Services", "Hospitality", "Services",
  "81", "Other Services (except Public Administration)", "Other Services", "Services",
  "92", "Public Administration", "Public Admin", "Public Services"
)

# ── Quadrant plot: Instrument strength vs. Second-stage significance ──────────
# Requires: results_df, f_stat_threshold (= 12), ggrepel
f_stat_threshold = 12
# Recover t-statistic from stored CIs
quad_data <- results_df %>%
  mutate(
    std_error = (ci_upper - ci_lower) / (2 * 1.96),
    t_stat    = estimate / std_error,
    quadrant  = case_when(
      ftest >= f_stat_threshold & p_val < 0.1 & estimate > 0 ~ "Strong IV + Significant",
      ftest >= f_stat_threshold & (p_val >= 0.1 | estimate <= 0) ~ "Strong IV + Non-significant",
      ftest <  f_stat_threshold & p_val >= 0.1                   ~ "Weak IV + Non-significant",
      ftest <  f_stat_threshold & p_val  < 0.1                   ~ "Weak IV + Significant"
    ),
    quadrant = factor(quadrant, levels = c(
      "Strong IV + Significant",
      "Strong IV + Non-significant",
      "Weak IV + Non-significant",
      "Weak IV + Significant"
    ))
  )

t_thresh <- 1.645   # 10% one-sided, matching p_val < 0.1 criterion
f_thresh <- f_stat_threshold

# X-axis limits for annotation positioning (log scale)
x_min  <- min(quad_data$ftest, na.rm = TRUE) * 0.8
x_max  <- max(quad_data$ftest, na.rm = TRUE) * 1.2
y_min  <- min(quad_data$t_stat, na.rm = TRUE)
y_max  <- max(quad_data$t_stat, na.rm = TRUE)

fig_quad <- ggplot(quad_data, aes(x = ftest, y = t_stat)) +

  # Quadrant shading (use finite xmin/xmax for compatibility with log scale)
  annotate("rect",
           xmin = f_thresh, xmax = x_max, ymin = t_thresh, ymax = y_max * 1.1,
           fill = "#d4edda", alpha = 0.5) +
  annotate("rect",
           xmin = f_thresh, xmax = x_max, ymin = y_min * 1.1, ymax = t_thresh,
           fill = "#fff3cd", alpha = 0.5) +
  annotate("rect",
           xmin = x_min, xmax = f_thresh, ymin = y_min * 1.1, ymax = t_thresh,
           fill = "#f8d7da", alpha = 0.4) +
  annotate("rect",
           xmin = x_min, xmax = f_thresh, ymin = t_thresh, ymax = y_max * 1.1,
           fill = "#cce5ff", alpha = 0.4) +

  # Reference lines
  geom_vline(xintercept = f_thresh, linetype = "dashed", colour = "grey40", linewidth = 0.6) +
  geom_hline(yintercept = t_thresh, linetype = "dashed", colour = "grey40", linewidth = 0.6) +
  geom_hline(yintercept = 0,        linetype = "dotted", colour = "grey60", linewidth = 0.4) +

  # Points
  geom_point(aes(colour = quadrant), size = 3.5, alpha = 0.9) +

  # State labels
  geom_text_repel(
    aes(label = state_name, colour = quadrant),
    size          = 3,
    fontface      = "bold",
    box.padding   = 0.4,
    max.overlaps  = 30,
    segment.size  = 0.3,
    segment.color = "grey60"
  ) +

  # Quadrant annotation labels
  annotate("text",
           x = x_max * 0.95, y = y_max * 1.0,
           label    = "Strong instrument\nIntact fiscal channel",
           hjust    = 1, vjust = 1, size = 3.2,
           colour   = "#1a7a3c", fontface = "italic") +
  annotate("text",
           x = x_max * 0.95, y = y_min * 0.95,
           label    = "Strong instrument\nBroken fiscal channel\n(equalization / centralization)",
           hjust    = 1, vjust = 0, size = 3.2,
           colour   = "#b07d00", fontface = "italic") +
  annotate("text",
           x = x_min * 1.05, y = y_min * 0.95,
           label    = "Weak instrument\nDoubly disconnected",
           hjust    = 0, vjust = 0, size = 3.2,
           colour   = "#9e2a2b", fontface = "italic") +
  annotate("text",
           x = x_min * 1.05, y = y_max * 1.0,
           label    = "Weak instrument\nUnreliable result",
           hjust    = 0, vjust = 1, size = 3.2,
           colour   = "#1a4f9c", fontface = "italic") +

  # Scales
  scale_colour_manual(
    values = c(
      "Strong IV + Significant"     = "#1a7a3c",
      "Strong IV + Non-significant" = "#b07d00",
      "Weak IV + Non-significant"   = "#9e2a2b",
      "Weak IV + Significant"       = "#1a4f9c"
    ),
    guide = "none"
  ) +
  scale_x_continuous(
    trans  = "log10",
    labels = scales::comma
  ) +

  labs(
    title    = "Instrument Strength vs. Second-Stage Significance by State",
    subtitle = str_wrap(
      "x-axis (log scale): first-stage F-statistic — how well the shift-share instrument predicts local wages.
       y-axis: t-statistic of the IV wage coefficient on public education spending.",
      width = 100),
    x       = paste0("First-stage F-statistic (log scale, threshold = ", f_thresh, ")"),
    y       = "t-statistic: IV wage coefficient on spending",
    caption = paste0(
      "Dashed vertical line = F = ", f_thresh, " (weak instrument threshold). ",
      "Dashed horizontal line = t = 1.645 (10% significance). ",
      "Oregon excluded from main IV results (sign inconsistency)."
    )
  ) +

  theme_minimal(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 10, colour = "grey30", margin = margin(b = 10)),
    plot.caption  = element_text(size = 8,  colour = "grey50", hjust = 0),
    panel.grid.minor = element_blank()
  )

print(fig_quad)
ggsave(here("output/fig_quadrant_iv_strength.png"), fig_quad,
       width = 12, height = 8, dpi = 300)

# ── 0. Load data ──────────────────────────────────────────────────────────────
df_ivs <- readRDS(here("data/temp/df_ivs_lags_2.RDS"))

# ── 1. State-level instrument dispersion (from regression data) ───────────────
instrument_labels <- tribble(
  ~state, ~instrument_strength_table,
  "ND", "High",  "LA", "High",  "OK", "High",  "TX", "High",  "WY", "High",
  "PA", "High",  "NM", "High",  "CO", "Moderate", "KY", "Moderate",
  "MT", "Moderate", "WV", "Moderate", "WA", "Moderate", "CA", "Moderate",
  "NY", "Moderate", "FL", "Moderate", "IN", "Moderate", "MI", "Moderate",
  "OH", "Moderate", "IL", "Moderate", "MN", "Moderate", "OR", "Moderate",
  "SD", "Low",   "AZ", "Low",   "AR", "Low",   "AL", "Low",   "GA", "Low",
  "ID", "Low",   "IA", "Low",   "KS", "Low",   "MS", "Low",   "MO", "Low",
  "NE", "Low",   "SC", "Low",   "UT", "Low",   "WI", "Low"
)

ss_variance <- df_ivs %>%
  filter(state %in% reg_states, !is.na(lev_gdp_ss_2d)) %>%
  group_by(state, year) %>%
  summarize(sd_cross_cz = sd(lev_gdp_ss_2d, na.rm = TRUE), n_czs = n(), .groups = "drop") %>%
  group_by(state) %>%
  summarize(mean_sd_cross_cz = mean(sd_cross_cz, na.rm = TRUE),
            mean_n_czs       = mean(n_czs), .groups = "drop") %>%
  left_join(instrument_labels, by = "state") %>%
  left_join(results_df %>% select(state, ftest, estimate, p_val) %>%
              group_by(state) %>% slice(1), by = "state") %>%
  mutate(
    state_name = sapply(state, get_state),
    sig_state  = state %in% sig_states,
    instrument_strength_table = factor(instrument_strength_table,
                                       levels = c("High", "Moderate", "Low"))
  ) %>%
  arrange(desc(mean_sd_cross_cz))

# ── FIGURE 1: Ranked bar chart of instrument dispersion ──────────────────────
fig_ss_var_1 <- ggplot(
  ss_variance,
  aes(x = fct_reorder(state_name, mean_sd_cross_cz), y = mean_sd_cross_cz,
      fill = sig_state)
) +
  geom_col(alpha = 0.85) +
  coord_flip() +
  scale_fill_manual(
    values = c("TRUE" = "#1a7a3c", "FALSE" = "grey70"),
    labels = c("TRUE" = "Included in main IV results", "FALSE" = "Not significant"),
    name   = NULL
  ) +
  labs(
    title    = "Cross-CZ Dispersion of Shift-Share Instrument by State",
    subtitle = str_wrap("Mean within-year SD of the shift-share instrument across CZs.
       Higher values indicate more heterogeneous industry exposure across
       CZs within the state. Green = state included in main IV results.", 90),
    x = NULL, y = "Mean cross-CZ SD of shift-share instrument",
    caption = "Instrument: lev_gdp_ss_2d. SD computed across CZs within each state-year, then averaged."
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, colour = "grey30"),
        legend.position = "top", panel.grid.minor = element_blank())

print(fig_ss_var_1)
ggsave(here("output/fig_ss_instrument_variance_simple.png"),
       fig_ss_var_1, width = 10, height = 10, dpi = 300)

# ── 2. Industry decomposition inputs — computed ONCE ─────────────────────────
share_cols   <- grep("^share_annual_avg_emplvl_",       names(ss_temp_filled), value = TRUE)
lwage_cols   <- grep("^log_natl_annual_avg_wkly_wage_", names(ss_temp_filled), value = TRUE)
strip_pfx    <- function(x, pfx) str_remove(x, paste0("^", pfx))
common_codes <- intersect(
  strip_pfx(share_cols, "share_annual_avg_emplvl_"),
  strip_pfx(lwage_cols, "log_natl_annual_avg_wkly_wage_")
)

cz_state_map <- df_ivs %>%
  select(unit, state) %>% distinct() %>% rename(cz_id = unit)

base_shares <- ss_temp_filled %>%
  group_by(cz_id) %>% filter(year == min(year)) %>% ungroup() %>%
  select(cz_id, all_of(paste0("share_annual_avg_emplvl_", common_codes))) %>%
  left_join(cz_state_map, by = "cz_id") %>%
  filter(state %in% reg_states)

natl_wage_growth <- ss_temp_filled %>%
  select(cz_id, year, all_of(paste0("log_natl_annual_avg_wkly_wage_", common_codes))) %>%
  group_by(cz_id) %>% arrange(year) %>%
  summarize(across(starts_with("log_natl_annual_avg_wkly_wage_"),
                   ~ last(na.omit(.x)) - first(na.omit(.x))), .groups = "drop") %>%
  summarize(across(-cz_id, ~ mean(.x, na.rm = TRUE))) %>%
  pivot_longer(everything(), names_to = "var", values_to = "g_k") %>%
  mutate(ind_code = strip_pfx(var, "log_natl_annual_avg_wkly_wage_")) %>%
  select(ind_code, g_k)

# g_k² × Var(z_{l,k}) per state × industry — ss_variance metadata joined here once
share_var <- base_shares %>%
  pivot_longer(cols = starts_with("share_annual_avg_emplvl_"),
               names_to = "var", values_to = "share") %>%
  mutate(ind_code = strip_pfx(var, "share_annual_avg_emplvl_")) %>%
  group_by(state, ind_code) %>%
  summarize(var_share = var(share, na.rm = TRUE), .groups = "drop") %>%
  left_join(natl_wage_growth, by = "ind_code") %>%
  mutate(ind_contribution = g_k^2 * var_share) %>%
  left_join(ss_variance %>% select(state, mean_sd_cross_cz, sig_state, state_name),
            by = "state") %>%
  mutate(state_name = fct_reorder(state_name, mean_sd_cross_cz))

# ── FIGURE 2: Top-8 industry decomposition ────────────────────────────────────
top_inds <- share_var %>%
  group_by(ind_code) %>%
  summarize(total = sum(ind_contribution, na.rm = TRUE), .groups = "drop") %>%
  slice_max(total, n = 8) %>% pull(ind_code)

ind_levels_fig2 <- c(
  share_var %>%
    filter(ind_code %in% top_inds) %>%
    group_by(ind_code) %>%
    summarize(t = sum(ind_contribution), .groups = "drop") %>%
    left_join(industry_mapping %>% select(industry_code, industry_name),
              by = c("ind_code" = "industry_code")) %>%
    arrange(desc(t)) %>% pull(industry_name),
  "Other"
)

variance_decomp_top8 <- share_var %>%
  mutate(ind_label = if_else(ind_code %in% top_inds, ind_code, "Other")) %>%
  group_by(state, state_name, mean_sd_cross_cz, sig_state, ind_label) %>%
  summarize(ind_contribution = sum(ind_contribution, na.rm = TRUE), .groups = "drop") %>%
  group_by(state) %>%
  mutate(prop = ind_contribution / sum(ind_contribution, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(industry_mapping %>% select(industry_code, industry_name),
            by = c("ind_label" = "industry_code")) %>%
  mutate(
    ind_label_named = case_when(
      ind_label == "Other"        ~ "Other",
      !is.na(industry_name)       ~ industry_name,
      TRUE                        ~ paste0("Industry ", ind_label)
    ),
    ind_label_named = factor(ind_label_named, levels = ind_levels_fig2)
  )

ind_pal_fig2 <- setNames(
  c("#e41a1c","#377eb8","#4daf4a","#984ea3",
    "#ff7f00","#a65628","#f781bf","#999999","grey85"),
  c(ind_levels_fig2[ind_levels_fig2 != "Other"], "Other")
)

fig_ss_var_2 <- ggplot(variance_decomp_top8,
                       aes(x = state_name, y = prop * mean_sd_cross_cz,
                           fill = ind_label_named)) +
  geom_col(alpha = 0.85) +
  geom_col(
    data        = ss_variance %>% filter(sig_state),
    aes(x = fct_reorder(state_name, mean_sd_cross_cz), y = mean_sd_cross_cz),
    fill = NA, color = "black", linewidth = 0.9, inherit.aes = FALSE
  ) +
  coord_flip() +
  scale_fill_manual(values = ind_pal_fig2, name = "Industry") +
  labs(
    title    = "Cross-CZ Dispersion of Shift-Share Instrument: Industry Decomposition",
    subtitle = str_wrap("Bar length = mean cross-CZ SD of shift-share instrument. Segments show each
       industry's proportional contribution (g_k² × Var(z_{l,k})). Bold outline = significant IV state.", 100),
    x = NULL, y = "Mean cross-CZ SD of shift-share instrument",
    caption = "g_k² × Var(z_{l,k}) decomposition. Top 8 industries by total contribution; remainder = Other."
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", size = 13),
        plot.subtitle = element_text(size = 9, colour = "grey30"),
        plot.caption = element_text(size = 8, colour = "grey50", hjust = 0),
        legend.position = "right", panel.grid.minor = element_blank())

print(fig_ss_var_2)
ggsave(here("output/fig_ss_instrument_variance_decomp.png"),
       fig_ss_var_2, width = 13, height = 11, dpi = 300)

# ── FIGURE 3: Two-category (cyclical/commodity vs other) ─────────────────────
CYCLICAL_INDS <- c("11", "21", "22", "23", "31_33", "42", "48_49", "72")

group_props <- share_var %>%
  group_by(state) %>%
  mutate(prop = ind_contribution / sum(ind_contribution, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(ind_group = if_else(ind_code %in% CYCLICAL_INDS,
                             "Cyclical / commodity", "Other industries")) %>%
  group_by(state, state_name, mean_sd_cross_cz, sig_state, ind_group) %>%
  summarize(group_prop = sum(prop, na.rm = TRUE), .groups = "drop") %>%
  mutate(state_name = fct_reorder(state_name, mean_sd_cross_cz))

bar_labels <- group_props %>%
  filter(ind_group == "Cyclical / commodity") %>%
  mutate(label = scales::percent(group_prop, accuracy = 1))

fig_ss_var_3 <- ggplot(group_props,
                       aes(x = state_name, y = group_prop, fill = ind_group)) +
  geom_col(aes(color = sig_state), width = 0.75, position = "stack") +
  geom_text(data = bar_labels,
            aes(x = state_name, y = group_prop / 2, label = label),
            inherit.aes = FALSE, size = 2.8, color = "white", fontface = "bold") +
  scale_fill_manual(values = c("Cyclical / commodity" = "#e07b39",
                                "Other industries"     = "grey80"),
                    name = "Industry group") +
  scale_color_manual(values = c(`TRUE` = "#1a1a2e", `FALSE` = NA), guide = "none") +
  scale_y_continuous(labels = scales::percent_format(),
                     expand = expansion(mult = c(0, 0.03))) +
  coord_flip() +
  labs(
    title    = "Share of Shift-Share Instrument Variance Driven by Cyclical / Commodity Industries",
    subtitle = str_wrap("Each bar = one state's cross-CZ dispersion decomposed by industry.
       Cyclical/commodity = Agriculture, Mining, Manufacturing, Construction, Wholesale Trade,
       Transportation, Hospitality. States ordered by total instrument dispersion. Bold border = significant IV state.", 100),
    x = NULL, y = "Share of cross-CZ variance explained",
    caption = "Cyclical/commodity (NAICS 11, 21, 23, 31-33, 42, 48-49, 72) in amber; all other industries in grey."
  ) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(face = "bold", size = 12),
        plot.subtitle = element_text(size = 9, color = "grey30", margin = margin(b = 8)),
        legend.position = "bottom", panel.grid.minor = element_blank(),
        panel.grid.major.y = element_blank(),
        plot.caption = element_text(hjust = 0, size = 8, color = "grey50",
                                    margin = margin(t = 8)))

print(fig_ss_var_3)
ggsave(here("output/fig_ss_instrument_variance_cyclical.png"),
       fig_ss_var_3, width = 10, height = 9, dpi = 300)


# ── FIGURE 4: All industries, above-average employment share highlighted ──────

# Cross-state average employment share per industry
avg_shares_national <- base_shares %>%
  pivot_longer(cols = starts_with("share_annual_avg_emplvl_"),
               names_to = "var", values_to = "share") %>%
  mutate(ind_code = strip_pfx(var, "share_annual_avg_emplvl_")) %>%
  group_by(ind_code) %>%
  summarize(avg_share = mean(share, na.rm = TRUE), .groups = "drop")

# State-level mean employment share per industry
state_ind_shares <- base_shares %>%
  pivot_longer(cols = starts_with("share_annual_avg_emplvl_"),
               names_to = "var", values_to = "share") %>%
  mutate(ind_code = strip_pfx(var, "share_annual_avg_emplvl_")) %>%
  group_by(state, ind_code) %>%
  summarize(state_share = mean(share, na.rm = TRUE), .groups = "drop") %>%
  left_join(avg_shares_national, by = "ind_code") %>%
  mutate(above_avg = state_share > avg_share)

# Build full decomposition from share_var (all industries, no "Other" bucket)
# Factor order: largest total contribution at the bottom of each bar
ind_order_all <- share_var %>%
  group_by(ind_code) %>%
  summarize(total = sum(ind_contribution, na.rm = TRUE), .groups = "drop") %>%
  arrange(total) %>%
  pull(ind_code)

variance_decomp_all <- share_var %>%
  group_by(state) %>%
  mutate(prop = ind_contribution / sum(ind_contribution, na.rm = TRUE)) %>%
  ungroup() %>%
  left_join(
    industry_mapping %>% select(industry_code, category),
    by = c("ind_code" = "industry_code")
  ) %>%
  mutate(
    ind_label = coalesce(category, paste0("Ind. ", ind_code)),
    ind_code  = factor(ind_code, levels = ind_order_all),
    ind_label = fct_reorder(ind_label, as.integer(ind_code))
  ) %>%
  left_join(state_ind_shares %>% select(state, ind_code, above_avg),
            by = c("state", "ind_code")) %>%
  mutate(above_avg = coalesce(above_avg, FALSE))

# 19-colour palette keyed to industry label
ind_labels_ordered <- levels(variance_decomp_all$ind_label)
n_inds <- length(ind_labels_ordered)
ind_pal_all <- setNames(
  colorRampPalette(
    c("#e41a1c","#ff7f00","#ffff33","#4daf4a","#377eb8",
      "#984ea3","#a65628","#f781bf","#17becf","#1a9641")
  )(n_inds),
  ind_labels_ordered
)

fig_ss_var_4 <- ggplot(variance_decomp_all,
                       aes(x     = state_name,
                           y     = prop,
                           fill  = ind_label,
                           alpha = above_avg)) +
  geom_col() +
  geom_col(
    data        = ss_variance %>% filter(sig_state),
    aes(x = fct_reorder(state_name, mean_sd_cross_cz), y = 1),
    fill = NA, color = "black", linewidth = 0.9, inherit.aes = FALSE
  ) +
  coord_flip() +
  scale_fill_manual(values = ind_pal_all, name = "Industry") +
  scale_alpha_manual(
    values = c(`TRUE` = 1.0, `FALSE` = 0.15),
    guide  = "none"
  ) +
  labs(
    title    = "Cross-CZ Dispersion of Shift-Share Instrument: Industry Specialisation",
    subtitle = str_wrap(
      "Bar length = mean cross-CZ SD of shift-share instrument. Solid segments = industries where
       the state's mean CZ employment share exceeds the cross-state average. Faded = below average.
       Bold outline = state included in main IV results.",
      100
    ),
    x       = NULL,
    y       = "Mean cross-CZ SD of shift-share instrument",
    caption = "All industries shown. Above-average defined as state mean CZ share > cross-state mean for that industry."
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 9, colour = "grey30"),
    plot.caption     = element_text(size = 8, colour = "grey50", hjust = 0),
    legend.position  = "right",
    legend.key.size  = unit(0.45, "cm"),
    legend.text      = element_text(size = 8),
    panel.grid.minor = element_blank()
  )

print(fig_ss_var_4)
ggsave(here("output/fig_ss_instrument_variance_specialisation.png"),
       fig_ss_var_4, width = 13, height = 11, dpi = 300)
```


```{r}

ss_variance %>% saveRDS(here("data/temp/ss_variance.RDS"))

df_dist_all <- readRDS(here("data/temp/funding_distributions_plotting_data.RDS")) %>% 
   mutate(grade_numeric = case_when(grade == "A" ~ 1, 
                                   grade == "B" ~ 2,
                                   grade == "C" ~ 3,
                                   grade == "D" ~ 4,
                                   grade == "F" ~ 5)) %>% 
  mutate(axis_placement = 1 + 0.05*(source_year - 2005)) %>% 
    left_join(
    fips_codes %>%
      distinct(state_name, state),  # keep only needed columns, remove duplicates
    by = c("state" = "state_name")
  ) %>%
  rename(state_name = state.y) %>% 
  select(-state)

df_dist_all_sum <- df_dist_all %>% 
    group_by(state_name, grade) %>% 
     summarise(n = n(), 
               max_grade = min(grade_numeric), 
               min_grade = max(grade_numeric), .groups = "drop") %>% 
     group_by(state_name) %>% 
     filter(n == max(n)) %>% 
     slice(1) %>% 
  mutate(fiscal_mech = case_when(state_name == "TX" ~ "Robin Hood Recapture",
                                 state_name == "NM" ~ "Public School Finance Act & State Equalization Guarantee",
                                 state_name == "WY" ~ "Redistribution via School Foundation Program Account",
                                 #state_name == "MN" ~ "Multiple Equalization Programs",
                                 #state_name %in% c("WI") ~ "Property Tax Cap & Recapture", # "WY", "UT", "KS"
                                 state_name == "WI" ~ "Revenue Limits & Negative Equalization Formula",
                                 state_name == "UT" ~ "Property Tax Cap & Recapture",
                                 state_name == "KS" ~ "Local Option Budget Recapture",
                                 #state_name %in% c("KY", "ND", "OK", "SC") ~ "Possible State Aid Deductions",
                                 state_name == "OH" ~ "DeRolph v. State",
                                 state_name == "NE" ~ "Tax Equity and Educational Opportunities Support Act (TEEOSA)",
                                 state_name == "CA" ~ "Constitutional Cap of Property Tax Expenditure & Local Control Funding Formula (2013)",
                                 state_name == "AZ" ~ "State Equalization Assistance",
                                 state_name == "MT" ~ "95 Mill School Equalization Tax & School Equalization Account",
                                 state_name == "IN" ~ "2008 General Fund Reform",
                                 state_name == "MI" ~ "State-wide Tax Reliance",
                                 state_name == "NY" ~ "Wealth-Adjusted Foundation Aid",
                                 state_name == "WA" ~ "Local Effort Assistance for Low-Wealth Districts",
                                 TRUE ~ "None or not yet identified"),
         
         fiscal_mech_color = as.factor(ifelse(fiscal_mech == "None or not yet identified", "No", "Yes")))


saveRDS(df_dist_all_sum, here("data/temp/df_dist_all_sum.RDS"))

grade_fill_colours <- c("A" = "#1D9E75",
                        "B" = "#7FB539",
                        "C" = "#9A9A9A",  # grey
                        "D" = "#D85A30",
                        "F" = "#E24B4A")

# Compute the state ordering ONCE, outside the plot, so both layers agree
state_order <- ss_variance %>%
  arrange(mean_sd_cross_cz) %>%
  pull(state_name)

saveRDS(state_order, here("data/temp/state_order.RDS"))

# Build the per-year grade data, restricted to states that appear as bars
df_grades_over_time <- df_dist_all %>%
  inner_join(ss_variance %>% select(state_name, mean_sd_cross_cz),
             by = "state_name") %>%
  mutate(state_name = factor(state_name, levels = state_order))

# Figure out a sensible x-range for the grade strip.
# Place the strip in the rightmost portion of the plot so it doesn't
# overlap the bars. Adjust `strip_start` and `strip_width` to taste.
y_max        <- max(ss_variance$mean_sd_cross_cz, na.rm = TRUE)
strip_start  <- y_max * 1.02                        # just past longest bar
strip_width  <- y_max * 0.25                        # how wide the strip is
year_range   <- range(df_grades_over_time$source_year)

df_grades_over_time <- df_grades_over_time %>%
  mutate(axis_placement = strip_start +
           strip_width * (source_year - year_range[1]) /
                         (year_range[2] - year_range[1]))

saveRDS(df_grades_over_time, here("data/temp/df_grades_over_time.RDS"))

# Width of each tile on the y-axis (in data units).
# The spacing between consecutive years is strip_width / (n_years - 1),
# so set tile_width to match that for a continuous bar with no gaps.
n_years    <- length(unique(df_grades_over_time$source_year))
tile_width <- strip_width / (n_years - 1)
# fig_ss_var_1 <- 
ggplot(
  ss_variance %>%
    left_join(df_dist_all_sum, by = "state_name") %>% 
        mutate(state_name = factor(state_name, levels = state_order)),
  aes(x = state_name, y = mean_sd_cross_cz, fill = sig_state)
) +
  geom_col(alpha = 0.85) +
  coord_flip(clip = "off") +
  scale_fill_manual(
    values = c("TRUE" = "#1a7a3c", "FALSE" = "grey70"),
    labels = c("TRUE" = "Included in main IV results",
               "FALSE" = "Not significant"),
    name   = NULL
  ) +
  # Per-year grade strip as shaded tiles
  ggnewscale::new_scale_fill() +
  geom_tile(
    data = df_grades_over_time,
    aes(x = state_name, y = axis_placement, fill = grade),
    width = 0.8, height = tile_width,
    inherit.aes = FALSE
  ) +
  scale_fill_manual(values = grade_fill_colours, name = "Grade") +
  # scale_y_continuous(
  #   expand = expansion(mult = c(0.01, 0.05)),
  #   limits = c(0, strip_start + strip_width * 1.05)
  # ) +
      geom_text(aes(y = 0.01, hjust = 0, label = fiscal_mech, color = fiscal_mech_color), size = 2.5, fontface = "bold", family = "LMRoman10-Bold") +
    scale_color_manual(values = c(                                       # ← second color scale
    "Yes" = "black",
    "No"  = "white"
  ), name = "Broken Fiscal Channel") +
  labs(
   title    = "Cross-CZ Variation in Shift-Share Instrument by State",
    subtitle = str_wrap("Mean within-year SD of the shift-share instrument across CZs.
       Higher values indicate more heterogeneous industry exposure across
       CZs within the state. The color of the bar displays whether a statistically significant (green) non-zero relationship between wages and public education expenditure per pupil. The colored tiles represent the state's national ranking in terms of funding distribution effectiveness as defined by the Education Law Center, where each tile represents a year between 2006-2021. ELC determines state funding distribution rankings by the ratio of state-level funds directed to low vs. high-poverty districts. Finally, the text label represents the presence of a legislative or constitutional redistribution mechanism, and the policy's name, if applicable.", 90),
    x = NULL, y = "Mean cross-CZ SD of shift-share instrument",
    caption = "SD computed across CZs within each state-year, then averaged."
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 13),
    plot.subtitle    = element_text(size = 9, colour = "grey30"),
    legend.position  = "none",
    panel.grid.minor = element_blank(),
    plot.margin      = margin(5, 30, 5, 5)
  ) + common_theme

ggsave(here("output/cross_cz_w_redist_policies.png"), height = 12, width = 8)

print(fig_ss_var_1)


```


```{r, results = 'asis', cache = FALSE}

source(here("code/heterogeneity_analysis.R"))

```

```{r, results = 'asis', cache = FALSE}

source(here("code/heterogeneity_mechanism_test.R"))

```

```{r, results = 'asis', cache = FALSE}

source(here('code/leverage_analysis_cz.R'))

```

## Mechanism Figures

```{r fig_mechanism_map, echo=FALSE, warning=FALSE, message=FALSE, fig.width=12, fig.height=14, cache = FALSE}
# ============================================================
# FIGURE 1: US CHOROPLETH MAP OF MECHANISM TYPES
# ============================================================
library(maps)

mech_colors <- c(
  "Amplification" = "#d7191c",
  "Insulation"    = "#fdae61",
  "Equalization"  = "#2c7bb6"
)

mech_df <- read_csv(here("output/mechanism_typology.csv"), show_col_types = FALSE) %>%
  rename(abbr = state_name)

state_lookup <- tibble(
  abbr      = state.abb,
  full_name = tolower(state.name)
)

mech_df <- mech_df %>%
  left_join(state_lookup, by = "abbr")

us_poly <- map_data("state") %>%
  left_join(mech_df, by = c("region" = "full_name"))

state_centroids <- us_poly %>%
  filter(!is.na(abbr)) %>%
  group_by(abbr, mechanism, sig_state) %>%
  summarize(
    long = (min(long) + max(long)) / 2,
    lat  = (min(lat)  + max(lat))  / 2,
    .groups = "drop"
  )

fig_map <- ggplot() +
  geom_polygon(
    data = us_poly %>% filter(!is.na(mechanism)),
    aes(x = long, y = lat, group = group, fill = mechanism),
    color = "white", linewidth = 0.35
  ) +
  geom_polygon(
    data = us_poly %>% filter(is.na(mechanism)),
    aes(x = long, y = lat, group = group),
    fill = "grey88", color = "white", linewidth = 0.35
  ) +
  geom_polygon(
    data = us_poly %>% filter(sig_state == TRUE),
    aes(x = long, y = lat, group = group),
    fill = NA, color = "black", linewidth = 1.0
  ) +
  geom_text(
    data = state_centroids,
    aes(x = long, y = lat, label = abbr,
        fontface = if_else(sig_state, "bold", "plain")),
    size = 2.4, color = "white"
  ) +
  scale_fill_manual(
    values   = mech_colors,
    name     = "Funding Mechanism",
    na.value = "grey88"
  ) +
  coord_fixed(1.3) +
  labs(
    title    = "Fiscal Transmission Mechanisms for K-6 Education Spending",
    subtitle = "States classified by conditional correlation between CZ wage growth and residual spending growth. Bold borders: significant IV elasticity. Grey: excluded states.",
    caption  = "Amplification: rho >= 0.20 (n=9); Insulation: |rho| < 0.20 (n=15); Equalization: rho <= -0.20 (n=11). Data 2001-2021."
  ) +
  theme_void(base_size = 11) +
  theme(
    plot.title       = element_text(face = "bold", size = 13, hjust = 0.5,
                                    margin = margin(b = 4)),
    plot.subtitle    = element_text(size = 9, hjust = 0.5, color = "grey30",
                                    margin = margin(b = 6)),
    plot.caption     = element_text(size = 7.5, hjust = 0.5, color = "grey45",
                                    margin = margin(t = 8)),
    legend.position  = "bottom",
    legend.title     = element_text(face = "bold", size = 10),
    legend.text      = element_text(size = 9),
    legend.key.width = unit(1.2, "cm"),
    plot.margin      = margin(10, 10, 10, 10)
  )

print(fig_map)
ggsave(here("output/fig_mechanism_map.png"), fig_map, width = 12, height = 7, dpi = 300)
```

```{r fig_mechanism_fingerprint, echo=FALSE, warning=FALSE, message=FALSE, fig.width=13, fig.height=14, cache = FALSE}
# ============================================================
# FIGURE 2: MECHANISM FINGERPRINT PANEL
# Requires: revenue_summary, transfer_results, cyclical_summary
# (created by heterogeneity_mechanism_test.R, sourced above)
# ============================================================
library(patchwork)

mech_colors <- c(
  "Amplification" = "#d7191c",
  "Insulation"    = "#fdae61",
  "Equalization"  = "#2c7bb6"
)
mech_levels <- c("Amplification", "Insulation", "Equalization")

# Panel A: Local revenue share
p_a <- revenue_summary %>%
  mutate(
    mechanism = factor(mechanism, levels = mech_levels),
    pct_label = paste0(round(mean_local_share * 100, 1), "%")
  ) %>%
  ggplot(aes(x = mechanism, y = mean_local_share * 100, fill = mechanism)) +
  geom_col(width = 0.55, alpha = 0.9) +
  geom_text(aes(label = pct_label), vjust = -0.45, size = 3.6, fontface = "bold") +
  scale_fill_manual(values = mech_colors) +
  scale_y_continuous(
    limits = c(0, 55),
    labels = function(x) paste0(x, "%"),
    expand = expansion(mult = c(0, 0.08))
  ) +
  labs(
    title = "A. Local Revenue Reliance",
    x     = NULL,
    y     = "Local revenue as % of total spending"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position    = "none",
    plot.title         = element_text(face = "bold", size = 11),
    axis.text.x        = element_text(face = "bold", size = 10),
    panel.grid.major.x = element_blank()
  )

# Panel B: Transfer elasticity with 95% CIs
p_b <- transfer_results %>%
  mutate(mechanism = factor(mechanism, levels = mech_levels)) %>%
  ggplot(aes(x = mechanism, y = elasticity, color = mechanism)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey55", linewidth = 0.6) +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.18, linewidth = 1.1) +
  geom_point(size = 4.5) +
  geom_text(aes(label = paste0(sprintf("%+.3f", elasticity), sig)),
            vjust = -1.0, size = 3.4, fontface = "bold") +
  scale_color_manual(values = mech_colors) +
  labs(
    title = "B. Transfer Elasticity to Wage Growth",
    x     = NULL,
    y     = "beta (IG revenue growth | wage growth)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position    = "none",
    plot.title         = element_text(face = "bold", size = 11),
    axis.text.x        = element_text(face = "bold", size = 10),
    panel.grid.major.x = element_blank()
  )

# Panel C: Levels-based IG redistribution test
# Correlates mean log wages vs. mean log IG per pupil across CZs within each mechanism group.
# Unlike growth-rate tests, this captures the structural redistribution that equalizing
# formulas create: low-wage CZs persistently receive more IG (negative correlation).
ig_levels_by_mech <- df_ivs %>%
  group_by(unit, state) %>%
  summarize(
    mean_log_wage = mean(log_weighted_annual_avg_wkly_wage, na.rm = TRUE),
    mean_log_ig   = mean(log_real_Total_IG_Revenue_pp,      na.rm = TRUE),
    .groups = "drop"
  ) %>%
  left_join(
    read_csv(here("output/mechanism_typology.csv"), show_col_types = FALSE) %>%
      select(state = state_name, mechanism),
    by = "state"
  ) %>%
  filter(!is.na(mechanism), is.finite(mean_log_wage), is.finite(mean_log_ig)) %>%
  group_by(mechanism) %>%
  filter(n() >= 3) %>%
  summarize(
    cor_ig_wage = cor(mean_log_wage, mean_log_ig),
    .groups = "drop"
  ) %>%
  mutate(
    mechanism = factor(mechanism, levels = mech_levels),
    bar_vjust = if_else(cor_ig_wage >= 0, -0.45, 1.4),
    cor_label = sprintf("%+.2f", cor_ig_wage)
  )

p_c <- ggplot(ig_levels_by_mech, aes(x = mechanism, y = cor_ig_wage, fill = mechanism)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey55", linewidth = 0.6) +
  geom_col(width = 0.55, alpha = 0.9) +
  geom_text(aes(label = cor_label, vjust = bar_vjust), size = 3.6, fontface = "bold") +
  scale_fill_manual(values = mech_colors) +
  scale_y_continuous(expand = expansion(mult = c(0.2, 0.2))) +
  labs(
    title = "C. IG Levels vs. Wage Levels (Structural)",
    x     = NULL,
    y     = "Cross-CZ correlation: mean wages vs. mean IG per pupil"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position    = "none",
    plot.title         = element_text(face = "bold", size = 11),
    axis.text.x        = element_text(face = "bold", size = 10),
    panel.grid.major.x = element_blank()
  )

fig_fingerprint <- (p_a | p_b | p_c) +
  plot_annotation(
    title    = "Structural Differences Across Fiscal Transmission Mechanisms",
    subtitle = "Amplification (red): local-revenue-heavy, transfers unresponsive to wages. Insulation (orange): moderate local reliance, stable transfers. Equalization (blue): transfers actively offset local wage gains.",
    caption  = "Panel A: mean local revenue as share of total spending, by mechanism group. Panel B: FE regression of IG transfer growth on wage growth, year FE, SE clustered by CZ; error bars = 95% CI. Panel C: cross-CZ correlation between total wage growth and IG revenue growth within each mechanism group; negative values indicate lower-wage CZs receive more IG funding.",
    theme = theme(
      plot.title    = element_text(face = "bold", size = 13, hjust = 0),
      plot.subtitle = element_text(size = 9, color = "grey30", hjust = 0,
                                   margin = margin(t = 2, b = 8)),
      plot.caption  = element_text(size = 7.5, color = "grey45", hjust = 0,
                                   margin = margin(t = 10))
    )
  )

print(fig_fingerprint)
ggsave(here("output/fig_mechanism_fingerprint.png"), fig_fingerprint,
       width = 13, height = 5.5, dpi = 300)
```

```{r, cache = FALSE}
# Safety check: heterogeneity_analysis.R adds mechanism to mechanism_typology
# but NOT back to state_correlations. If the join in leverage_analysis_cz.R
# therefore failed to attach mechanism to cz_conditional_all, do it here.
# if (!"mechanism" %in% names(cz_conditional_all)) {
#   mech_lookup <- read_csv(here("output/mechanism_typology.csv"), show_col_types = FALSE) %>%
#     rename(state = state_name) %>%
#     select(state, mechanism) %>%
#     mutate(mechanism = factor(mechanism, levels = c("Amplification", "Insulation", "Equalization")))
#   cz_conditional_all <- cz_conditional_all %>%
#     left_join(mech_lookup, by = "state")
# }

print_lev_sum <- function(state){
    show_state_drivers(state, n = 60) -> state_driver_temp
    #print(state_driver_temp)
    temp_res <- state_driver_temp$cz_id %>%
    compare_czs(., state_name = get_state(state))
    
    # map_wage_bust_episodes(wage_pctile = 0.25, spend_pctile = 0.25, include_states = state, 
    #                        sig_states = ifelse(state %in% sig_states, state, NULL), include_cities = TRUE, city_pop = 100000)
    map_wage_bust_episodes(wage_method = "raw",  include_states = state, 
                           sig_states = ifelse(state %in% sig_states, state, NULL), include_cities = TRUE, city_pop = 100000)
    map_wage_bust_episodes(wage_method = "iv_pooled",  include_states = state, 
                               sig_states = ifelse(state %in% sig_states, state, NULL), include_cities = TRUE, city_pop = 100000)
    map_wage_bust_episodes(wage_method = "iv_state",  include_states = state, 
                               sig_states = ifelse(state %in% sig_states, state, NULL), include_cities = TRUE, city_pop = 100000)
}

# Allows to iterate over a list across separate chunks - pops a list element before the next one. 
pop_state <- function(state_list, temp_state_code){
  if(length(state_list) > 1){
      remaining_state_list <- setdiff(state_list, temp_state_code)
  }else{
    return(NULL)
  }
  
  return(remaining_state_list)
}

```

```{r, fig.height=14, fig.width = 10}

source(here("code/heterogeneity_industrial_share_analysis.R"))

```

## Mapping All States' Booms & Busts

```{r}

map_wage_bust_episodes(wage_pctile = 0.5, spend_pctile = 0.5, include_states = reg_states, 
                       sig_states = sig_states, include_cities = TRUE, city_pop = 100000, wage_method = "iv_pooled")

```


Note that any district and school information is sourced from the National Center for Education Statistics' "Search for Public School Districts" database: https://nces.ed.gov/ccd/districtsearch/.

## Significant States

<!-- ::: {.panel-tabset} -->

### Colorado

Colorado is a remarkable state precisely because education spending almost perfectly traces wages.

Colorado is one of ten states identified has having a positive causal wage-education spending relationship. Examining its commuting zones in detail confirms this pattern, wherein commuting zones whose wages follow a stable pattern exhibit stable spending, whereas those whose wages fluctuate considerably from a smooth trend show more volatile spending patterns. Though there is evidence that the spending relationship is asymmetric. In other words, local sources increase education spending in "good" times and are only partially offset by intergovernmental spending in "bad" times. However, this is far from a consistently responsive mechanism, as can be seen in commuting zone 663 where IG expenditure increases despite significant local booms to spending. Altogether, Colorado's pattern is consistent with an amplification mechanism, wherein local wage booms and busts contribute to spending increases and cuts, though the cuts are offset in certain commuting zones by intergovernmental expenditure responses. The concern remains that this intergovernmental response seems to be inconsistent across commuting zones. For example, in CZ 663 (Jackson County, CO (1 school - 1 district) & Carbon County, WY (11 schools - 2 districts)), IG expenditure increases despite significant local booms to spending, whereas in CZ 633 (Gunnison (1 district - 7 schools) and Hinsdale (1 school - 1 district) Counties), IG expenditure steps in to correct for local resource losses post-2012; and CZ 693 (Baca County - 11 schools across 5 public school districts) sees both IG spending move in tandem with local resource changes, meaning that the intergovernmental response is not reliable during local fiscal downturns.  

```{r, cache = FALSE}

curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r, cache = FALSE}

print_lev_sum(curr_state)

temp_trash <- show_state_drivers("08", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("08"), highlight = c(663, 5, 633, 691, 693, 697))

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, "08", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 1)  

show_fiscal_decomposition(c(663, 5, 693, 633, 691), "08", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 1)  

show_cz_industry_profile(c(414, 663, 691, 183, 677, 633), state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

```

### Florida

The spending jump Franklin County, Florida, likely comes from a significant capital expenditure allocated in the 2007-2008 budget: https://www.leg.state.fl.us/data/committees/Senate/EA/EdFundingsummary_2007.pdf.

Florida represents a state in which spending co-moves with local economic conditions (GDP and wages). Essentially, spending increases prior to 2008, decreased after, and has slowly been recovering since. However, certain commuting zones still spend more per student than others, largely as a factor of uneven local revenues feeding into school systems in boom periods without equalizing intergovernmental revenues. Notably, intergovernmental revenue has become more correcting post-2008. Debate within Florida has been waged regarding the burden of public education funding (taxation at the local or state level), with state legislators preferring to push the burden to local communities, aligning with the data visualized in the below figure, where spending growth lags due to slowly recovering local economic conditions post-2008 and intergovernmental revenue decreasing across nearly all commuting zones. Essentially, in Florida (mainly post-2008), intergovernmental revenues have not served an equalizing purpose, but has rather dwindled in favor of a reliance on local funding mechanisms, leaving budgets susceptible to fluctuations in local conditions and the economic geography within the state.

```{r}

curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}


```

```{r, fig.width = 16, fig.height=14}

print_lev_sum("12")

florida_czs <-  c(62, 670, 410, 307, 306, 522, 334, 325, 374)

temp_trash <- show_state_drivers("12", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("12"), highlight = florida_czs)

all_florida_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_florida_czs, "12", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 1)  
show_cz_industry_profile(florida_czs, state_code = "12")

```

### South Dakota

Dips in total commuting-zone level wages in South Dakota are linked to wage losses in the construction industry (CZ 556 and 680); whereas other regions seeing similar construction-wage based dips are outweighed by the stability of wages in other sectors. Interesting cases are Pierre, SD (CZ6); Watertown, SD, whose spending increases significantly post-2008 (CZ301); Aberdeen, SD (324); Brookings, SD (467); CZ646 Boom in 2016/17; CZ 260 - major local resource losses offset by intergovernmental revenues. 

There are four commuting zones that receive above-mean levels of intergovernmental transfers (260 (Dewey and Ziebach Counties), 556 (Bennett and Shannon Counties), 459 (Mellette & Todd Counties), 474 (Sioux, Cambell, Corson, Walworth Counties)). These are school districts that lie within Indian/Native American Reservations within the rural corridor of the state.

The figure demonstrating the fiscal decomposition of spending changes displays commuting zones, ordered from highest to lowest spending per pupil where the black outline demonstrates the zones in the top 25th percentile of wage growth variance. Of the top five commuting zones in terms of wage variance, three include those counties within Reservations. The remaining two rely largely on wages from Construction and Agricultural industries. 

```{r}

curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}


```

```{r}

print_lev_sum(curr_state)

if(!exists("south_dakota_czs")){south_dakota_czs <- NULL}

temp_trash <- show_state_drivers("46", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("46"), highlight = south_dakota_czs)

south_dakota_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:round(0.25*nrow(temp_trash))) %>% pull(cz_id)
all_sd_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_sd_czs, "46", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = south_dakota_czs)  
show_cz_industry_profile(south_dakota_czs, state_code = "46")
map_state(state_name = "46", include_cities = TRUE, highlight_czs = south_dakota_czs)

```

### Kentucky

Additionally, 72 (Frankfort, KY, Lexington-Lafayetta) has extremely low IG spending but high wages. As is common across the United States, healthcare services dominate the commuting zones in the 25th percentile of wage growth variance. However, in these commuting zones, either natural resources (including mining) and hospitality immediately follow making up nearly 50% of employment per commuting zone. These commuting zones exhibit clear wage stagnation following the 2008 financial crisis and post-2010 decline in coal production, likely explaining these wage patterns. As can be seen in the map the commuting zones with greatest wage growth variation are clustered in the eastern coal-producing region of Kentucky, northern urban areas near Louisville, Kentucky and Cincinnati, Ohio; and on the border with Union City and Dyersburg, Tennessee, where hospitality and public administration industries are major employers.

```{r}
curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("kentucky_czs")){kentucky_czs <- NULL}

temp_trash <- show_state_drivers("21", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("21"), highlight = kentucky_czs)


kentucky_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:round(0.25*nrow(temp_trash))) %>% pull(cz_id)
all_ky_czs <- temp_trash %>% pull(cz_id)

# c(117, 319, 463, 450, 302))
show_fiscal_decomposition(all_ky_czs, "21", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = kentucky_czs)  
show_cz_industry_profile(kentucky_czs, state_code = "21")
map_state(state_name = "21", include_cities = TRUE, highlight_czs = 89)

```

### North Dakota

North Dakota demonstrates the relevance of stable wage growth. The commuting zones with greatest wage growth variation see a significant boost in local wages around 2010. However, wages in these areas stagnate within 2-3 years, achieving a near-zero rate of growth, whereas the remaining commuting zones largely have lower wage levels, but stable growth. The map indicates that the commuting zones seeing wage boosts with subsequent stagnation are concentrated in the western half of the state near the border with Montana. Examining the industrial composition of these two types of North Dakotan commuting zones, the western zones' top industries are healthcare, wholesale trade, and public administration; whereas in the eastern zones, the top industries behind healthcare are public administration followed by wholesale trade. Essentially, wholesale trade is a strong employer in the west due to the presence of a large agricultural sector. The shape of wages in the western commuting zones aligns with that of real value added from the agriculture, forestry, fishing, and hunting as a percentage of GDP. 

The fiscal decomposition of education spending per pupil across North Dakotan commuting zones indicates a pattern in which the zones experiencing wage booms and subsequent stagnation, have seen spending booms with non-enduring effects. On the other hand, other zones have seen achieved a stable higher level of expenditure per pupil.

```{r}
curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("nd_czs")){nd_czs <- NULL}

temp_trash <- show_state_drivers("38", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("38"), highlight = nd_czs)

nd_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_nd_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_nd_czs, "38", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = nd_czs)  
show_cz_industry_profile(nd_czs, state_code = "38")
show_cz_industry_profile(setdiff(all_nd_czs, nd_czs), state_code = "38")
map_state(state_name = "38", include_cities = TRUE, highlight_czs = nd_czs)

```

### Louisiana

All commuting zones with greatest wage growth variance have a high proportion of employment in the mining sector, aligning with the presence, whereas this is only a small proportion of employment in other commuting zones. The map demonstrates the location of commuting zones with considerable wage variance along the "Industrial Corridor" of Louisiana which follows the south end of the Mississippi River approaching the Gulf Coast. This corridor houses several petrochemical refineries which serve as a source of local employment while also transmitting variability in natural resource wages to commuting zones in this region. 

```{r}
curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("la_czs")){la_czs <- NULL}

temp_trash <- show_state_drivers("22", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("22"), highlight = la_czs)

la_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_la_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_la_czs, "22", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = la_czs)  
show_cz_industry_profile(la_czs, state_code = "22")
show_cz_industry_profile(setdiff(all_la_czs, la_czs), state_code = "22")
map_state(state_name = "22", include_cities = TRUE, highlight_czs = la_czs)


```

### Pennsylvania

Pennsylvania's statistically significant relationship stems largely from a discernible connection between spending and state-level economic conditions, rather than clear delineations between diverging wage trajectories. As evidenced in the fiscal decomposition figure below, educational spending growth is largely consistent across commuting zones in the state, however the relative share of intergovernmental versus local revenue shares varies remarkably. In CZ 215 (Chautauqua County, Clarion County, Crawford County, Erie County, Forest County, Venango County, Warren County), CZ 358 (Bradford County, Clinton County, Lycoming County, Sullivan County), CZ 524 (Cameron County, Elk County, Mc Kean County, Potter County), spending growth has kept up with state-wide rates due largely to intergovernmental transfers. 

```{r}
curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("pa_czs")){pa_czs <- NULL}
temp_trash <- show_state_drivers("42", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("42"), highlight = pa_czs)

pa_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_pa_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_pa_czs, "42", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = pa_czs)  
show_cz_industry_profile(pa_czs, state_code = "42")
show_cz_industry_profile(setdiff(all_pa_czs, pa_czs), state_code = "42")
map_state(state_name = "42", include_cities = TRUE, highlight_czs = pa_czs)


```

### Oregon

The commuting zones exhibiting greatest wage growth variation in Oregon are those whose employment relies heavily on the Agriculture, Forestry, Fishing, and Hunting industry, as compared to the remaining commuting zones. Notably, education spending in these commuting zones correlate strongly with wage growth rates as indicated by the red and green shading in the fiscal decomposition plots. Furthermore, examining CZ 662 (Lincoln and Tillamook Counties) with a notably strong co-movement between wage growth and education spending, these areas rely greatly on income from Accommodation and Food Services given their proximity to tourist destinations along the Pacific Coast including state forests. 

```{r}
curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("or_czs")){or_czs <- NULL}
temp_trash <- show_state_drivers("41", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("41"), highlight = or_czs)

or_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_or_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_or_czs, "41", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = or_czs)  
show_cz_industry_profile(or_czs, state_code = "41")
show_cz_industry_profile(c(174, 662), state_code = "41")
map_state(state_name = "41", include_cities = TRUE, highlight_czs = or_czs)

```

### Oklahoma

The commuting zones exhibiting greatest wage growth variation overlap with Native American reservations and have an above-state-average share of employment in mining and public administration.  

Other notable commuting zones: CZ 370 (Blaine, Dewey, Ellis, Harper, Woodward Counties) lies mainly within the Cheyenne-Arapaho Tribal Jurisdiction of Oklahoma. This commuting zone has received a decreasing amount of intergovernmental revenue since 2008 though overall spending has increased as a result of rising wages in the area. 

```{r}
curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("ok_czs")){ok_czs <- NULL}
temp_trash <- show_state_drivers("40", n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state("40"), highlight = ok_czs)

ok_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_ok_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_ok_czs, "40", decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = ok_czs)  
show_cz_industry_profile(ok_czs, state_code = "40")
show_cz_industry_profile(setdiff(all_ok_czs, ok_czs), state_code = "40")
map_state(state_name = "40", include_cities = TRUE, highlight_czs = ok_czs)

```

### Indiana

Indiana's public education system underwent two major shifts to its funding landscape during the time period analyzed in this study. First, the Indiana General Assembly passed Public Law 148 n 2008 which eliminated property taxes as a revenue source for the state General Fund, the main vehicle funding school operating expenses. This legal change is demonstrated clearly in the figures below, wherein the intergovernmental transfers across all commuting zones exhibits a distinct mean shift [Michael 2009; Egalite & Catt 2025]. . Second, Indiana has championed "school choice" reform programs wherein the state has, among other reforms, subsidized access to private education through publicly funded voucher programs and established several charter schools which aim to provide independently operated public schools. Critics of "school choice" reform programs cite the potential for charter schools or voucher programs to crowd out traditional public school funding (Wang et al; Shaffer; Bifulco), while proponents cite the ability to adapt to heterogeneous student needs, the benefit of a more direct student funding channel (families and students themselves decide where to spend their allotted funding dollars); and scant empirical evidence of harms to students remaining in traditional public schools (Bitler; Egalite; Wang; Egalite & Catt).

However, evident in the below figures, public school funding has decreased significantly across nearly all commuting zones within Indiana following 2008. Though intergovernmental revenues have largely balanced the losses in local funding sources, public education expenditure remains at a lower level in both real and nominal terms toward the end of the time series analyzed in this study. 

```{r}

curr_state <- sig_states_temp[1]

if(!is.null(sig_states_temp)){
  sig_states_temp <- pop_state(sig_states_temp, curr_state)
  print(paste0("Remaining list: ", paste(sig_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)


```

<!-- ::: -->

## All Other States

For the states where no statistically significant causal effect is detected, we proceed with an analysis of the potential source of the negligible effect. This diagnosis is important for two reasons. First, a near-zero elasticity can indicate the presence of strong intergovernmental equalization tools to alleviate the spill-over consequences of uneven wage distributions within a state. Second, a near-zero elasticity can indicate a fundamental decoupling between commuting-zone level economic conditions. Both of these sources of a negligible elasticity would provide potential support for the wider implementation of the relevant state or local public education funding mechanisms as they effectively insulate commuting zones from a wage-based channel for public education inequity. However, a near-zero elasticity could additionally indicate either a poorly estimated functional form unfit for the research question or insufficient within-state variation to detect meaningful differences between higher- versus lower-wage commuting zones. In either of these latter cases, it will be important to highlight any statistical-, data-, or estimation-related obstacles to identification to ensure that undue conclusions about the quality of local public education funding systems are not drawn from this effect. In other words, distinguishing between states in which a non-zero elasticity has meaningful policy relevance or is simply an artifact of a poor econometric specification or insufficient data is imperative to ensure the integrity of conclusions drawn. 

<!-- ::: {.panel-tabset} -->

### Alabama

Panel C demonstrates that the relationship between wage growth and spending growth is near flat while Panel A indicates that there is little variation in wage growth across Alabama commuting zones. Panel B demonstrates that the intergovernmental expenditure shows little evidence of responding to wage growth in an insulating or equalizing way. Combined, this indicates that the inability to detect an effect likely stems from an identification challenge rather than evidence of an effective equalization system. 

```{r}

rm(temp_czs)
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

#print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)


diagnose_no_effect(curr_state)


```

### Arizona

Arizona has considerable variation in commuting zone level wages, aligned with the distribution of the country as a whole. However, wage growth appears to have a negligible relationship with both intergovernmental funding and spending growth indicating that commuting zones are largely decoupled from commuting-zone specific wage dynamics. Nonetheless, all six commuting zones remain strongly linked to higher-level economic indicators that appear to affect local and intergovernmental funding evenly. 

The commuting zone containing Graham and Greenlee counties exhibits a wage growth rate ~50% higher than the rest of the commuting zones in the state, which can likely be attributed to the reliance of Greenlee county on employment by the Morenci Mine, the largest copper mining operation in North America. In 2024, 90% of the county's GDP and 70% of the county's total jobs relied on mining, according to a report produced by Better City and sponsored by the holding company of the Morenci Mine (Freeport-McMoran)  [source](https://www.freeportinmycommunity.com/uploads/2024.05_Greenlee_RAP_Full.pdf). We therefore exclude this commuting zone when evaluating the relationships below. 

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state, exclude = 207)


```

### Arkansas

Arkansas exhibits wage variation similar to that of the full-country sample; wage growth and spending are largely decoupled; and intergovernmental revenue neither equalizes nor amplifies any wage-spending link. 

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### California

California exhibits slightly low wage variation, especially when excluding the two outlier commuting zones surrounding San Francisco (including some of the highest-income counties in the United States like Napa, Marin, and Monterey Counties). However, excluding these outliers, there is considerable evidence of a strong equalizing mechanism in the state whereby wage growth and spending growth exhibit a positive relationship, counter-balanced by a negative relationships between wage growth and intergovernmental expenditure. 

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state, exclude = c(218, 294))

```

### Georgia

Georgia seems to exhibit strong equalization as well wherein the relationship between wage and spending growth is positive but wage and intergovernmental revenue is negative. Most of the growth in spending across commuting zones comes from local sources. 

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Idaho

In 2006, Idaho state legislators passed House Bill 1 which cut property taxes for public schools by nearly $USD 260 million [CITE RICHERT KEVIN - House Bill is unable to load]. This policy change is evident in the figures below, where state funding (mainly sourced from state sales and income taxes) replaced this funding loss for a brief period. However, the impact of the Great Recession on state revenues depressed the intergovernmental resources flowing to local school districts. Idaho presents an instance in which the attempted equalization via a shift from property taxes to increased intergovernmental funding as a source of public school revenues ultimately backfired due to the uniform impact of the Great Recession on the state's funding. This demonstrates that the centralization of public revenue collection and dispensation cannot insulate communities from the effects of broader economic downturns.

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Illinois

Wages have a near-zero direct relationship with spending and, excluding an outlier commuting zone that crosses the state boundary with Iowa and Montana, a negative relationship with intergovernmental spending growth indicating a strong channel of equalization between Illinois commuting zones. 

```{r}

curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state, exclude = c(192))

```

### Iowa

PN: Discard

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Kansas

Kansas represents a strong example of fiscal equalization wherein the relationship between wages and spending growth is negative largely due to a negative relationship between intergovernmental revenue and local wages. CZ 687 (Ness County, Kansas) is exemplary of this, where despite a strong reliance on wholesale trade form agriculture and beef and cattle ranching which experienced significant declines in demand following the 2008 crisis, education spending per pupil remained steady during this period. 

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Michigan

Amplified bust near Detroit, Michigan! Intergovernmental revenue is equalizing - wage and spending growth have a positive relationship but wage and IG growth have a strong negative relationship. Nonetheless, the bust following the bankruptcy declarations of GM and Chrysler in 2009 took a toll on local finances in the Detroit Michigan area, as evidenced by the red region in the map below. 

Counties experiencing fluctuations in CZ 395 (Charlevoix; Cheboygan; Emmet counties) and CZ 172 (counties surrounding Big Rapids, Cadillac, MI) are all heavily reliant on the accommodation and food services industries which experienced significant tourism decline in the recovery from the 2008 recession.

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Minnesota

Minnesota demonstrates a state in which wage changes likely filter through to public education budgets via intergovernmental revenue (state income taxes). As is demonstrated in the figures below, commuting zones with higher wage growth during the period studied are associated with higher levels of intergovernmental expenditure. As is demonstrated in the map below, there are several both boom and bust episodes that are amplified across the state.

In the figure below, CZ 437 (counties surrounding Marshall, MN) experienced a wage boom from 2018-2020 which boosted local public education expenditure. This wage boom is likely linked to the sale of Schwan's Company, a major employer in Marshall, MN to South Korea's CJ CheilJedang. [Source (https://royalexaminer.com/schwans-home-delivery-shuts-down-after-72-years-of-service/)]. Additionally, CZ 305 (counties surrounding Brainerd, MN) saw a wage boom toward the end of the time period analyzed in this study, likely attributable to a building boom around the lakes in the area and outdoor tourism during and after COVID-19 lockdowns.

CZ 144 containing the counties surrounding Rochester, MN and Red Wing, MN are likely attributable to a significant contraction and subsequent closing of the Rochester IBM campus. Sources:  https://www.mprnews.org/story/2012/02/28/ibm-rochester-layoffs; https://www.mprnews.org/story/2013/03/07/stunned-silence-as-ibm-breaks-news-to-rochester-employees. 

CZs 683 (Koochiching County) and 496 (Lake of the Woods an Roseau counties) suffered wage losses where intergovernmental funding insulated significant fiscal losses. Koochiching County saw significant layoffs at around 2013 due to a declining timber and mill industry, mainly operated by Boise Inc. (source: https://www.mprnews.org/story/2013/05/02/boise-mill-layoff-of-265-workers-devastating; https://www.mprnews.org/story/2013/09/26/as-boise-layoffs-hit-trying-to-make-new-jobs-in-international-falls; https://www.twincities.com/2013/05/07/community-reacts-responds-to-boise-inc-loss-of-jobs/). CZ 496's wage losses are likely attributable to a decline in local manufacturing. Roseau's main employer is a recreational sports equipment employer, Polaris, which struggled to recover after the 2008 Recession (source: https://www.mprnews.org/story/2009/01/21/polaris-more-than-100-hourly-workers-laid-off). In both cases, as demonstrated in the figures below, intergovernmental revenue largely balanced the losses in local revenue to avoid significant impact to local public education budgets. 

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Mississippi

Mississippi is one of the poorest states in the entire country. The state faced major employment losses In the figures below, it demonstrates a strong equalizing relationship between commuting zones where the relationship between wage growth and spending growth is positive but intergovernmental revenue favors areas with low wage growth. 

We se four cases of wage busts that were equalized by intergovernmental funding structures. 

CZ 416 (Coahoma; Lafayette; Panola; Quitman; Tallahatchie; Yalobusha counties surrounding Clarksdale and Oxford, MS) saw a wage loss episode from 2016-2018. The healthcare sector, Coahoma County's largest employer, was destabilized when Community Health Systems transferred the lease of the county's only hospital to Curae Health in 2017, which filed for Chapter 11 bankruptcy in 2018 and issued WARN notices to nearly 500 workers.

CZ 156 (Carroll; Grenada; Leflore; Montgomery counties surrounding Greenwood and Grenada, MS) saw wage losses from 2012-20214. Viking Range, Greenwood's dominant manufacturer, was acquired by The Middleby Corporation in December 2012 and followed the acquisition with rounds of layoffs; simultaneously, the Delta's farm-raised catfish industry (Source: https://mississippiencyclopedia.org/entries/catfish-farming/), which employed more workers in Leflore County than Viking did, was in steep decline due to competition from Vietnamese imports (Source: https://mississippitoday.org/2021/06/21/mississippi-is-losing-the-catfish-wars/), with hundreds of ponds converted or abandoned across the region by this period.

CZ 589 (Attala and Holmes Counties) saw a wage bust episode between 2002–2004. This window coincides with the wave of rural Southern manufacturing closures that followed China's WTO accession in 2001, which eliminated many of the light assembly and other manufacturing jobs these counties depended on; Holmes County was further exposed through the early contraction of the Delta catfish processing industry, which had provided a modest employment base through the 1990s.

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state, exclude = c(1, 389))

```

### Missouri

Missouri exhibits a positive relationship between wages and spending growth. However, intergovernmental revenues do

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Montana

PN: Discard

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Nebraska

PN: Discard; potentially look at CZ with above-average IG spending and increasing IG spending (two pink lines)

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### New Mexico

Prelim notes: pink commuting zones - low wage, high spend

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### New York

PN: Orange and gold CZ with high wages, high spending per pupil, and low IG - PERFECT EXAMPLE OF EQUALIZATION!

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Ohio

Prelim Notes: Variability in IG Spending; Orange CZ with higher spending; blue and green with collapsing spending.

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### South Carolina

PN: Deemed potential outlier...although I now see that the CZ with lowest GDP per capita gets the most IG revenue.

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Texas

PN: none...

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Utah

PN: teal commuting zone has high spending and increasing IG at the same time as GDP per capita nd wages are stagnating/decreasing! Interesting case as well

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Washington

PN: Red CZ has highest GDP, enrollment, and wages, but lowest IG - also potentially a good example of equalization at play.

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### West Virginia

PN: Purple line is potential outlier

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Wisconsin

PN: Discard

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```

### Wyoming

```{r}
curr_state <- reg_states_temp[1]

if(!is.null(reg_states_temp)){
  reg_states_temp <- pop_state(reg_states_temp, curr_state)
  # print(paste0("Remaining list: ", paste(reg_states_temp, collapse = ",")))
}

```

```{r}

print_lev_sum(curr_state)

if(!exists("temp_czs")){temp_czs <- NULL}
temp_trash <- show_state_drivers(curr_state, n = 60) %>% .$cz_id %>% 
compare_czs(., state_name = get_state(curr_state), highlight = temp_czs)

temp_czs <- temp_trash %>% arrange(desc(sd_gr_wage)) %>% slice(1:(round(0.25*nrow(temp_trash)) + 1)) %>% pull(cz_id) 
all_temp_czs <- temp_trash %>% pull(cz_id)

show_fiscal_decomposition(all_temp_czs, curr_state, decomp_bars = TRUE, add_wages = TRUE, wage_lag = 2, highlight_czs = temp_czs)  
show_cz_industry_profile(temp_czs, state_code = curr_state)
show_cz_industry_profile(setdiff(all_temp_czs, temp_czs), state_code = curr_state)
map_state(state_name = curr_state, include_cities = TRUE, highlight_czs = temp_czs)

diagnose_no_effect(curr_state)

```
<!-- ::: -->
