---
title: "Uneven Wage Growth and Public Goods"
subtitle: "The Case of US Public Education"
author:
  - name: Ebba Mark
    affiliation: University of Oxford
abstract: |
  Explores the effect of uneven wage growth on public education expenditure through its effect on local property values in the US. The work aims to illuminate the elasticity of public education expenditure to changes in local livelihoods and economic conditions with implications for the delivery of public services in a political economy defined by greater income and wealth inequality. Employing various methods to account for observed and unobserved heterogeneity in state-level tax regimes, income, and economic growth rates we find that this heterogeneity almost entirely determines the elasticity of public education expenditure to uneven wage growth....[These results provide insight into region- and state-specific adjustments that can be made to ensure that uneven economic development and structural transformation does not exacerbate existing in equalities in public service delivery in the United States?? (We shall see...)]
format:
  pdf:
    documentclass: article
    toc: true
    number-sections: true
    fontsize: 10pt
    pdf-engine: xelatex
    include-in-header: header.tex
editor: 
  markdown: 
    wrap: 72
bibliography: pub_fin_md.bib
---

```{r setup-appendix-writer, include=FALSE}

library(here)

appendix_file <- "appendix_generated.qmd"

# Initialize the file with appendix header only (overwrite if exists)
writeLines(c(
  "# Appendix ",
  "",
  "```{=latex}",
  "\\appendix",
  "```",
  ""
), here(paste0("code/", appendix_file)))

# Define function to append a rendered chunk to it
append_to_appendix <- function(title, description) {
  chunk_label <- knitr::opts_current$get("label")
  lines <- c(
    sprintf("%s", title),
    "",
    description,
    "",
    sprintf("```{r, ref.label='%s', echo = FALSE, eval=TRUE, results = 'asis'}", chunk_label),
    "```",
    ""
  )
  write(lines, file = here(paste0("code/", appendix_file)), append = TRUE)
}

```

# Working Notes {#sec-procedural-notes}

\textcolor{red}{The following are notes to keep in mind while the project is still underway.}

The following document summarises the progress made thus far on Chapter
1: Local Fiscal Risks of Decarbonisation of my DPhil. The work aims to
pursue a better understanding of how industrial transformation impacts
local well-being. From an original interest in looking at all aspects of
local public finance, the project has narrowed to focus on expenditure
on public education and its connection to industrial prosperity and
transformation.

Current strategy/research plan:\
1. Outcome: Educational Expenditure\
2. Treatment (endogenous): Wages, Economic Growth, Property Values,
Property Taxes\
3. Instrument: Industry Shares of Employment in high vs. low wage growth
industries/sectors

Items to be adjusted:
\begin{itemize}
  \item AR(1) bias correction to account for fixed effect interaction. 
  \item Include additional control variables (migration, poverty, race, rurality, home ownership rates, private school enrollment)
  \item Spatial autocorrelation term
  \item Run gets on central model with CFESIS (preliminary testing indicates coefficient breaks in 2006 and 2013 which neatly bookends the financial crisis/housing market collapse - interesting?)
  \item Separate shift-share on more granular industrial categories (ie. high and low-wage areas correcting for local CPI)
\end{itemize}

Note: Any warnings about "missing observations" or "NA being removed"
relates to the lags incorporated, except in the Bartik estimations.

# Introduction

Though considerable debate has recently been stoked in the United States
regarding the inherent value of public goods provision, majority of
developed economies around the world agree on the importance of securing
a reasonable standard of living for all.

## Uneven Public Goods Expenditure

## Uneven Wage Growth

## Economic Health & The Elasticity of Public Education Expenditure

## Risks for Increasing Inequality and Structural Transformation

**In many parts of the United States, local well-being is heavily
cointegrated with industrial prosperity and stability, not least through
the channel of public expenditure.**

\begin{itemize} 
  \item Proof of cointegrating relationship 
  \item Economic diversity and public well-being 
\end{itemize}

A recent study by Resources for the Future found that US revenues from
fossil fuels generated about \$ 138 billion annually for US localities,
states, tribes, and the federal government @raimi2022 . This amount is
forecast to decline by 2050 even in a business-as-usual scenario
assuming no changes in climate policy stringency. The study also finds
that Wyoming, North Dakota, Alaska, and New Mexico are the states most
dependent on fossil fuel revenues with at least 14% of state and local
revenues generated from the fossil fuel industry (Wyoming's dependence
is above 50%). The work makes a demonstrative statement about the link
between this revenue stream and essential services like schools, public
health, and infrastructure, but stops short of an empirical analysis
into the impact of fossil fuel decline on revenues and associated
expenditure.

**One channel through which such challenges will present themselves is
the potential hollowing out of local public revenues.** In many parts of
the United States, local well-being is heavily cointegrated with
industrial prosperity and stability, not least through the channel of
public expenditure. US revenues from fossil fuels generated about \$138
billion annually for US localities, states, tribes, and the federal
government @raimi2022 . This amount is forecast to decline by 2050 even
in a business-as-usual scenario (assuming no changes in climate policy
stringency). Wyoming, North Dakota, Alaska, and New Mexico are the
states most dependent on fossil fuel revenues with at least 14% of state
and local revenues generated from the fossil fuel industry (Wyoming's
dependence is above 50%). The work makes a demonstrative statement about
the link between this revenue stream and essential services like
schools, public health, and infrastructure, but stops short of an
empirical analysis into the impact of fossil fuel decline on revenues
and associated expenditure, even at the state level.

**Furthermore, (adding insult to injury) community well-being and public
expenditure in the US is already characterised by a high degree of
spatial heterogeneity.** Not only does the US consistently rank among
the top 5 most unequal OECD countries [^1].

[^1]: The US consistently ranks among the top 5 most unequal countries
    in the OECD alongside Turkey, Mexico, Chile, and Costa Rica across
    all relevant indicators reported by the OECD: Gini coefficient,
    three interdecile income ratios (P50/P10; P90/P10; P90/P50), Palma
    ratio, S80/S20 quintile share.}, this inequality is further
    reflected in uneven and unequal quality of infrastructure,
    education, healthcare leading to real consequences for particular
    people and places @chetty2016, @logan2012, @semuels2016,
    @Avancena2021, @flavin2009 .

**One public service that has particularly important ties to ensuring
generational resilience to economic decline is education.** Public
schools around the US are responsible for educating over 80% of
school-age children. In 2019, governments around the US (including the
federal government) spent a total of \$870 billion on public education,
roughly \$17,013 per pupil @nationalcenterforeducationstatistics2023 .
However, the quality of services delivered varies widely across the
country. [^2]

[^2]: In 2016, for example, the Connecticut State Department of
    Education reported that the town of Greenwich, one of the
    highest-income towns in the country, spent \\\$8,000 more per pupil
    than Bridgeport (\\\$21.9k versus \\\$13.7k per pupil), despite both
    towns being part of the same county, located less than 40 kilometers
    apart @semuels2016 .

The quality of public education, especially at an early age, can have
long-lasting consequences for personal and economic well-being over an
individual's lifetime as well as generations following them @alfonso2020
. Therefore, ensuring that local or regional economic decline does not
disrupt or worsen the quality of education delivered to
resource-dependent communities will be of paramount importance in a
net-zero transition. [^3]

[^3]: Perhaps the most prominent and often-cited relationship between
    education and extractive industries is through the lens of the
    'resource curse.' The validity and empirical existence of a
    'resource curse' has been tested since its conception with disparate
    results [@ahlerup2020, @badeeb2017, @blanco2012, @borge2015,
    @brunnschweiler2008, @cockx2014, @cockx2016, @deacon2011,
    @dialga2022, @douglas2017, @haber, @menaldo2016, @ross2018,
    @ross2015, @sincovich2018, @wiens2014]. The literature is divided
    into two strands focusing on either political (the relationship
    between resource wealth and governance) [@ross2015, @ross2015,
    @wiens2014, @deacon2011] or economic (the relationship between
    resource wealth and economic growth or human capital) resource
    curses. Empirical investigation of the economic resource curse has
    explored the effect of resource dependence on economic growth,
    public health and education expenditure and outcomes, mainly at a
    national level [@borge2015, @cockx2016, @douglas2017,
    @sincovich2018]. In the case of education, the distinct outcome
    measured is level of educational attainment, in other words, whether
    the presence of a booming resource extraction economy provides
    disincentives to education for young people. It is worth noting that
    this literature has been repeatedly questioned on theoretical and
    conceptual grounds as institutional context often dictates whether a
    resource curse exists and empirical analyses seem to be very
    sensitive to methodological choices [@ross2018, @ross2015,
    @dialga2022]. Although awareness of this strand of literature is of
    relevance to this work, the unresolved nature of the 'debate'
    surrounding its existence requires caution if eventually utilised as
    a theoretical framework for answering the research question.

@ahlerup2020 find that for 30 countries in Africa, the presence of gold
mines during adolescence have a significant effect on educational
attainment. @badeeb2017 investigates whether resource dependence slows
economic growth with no explicit mention of education. @blanco2012 find
that in Latin America, petroleum export has a significant long-run
negative relationships with human capital. @borge2015 find support for
the paradox of plenty hypothesis in Norway - that higher local public
revenue negatively affects the efficiency of local public good
provision. @brunnschweiler2008 critically evaluate 'the empirical basis
for the so-called resource curse and find that, despite the topic's
popularity in economics and political science research, this apparent
paradox may be a red herring. The most commonly used measure of
“resource abundance” can be more usefully interpreted as a proxy for
“resource dependence”—endogenous to underlying structural factors. In
multiple estimations that combine resource abundance and dependence,
institutional, and constitutional variables, we find that (i) resource
abundance, constitutions, and institutions determine resource
dependence, (ii) resource dependence does not affect growth, and (iii)
resource abundance positively affects growth and institutional quality.'
@cockx2014 use a panel on 140 countries from 1995-2009 and find an
inverse relationship between resource dependence and and public health
spending over time. @cockx2016 investigate a panel of 140 countries from
1995-2009 to find an adverse effect of resource depdence on public
education expenditures relative to GDP. @dialga2022 find disparate
results for health and education controlling for institutional quality.
@douglas2017 "measure the effect of resource-sector dependence on
long-run income growth using the natural experiment of coal mining in
409 Appalachian counties selected for homogeneity. Using a panel data
set (1970–2010), we find a one standard deviation increase in resource
dependence is associated with 0.5–1 percentage point long-run and a 0.2
percentage point short-run decline in the annual growth rate of per
capita personal income. We also measure the extent to which the resource
curse operates through disincentives to education, and find significant
effects, but this “education channel” explains less than 15 percent of
the apparent curse.' @haber focus on authoriarian regimes. @menaldo2016
argues again that this is an institutions curse and not a resource curse
issure. @sincovich2018 provide a literature review of resource curse
investigations in the Australian context

Altogether, this evidence points to the value of identifying the extent
to which expenditure on public education is reliant on fossil fuel
revenues in particular geographies. In other words, this work aims to
answer the following research questions:

*RQ1:* How does uneven industrial wage growth affect public education
expenditure? *RQ2:* Do intergovernmental transfers alleviate
wealth-driven inequalities in public ed exp? *RQ3:* Can accounting for
non-constant relationships between explanatory variables improve our
understanding of the relationship between public goods and uneven
economic growth?

In the sections that follow, I outline in @sec-data the data to be used
in the analysis; @sec-methods the proposed methods;

# Data {#sec-data}

This work will make use of Willamette University's Annual Government
Finance Database. This resource is a harmonised repository of the data
collected annually as part of the US Census Bureau's Annual Survey of
State & Local Government Finances, the 'only comprehensive source of
information on the finances of local governments in the United States'
@pierson . I aggregate school district measures up to the county-level
to ensure the availability of adequate control and treatment variables.
[^4]

[^4]: The database is provided for six different levels of government:
    state, county, municipal, township, special district, and school
    district. Reporting is only mandated in Census years (every five
    years), and even then missing data remains a challenge. This means
    that data provided at any other level of government suffers from
    significant levels of missing data, with a high level of selection
    bias correlated with administrative capacity. However, strengthened
    by a partnership with the National Center for Education Statistics,
    observations for US school districts exhibit near-complete coverage
    between 1997-2021 @pierson . I choose to conduct the analysis on the
    commuting zone level because of a lack of availability of control
    variables at a school district level.

Thus, this dataset provides estimates in current \$USD on total public
school revenue disaggregated by source (federal, state, local
intergovernmental versus own local sources) and expenditure
disaggregated by item (level of schooling, teacher salaries, debt,
etc.). Finally, I gather GDP control variables from the Bureau of
Economic Analysis (BEA). This BEA data is only available after 2001,
therefore the panel reported and used below is restricted to 2001-2021.
This results in a complete and balanced panel of 2,710 of 3,143 US
counties between 2001-2021. [^5] [^6]

[^5]: The reason 13% of counties are missing from the dataset is because
    of (1) the exclusion criteria already outlined; (2) Hawaii and
    Alaska have been excluded due to the methodological challenge of
    incorporating their school districts into spatial econometric work;
    and (3) Connecticut, Maryland, North Carolina, and Virginia have
    been excluded due to unconventional or incomplete public school
    district reporting. I aim to resolve this, especially in the case of
    Virginia given its relatively high rates of employment in the coal
    sector.

[^6]: Given the work's intent to rely on data on property taxes
    collected, any county that reports more than five 0 values for
    property taxes collected is excluded.

```{r, echo = FALSE, message = FALSE, cache = FALSE}

options(width = 500)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(conflicted)
library(stargazer)
library(quantreg)
library(purrr)
library(zoo)
library(tidycensus)
library(here)
library(viridis)
library(datasets)
library(broom)
library(xtable)
library(fipio)
library(panelvar)
library(fixest)
library(paletteer)
library(readxl)
library(ggrepel)
library(plm)
library(getspanel)
library(usmap)
library(patchwork)
library(kableExtra)
library(lattice)
library(lubridate)
library(stringr)
library(xts)
library(urca)
library(dynlm)
library(tseries)
library(quantmod)
library(lmtest)
library(sandwich)
library(patchwork)
library(vars)
library(assertthat)
library(ggridges)
library(grid)
library(ggplotify)
source(here('code/source_code/useful_functions.R'))
source(here('code/source_code/dicts.R'))
source(here('code/reg_forms.R'))
source(here('code/source_code/cz_cleaning.R'))
conflict_prefer_all("dplyr", quiet = TRUE)

setFixest_etable(fitstat = ~ . + ivf1 + ivf1.p + ivf2 + ivf2.p + wh + wh.p + ivwaldall + ivwaldall.p, se.below = TRUE)

unit_id = "cz_id"

source(here("code/source_code/pull_data.R"))

```

All data used is reported annually at the commuting zone level. [^7]

[^7]: In line with similar work on US economic geography, commuting
    zones were chosen as the unit of analysis as they are a far less
    arbitrary and more accurate representation of local labour market
    areas/economies
    \\href{https://www.ddorn.net/data.htm\\#Local\\%20Labor\\%20Market\\%20Geography}{David
    Dorn's Resource Page},
    \\href{https://www.nature.com/articles/s41597-024-03829-5}{Fowler et
    al. 2024}.}

Therefore, no time-invariant variables are included (apart from the
State in which a commuting zone is in). Finally, our data represents 636
commuting zones in 40 states between 2001-2021.

**Expenditure and Revenue:** The dependent variables of interest come
from [Willamette University's Government Finance
Database](https://willamette.edu/mba/research-impact/public-datasets/index.html).
The data includes commuting-zone level revenue and expenditure on public
education including disaggregated values by revenue source (federal,
state, or other intergovernmental revenue) and expenditure item
(lunches, wages, debt). All values are reported in real US dollars. The
data for property taxes collected used in regressions below also come
from this dataset. Expenditure on vocational training and from
Educational Service Agencies (ESAs) are also sourced from this dataset.

**GDP Controls:** US Bureau of Economic Analysis. Values are also
reported in current US dollars (real GDP values exist). The controls
used in the below are total, private industry, and oil, gas, mining and
quarrying commuting zone-level GDP.

**Population controls:** US Census Bureau.

**Property Prices:** The US Federal Housing Financy Agency provides a
geographically linked data on single-family house prices called the
Housing Price Index. HPI is a broad measure of the movement of
single-family house prices. The FHFA HPI is a weighted, repeat-sales
index, meaning that it measures average price changes in repeat sales or
refinancings on the same properties. This information is obtained by
reviewing repeat mortgage transactions on single-family properties whose
mortgages have been purchased or securitized by Fannie Mae or Freddie
Mac since January 1975 [Source](https://www.fhfa.gov/data/hpi). It is
reported at the county level at an annual frequency. We aggregate to the
commuting zone level via a mean.
\textcolor{red}{should do a population weighted average!}

<!-- **Coal mine activity and production levels:** Mine Safety and Health -->

<!-- Administration -->

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.height = 10, cache = TRUE}

mean_df <- mines_cz %>% 
  group_by(year) %>%
             summarise(real_Total_Educ_Total_Exp = mean(real_Total_Educ_Total_Exp, na.rm = TRUE),
                       real_Total_Educ_Total_Exp_pp = mean(real_Total_Educ_Total_Exp_pp, na.rm = TRUE),
                       real_log_Total_Educ_Total_Exp = mean(log_real_Total_Educ_Total_Exp, na.rm = TRUE),
                       real_log_Total_Educ_Total_Exp_pp = mean(log_real_Total_Educ_Total_Exp_pp, na.rm = TRUE))

```

## Summary statistics

```{r, eval = FALSE, echo = FALSE, cache = FALSE}

is.pbalanced(mines_cz, index = c("cz_id", "year"))

```

All dollar values are reported in real 2017-chained thousands.

```{r, results = 'asis', cache = TRUE, message = FALSE}

# Variables of interest for descriptive statistics
vars <- c("Property_Tax", "Total_IG_Revenue", "Total_Fed_IG_Revenue", "Total_State_IG_Revenue", "Total_Educ_Total_Exp", "Elem_Educ_Total_Exp") # Total_Revenu

mines_cz %>%
  select(Enrollment, pop_total, real_Elem_Educ_Total_Exp_pp, real_Property_Tax_pp, real_Total_IG_Revenue_pp,# real_Total_Fed_IG_Revenue_pp, 
         real_Total_State_IG_Revenue_pp, 
         real_gdp_total_pc, real_gdp_priv_ind_pc, #real_gdp_o_g_mining_quarr_21_pc, 
         hpi) %>% 
         #real_Elem_Educ_Total_Exp, real_Property_Tax, real_Total_IG_Revenue, real_Total_Fed_IG_Revenue, 
         #real_Total_State_IG_Revenue, real_gdp_total, real_gdp_priv_ind, real_gdp_o_g_mining_quarr_21, total_active_n, total_active_prod) %>% 
  mutate(across(c(Enrollment, pop_total), ~./1000)) %>%  # gdp_govt, 
          data.frame() %>%
          stargazer(digits = 0, digits.extra = 3, covariate.labels = c(
         "Enrollment",
         "Population",
         "Elem. Expenditure per pupil",
         "Property Tax per pupil",
         "IG Revenue per pupil", 
         #"Federal IG Revenue per pupil", 
         "State IG Revenue per pupil",
         "GDP per capita",
         "GDP pc - Private Industry",
         "House Price Index"),
         header = FALSE)
         #"GDP pc - Oil, gas, mining",
         #"Elem. Expenditure",
         #"Property Tax",
         #"IG Revenue", 
         #"Federal IG Revenue", 
         #"State IG Revenue",
         #"GDP",
         #"GDP - Private Industry", 
         #"GDP - Oil, gas, mining",
         #"Active Coal Mines",
         #"Coal Produced (k short tons)"))


```

<!-- ##Descriptive Regression Results - Appendix -->

```{r appendix1, include = FALSE, echo = FALSE, results = 'asis', cache = TRUE}

append_to_appendix("## Descriptive Regression Results",
"In the following set of results, I report descriptive regressions to establish relationships between property taxes, education expenditure, GDP (total, private industry, O&G&mining), etc.
All regression models that follow include TWFE (CZ- and year- fixed effects) and standard errors clustered by commuting zone.
All functional forms in the feols() functions below are of the form Y ~ X In the cases in which multiple estimations are included via sw(Xa, Xb, Xc + Xd), the function will return results for Y~Xa, Y~Xb, Y ~ Xc + Xd.")


```

```{r appendix2, include = FALSE, echo = FALSE, results = 'asis', cache = TRUE}

append_to_appendix("### Property Tax ~ GDP",
               "GDP has a highly relevant relationship to property taxes. A 1% increase in GDP (per capita) leads to a 0.38% (0.32%) increase in property taxes collected (per capita).")

run_model(prop_taxes, mines_cz) %>%
  etable(tex = TRUE, adjustbox = TRUE)

```

```{r appendix3, include = FALSE, echo = FALSE, results = 'asis', cache = TRUE}

append_to_appendix("### Education Expenditure ~ Revenue Sources",
               "The below regressions are included to establish the relationship between education expenditure and its component parts. These regressions simply corroborate what is displayed in the section on Key Relationships in [LINK](https://ebbam.github.io/pub_fin_schools/code/cz_results.html) (ie. that the largest form of IG revenue is state funding and Own Source revenue is largely sourced from Property Taxes).")

run_model(educ_source, mines_cz) %>% etable(adjustbox = TRUE, tex = TRUE)

```

```{r appendix4, include = FALSE, echo = FALSE, results = 'asis', cache = TRUE}
append_to_appendix("### Education Expenditure ~ GDP",
               "A 1% increase in GDP pc is associated with a 0.19% increase in education expenditure per pupil, dominated by the effect of GDP from private industry (0.16%). I include here also the GDP generated from the oil, gas, mining, and quarrying sector. The effect is small and statistically insignificant.")

run_model(educ_gdp, mines_cz) %>% etable(adjustbox = TRUE, tex = TRUE)
```

```{r, include = FALSE, echo = FALSE, results = 'asis', cache = TRUE}

common_theme <- theme(
  axis.title.x = element_text(size = 16),
  axis.title.y = element_text(size = 16),
  axis.text.x = element_text(size = 14),
  axis.text.y = element_text(size = 14)
)

gdp_plot <- mines_cz %>%
  ggplot(aes(x = log_real_gdp_priv_ind_pc, y = log_real_Elem_Educ_Total_Exp_pp)) +
    geom_jitter(, colour = "darkorange2", alpha = 0.5) +
    geom_smooth(method=lm, colour = "black")  +
    labs(x = "log Priv. Industry GDP pc", y = "log Education Expenditure pp") +
  common_theme

hpi_plot <- mines_cz %>%
  ggplot(aes(x = log_hpi, y = log_real_Elem_Educ_Total_Exp_pp)) +
    geom_jitter(, colour = "maroon", alpha = 0.5) +
    geom_smooth(method=lm, colour = "black")  +
    labs(x = "log House Prices (Index Value)", y = "log Education Expenditure pp")+
  common_theme

wage_plot <- mines_cz %>%
  ggplot(aes(x = log_weighted_annual_avg_wkly_wage, y = log_real_Elem_Educ_Total_Exp_pp)) +
    geom_jitter(, colour = "darkorchid", alpha = 0.5) +
    geom_smooth(method=lm, colour = "black")  +
    labs( x = "log Annual Average Weekly Wage", y = "log Education Expenditure pp")+
  common_theme

pop_plot <- mines_cz %>%
  ggplot(aes(x = log_Enrollment, y = log_real_Elem_Educ_Total_Exp_pp)) +
    geom_jitter(, colour = "steelblue", alpha = 0.5) +
    geom_smooth(method=lm, colour = "black")  +
    labs(x = "log Enrollment", y = "log Education Expenditure pp")+
  common_theme

wage_prop_values <- mines_cz %>%
  ggplot(aes(x = log_weighted_annual_avg_wkly_wage, y = log_hpi)) +
    geom_jitter(, colour = "darkgreen", alpha = 0.5) +
    geom_smooth(method=lm, colour = "black")  +
    labs( x = "log Wage", y = "log House Prices (Index Value)")+
  common_theme

(gdp_plot + hpi_plot) / (wage_plot + pop_plot)+ plot_annotation(title = "Key Relationships Between Economic Varables and Ed.Exp.",
                                                                 subtitle = "Commuting Zone Level. 2001-2020.",
                                                                 theme = theme(plot.title = element_text(size = 18)))

mines_cz %>%
  ggplot(aes(x = log_real_gdp_priv_ind_pc, y = log_real_Total_IG_Revenue_pp)) +
    geom_jitter(, colour = "brown", alpha = 0.5) +
    geom_smooth(method=lm, colour = "black")  +
    labs(title = "IG Transfers (pp) vs. GDP pc", x = "log Priv. Industry GDP pc", y = "log IG Revenue pp") +
  common_theme

```

First, shift-share or *Bartik* instruments have gained popularity in
empirical work as a method of handling endogeneity issues in panel data
[@ferri2022, @goldsmith-pinkham2020, @bartik1991]. Effectively, they
combine time-variant yet unit-invariant changes in aggregate economic
variables (ie., changes in national fossil fuel employment levels) with
time-invariant yet unit-variant shares in exposure to these macro-level
changes (ie., local shares of employment in fossil fuels). This
decomposition of local-level changes via a delocalisation over space and
time allows for a defensible 'de-endogenising' of the treatment.
Notably, the method can also be considered to serve a further purpose,
allowing for, by construction, the examination of a macro phenomenon's
effect on more local units.

\\footnote{Autor et al use a shift-share instrument to assess the effect
of Chinese import competition on manufacturing employment in US counties
@autor2013 . As an extension, @feler2017 use a similar shift-share
instrument to assess the effect of the same shock on the size of local
government. @baccini2021 employ a shift-share instrument for
manufacturing layoffs to tease out the effect of a decline in
manufacturing on both economically motivated and racial identity voting
patterns in the US.}

Second, an additional popular indicator for modelling industrial shocks
is *oil price* as values are often assumed to be exogenous to local and
even national conditions @scheer2022 . Third, separate from specifically
green transformation shocks, various indicators for measuring
*deindustrialisation* have been proposed including the manufacturing
share of employment, value added, and GDP [@tregenna2009,
@tregenna2020]. These deindustrialisation metrics could be used in
combination with a shift-share instrument. Finally, in rare instances,
exogeneity can be secured due to *geographical, climatological, or
geological factors*. For example, @borge2015 obtain an exogenous measure
of local revenue by "instrumenting the variation in hydropower revenue,
and thus total revenue, by topology, average precipitation and meters of
river in steep terrain." Certain authors have argued that the fact that
the location of hydrocarbon deposits is dictated by geomorphological
processes provides a plausible argument for exogeneity [@esposito2021,
@chen2022].

In order to arrive at a final estimation strategy, I plan to explore
various treatment variables to proxy industrial transformation from the
outlined alternatives. Thus far, I have employed data on coal mining
activity in a two-way fixed effects ordinary least squares framework and
an instrumental variable approach as well as constructed, although not
yet estimated, shift-share instruments using employment and wage shares
in the coal, oil and gas, and extraction sectors. I present preliminary
results from these estimation efforts below. I do not yet claim that
these results provide a causal interpretation nor that they are robustly
specified. However, I provide these results as evidence of preliminary
attempts to answer the outlined research questions as well as important
building blocks toward eventually arriving at a credible estimation
strategy.

<!-- ### Coal Mining Activity -->

<!-- First, the Mining Safety and Health Administration (MSHA) provides data -->

<!-- at the mine level on activity and production. I have thus far aggregated -->

<!-- this data for a panel of annual county-level estimates of the number of -->

<!-- active coal mines and production volumes (reported in thousand short -->

<!-- tons) between 1996-2021. I restrict the sample to 2001-2021 to align -->

<!-- with the county finance -->

<!-- data.\footnote{I do not yet make a distinction between different types of mining or coal product (ie. surface and undeground mines; (sub)bituminous, lignite, versus anthracite coal), although such distinctions are possible using the data collected.} -->

# Analysis {#sec-methods}

Given the high degree of both structural (state-specific tax,
regulatory, and legislative regimes) and evolved heterogeneity
(industrial activity, income, inequality, economic diverrsity) the
following analysis only briefly explores the potential to arrive at
national-level average treatment effects using various pooled estimation
strategies. These serve to establish foundational relationships between
local economic conditions that seem to reasonably generalise across the
country.

Considerable investigation uncovered the need for a regional or
state–by-state estimation strategy to truly account for the
heterogeneity across the units of observation. Therefore, the body of
the analysis is dedicated to state-by-state estimation of the relevant
econometric specifications. We dedicate majority of this manuscript to
discussing the consequences of an uneven legislative landscape on
educational expenditure outcomes.

## Descriptive Regressions

### Baseline

First, I employ a two-way fixed effects ordinary least-squares panel
model with standard errors clustered by commuting zone. I outline the
model specification immediately below:

\begin{equation}
    Y_{it} = \beta_{0} + \beta_{x}X_{it} + \delta_1 Enrollment_{it} + \delta_2 IGR_{it} +\alpha_i + \gamma_t + \varepsilon_{it} 
\label{desc_function}
\end{equation}
\begin{equation}
    Y_{it} = \beta_{0} + \beta_{x}X_{it} + \delta_1 Enrollment_{it} + \delta_2 IGR_{it} + \lambda Y_{it-1} +\alpha_i + \gamma_t + \varepsilon_{it} 
\label{desc_functio_ar1}
\end{equation}

$Y_{it}$ is the natural logarithm of elementary (serving ages 6-12) education expenditure per pupil for
county $i$ in year $t$. $\alpha_i$ and $\gamma_t$ represent county- and
year-fixed effects, respectively, and $\varepsilon_{it}$ represents the
error term. We control for enrollment to account for scaling factors in education expendture and intergovernmental transfers to account for the significant role of such transfers in funding education expenditure. Additionally, we optionally include a lagged dependent variable to account for the persistence of education expenditure. $X_{it}$ takes three forms represented by
\@ref(gdp_function), \@ref(wage_function), \@ref(hpi_function) where $h = [0,2]$ in the specifications that
incorporate time lagged variables. We estimate all equations in both levels and growth rates. 

\begin{equation}
    X_{it}^{GDP} = \sum_{h=0}^{2} \beta_k^{\text{GDP}} \log(\text{GDP}_{it-h})
\label{gdp_function}
\end{equation}
\begin{equation}
    X_{it}^{Wage} = \sum_{h=0}^{2} \beta_k^{\text{Wage}} \log(\text{Wage}_{it-h})
\label{wage_function}
\end{equation}
\begin{equation}
    X_{it}^{HPI} = \sum_{h=0}^{2} \beta_k^{\text{HPI}} \log(\text{HPI}_{it-h})
\label{hpi_function}
\end{equation}

Education expenditure has a highly relevant time dependence. The effect
of increases in GDP two years prior has the greatest effect on current
education expenditure, implying a delayed effect of commuting zone-level
economic growth on public education expenditure.

Table X reports a series of fixed effects regressions estimating the
association between local economic conditions and elementary education
expenditure per pupil, expressed in both levels (left panel) and growth
rates (right panel). Across all specifications, the results underscore
the central role of intergovernmental transfers and broader local
economic fundamentals in shaping patterns of education spending.

In the levels regressions, intergovernmental revenue per pupil emerges
as the strongest and most consistent predictor of education expenditure.
A 1% increase in intergovernmental transfers is associated with
approximately a 0.35–0.38% increase in per-pupil education spending,
controlling for unit and year fixed effects. This finding highlights the
importance of state and federal aid in sustaining local education
budgets. Lagged economic indicators, particularly private industry GDP
and average weekly wages, are also positively and significantly
associated with education spending. The magnitude of the coefficients
increases with the number of lags, suggesting a gradual adjustment
process by which local economic growth translates into higher public
investment in education over time. For example, a 1% increase in lagged
(t–2) real private GDP per capita is associated with a 0.14% increase in
per-pupil spending, while lagged wage growth shows similar cumulative
effects.

The house price index also enters positively and significantly in
contemporaneous and short-lag specifications (up to t–3), but its
influence diminishes and turns negative at longer lags. This pattern is
consistent with the hypothesis that local housing markets influence
education budgets through changes in property tax bases and local fiscal
capacity, but that these effects are most salient in the short run and
may attenuate over time due to institutional lags in budgeting and
allocation.

The growth rate regressions, while explaining less variance overall (as
expected), largely confirm the patterns observed in the levels
specifications. Intergovernmental revenue growth remains a strong and
highly significant determinant of education expenditure growth, with
coefficients ranging from 0.33 to 0.40. Lagged wage and GDP growth also
emerge as important predictors, particularly at longer lags. Notably,
wage growth two years prior is associated with a 0.31% increase in
education spending growth, suggesting that labor market improvements
take time to materialize in local education budgets. Growth in house
prices also predicts increases in education spending, but these effects
are concentrated in the contemporaneous and first-lag specifications.

Taken together, these results offer three key insights. First, public
education investment is strongly mediated by external fiscal flows,
reaffirming the role of intergovernmental transfers in equalizing local
education finance. Second, local labor market conditions—captured
through wages and GDP—exert lagged, cumulative effects on education
spending, consistent with fiscal inertia and political delays in
budgetary response. Third, local housing markets play a more modest and
short-term role in shaping education budgets, reflecting the link
between property values and tax revenues.

```{r, message = FALSE, warning = FALSE, results = 'asis'}

run_model(c(log_real_Elem_Educ_Total_Exp_pp ~ sw(
  log_real_gdp_priv_ind_pc + l1_log_real_gdp_priv_ind_pc + l2_log_real_gdp_priv_ind_pc,
   log_real_gdp_priv_ind_pc + l1_log_real_gdp_priv_ind_pc + l2_log_real_gdp_priv_ind_pc + l1_log_real_Elem_Educ_Total_Exp_pp,
  log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage, 
  log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage + l1_log_real_Elem_Educ_Total_Exp_pp,
  log_hpi + l1_log_hpi + l2_log_hpi + l3_log_hpi + l4_log_hpi, 
  log_hpi + l1_log_hpi + l2_log_hpi + l3_log_hpi + l4_log_hpi + l1_log_real_Elem_Educ_Total_Exp_pp
) + log_real_Total_IG_Revenue_pp + log_Enrollment | unit + year), mines_cz) %>% etable(adjustbox = TRUE, tex = TRUE)

run_model(c(diff_log_real_Elem_Educ_Total_Exp_pp ~ sw(
  diff_log_real_gdp_priv_ind_pc + l1_diff_log_real_gdp_priv_ind_pc + l2_diff_log_real_gdp_priv_ind_pc,
  diff_log_real_gdp_priv_ind_pc + l1_diff_log_real_gdp_priv_ind_pc + l2_diff_log_real_gdp_priv_ind_pc+ l1_diff_log_real_Elem_Educ_Total_Exp_pp,
  gr_weighted_annual_avg_wkly_wage + l1_gr_weighted_annual_avg_wkly_wage + l2_gr_weighted_annual_avg_wkly_wage, 
  gr_weighted_annual_avg_wkly_wage + l1_gr_weighted_annual_avg_wkly_wage + l2_gr_weighted_annual_avg_wkly_wage + l1_diff_log_real_Elem_Educ_Total_Exp_pp,
  gr_hpi + l1_gr_hpi + l2_gr_hpi + l3_gr_hpi + l4_gr_hpi,
  gr_hpi + l1_gr_hpi + l2_gr_hpi + l3_gr_hpi + l4_gr_hpi+ l1_diff_log_real_Elem_Educ_Total_Exp_pp) + diff_log_real_Total_IG_Revenue_pp + diff_log_Enrollment | unit + year), mines_cz) %>% etable(adjustbox = TRUE, tex = TRUE)

```

Furthermore, given the heterogeneity in reliance on intergovernmental transfers (largely coming from the state), we interact all economic predictors above with a variable that represents the share of total elementary education expenditure coming from state-level funding. 

```{r, message = FALSE, warning = FALSE, results = 'asis'}

selected_twfe_models_state_share_interaction = 
  c(log_real_Elem_Educ_Total_Exp_pp ~ sw(state_share*(log_real_gdp_priv_ind_pc + l1_log_real_gdp_priv_ind_pc + l2_log_real_gdp_priv_ind_pc), state_share*(log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage), state_share*(log_hpi + l1_log_hpi + l2_log_hpi + l3_log_hpi + l4_log_hpi + l5_log_hpi)) + state_share*log_real_Total_Fed_IG_Revenue_pp + state_share*log_Enrollment | unit + year,
    log_real_Elem_Educ_Total_Exp_pp ~ sw(state_share*(log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage), state_share*(log_hpi + l1_log_hpi + l2_log_hpi + l3_log_hpi + l4_log_hpi + l5_log_hpi)) + state_share*log_real_Total_Fed_IG_Revenue_pp + state_share*log_Enrollment + l1_log_real_Elem_Educ_Total_Exp_pp | unit + year)


run_model(selected_twfe_models_state_share_interaction, mines_cz) %>% etable(adjustbox = TRUE, tex = TRUE)
#run_model(selected_twfe_models_levs_state_share_interaction[1:2], mines_cz) %>% etable(tex = TRUE)
#run_model(selected_twfe_models_levs_state_share_interaction[3:4], mines_cz) %>% etable(tex = TRUE)

```

<!-- ### Incorporating state-level trends -->

<!-- The below take the Education Expenditure \~ GDP models and incorporate -->

<!-- deterministic state time trends. -->

<!-- ```{r, message = FALSE, warning = FALSE, results = 'asis'} -->

<!-- # run_model(selected_twfe_models_state_trend[1:2], mines_cz) %>% -->

<!-- #   etable(drop = "*time", tex = TRUE) -->

<!-- run_model(selected_twfe_models_state_trend[3:4], mines_cz) %>% -->

<!--   etable(drop = "*time", tex = TRUE, adjustbox = TRUE) -->

<!-- ``` -->

## Instrumental Variable Approach

There is a significant endogeneity concern in using wage growth as the
treatment variable. Therefore, I have tried two instrumental variable
approaches below and aim to add results using production- and
employment-based Bartik instruments.

A critical break-through in this work came from understanding the
structure of public financing (described in further detail in Section
\ref{si_section:Challenges} of the Supplementary Materials). In brief,
revenue for public education comes from a combination of
intergovernmental and local sources and revenue generated from local
sources comes almost entirely from property taxes. Given this, I can
isolate the channel through which I assume our treatment (industrial
production) will affect our outcome variable using an instrumental
variable approach. I provide the results of an initial attempt at this
theory below including the functional form and underlying theoretical
path diagram of the econometric specification.

We consider using XX as an instrument affecting education expenditure
through property taxes or GDP. We know that property taxes have an
endogenous relationship with education expenditure, however, in theory,
XX is unlikely to affect education expenditure, except via property
taxes. We test this hypothesis below.

As a reminder, the intuition behind the idea is

\begin{figure}[H]
\centering
\caption{Instrumental Variable Path Diagram}
\begin{tikzpicture}
% nodes %
\node[rectangle] (z) at (0,0) {Instrument};
\node (t) at (4,0) {Property Values};
\node (y) at (8,0) {Education Expenditure};
\node (u) at (4,2) {Omitted variable};

% edges %
\draw[->, line width=1] (z) -- (t);
\draw[->, line width=1] (t) -- (y);
\draw[->, line width=1, red, dashed] (u) -- (t);
\draw[->, line width=1, red, dashed] (u) -- (y);
\end{tikzpicture}
\label{fig:iv_approach}
\end{figure}

As seen in Figure \ref{fig:iv_approach}, I hypothesize that property
taxes have an effect on education expenditure. However, there is
significant concern of a reverse causal effect as higher income families
likely gravitate towards school districts with higher levels of
expenditure per pupil, driving up property values. However, we assume
that coal mining activity has no other effect on education expenditure
except through the channel of property taxes.

A more commonly used identification strategy is via a shift-share or
Bartik instrument. A shift-share instrument interacts local industry
shares with national industry-level growth rates to attain a plausibly
exogenous local shock. In the context of this work, we intend to create
a unit-specific time-varying treatment variable by interacting a
unit-specific, time-invariant industrial employment share variable with
a national-level time-varying wage growth rate.

$Y_{it}$ takes the same two values as above. $Z_{i,t}$ takes the value
of either active coal production or the number of active coal mines in
county $i$ in year $t$. I further test for a time lagged effect and run
alternate specifications where a one- and two-year lag $Z_{i,t}$
replaces the contemporaneous value. $X_{i,t}$ is the amount of property
taxes collected per pupil in county $i$ in year $t$.

The literature on Bartik instruments derives plausible exogeneity from
two sources. First, authors argue that local industry shares are
exogenous by imposing that shares be fixed to a particular base year and
are therefore unable to adapt to changes in national-level growth rates.
Such a shift-share instrument would look as follows:

\begin{equation}
Z_{it} = \sum_{j=1}^{k} S_{ij\tau}G_{njt}
\end{equation}

where $S_{ij0}$ is the local share of unit $i$'s economy (potentially
measured by metrics like employment, wages, revenue) in industry $j$ at
a fixed base year $\tau$ and $G_{njt}$ is the growth rate of industry
$j$ at a national level $n$ at time $t$.

Alternatively, authors may argue that the national-level growth rates
are exogenous allowing the shares to vary over time, constructing the
shift-share instrument as follows:

\begin{equation}
Z_{it} = \sum_{j=1}^{k} S_{ijt}G_{njt}
\end{equation}

Finally, authors might be concerned about the implausible exogeneity of
both shares and national-level growth rates in which case they could
construct the instrument as follows where the local shares are fixed at
a common base year and industry-specific growth rates $G$ are derived
from data on other similar regions $o$ rather than national-level
changes that are inherently comprised of local-level shifts. This
approach likely comes at significant expense to instrument relevance.

\begin{equation}
Z_{it} = \sum_{j=1}^{k} S_{0jt}G_{ojt}
\end{equation}

Finally, the authors can make an additional design choice about whether
the effect of these instruments should be assumed common to an aggregate
local-level wage growth indicator or allowed to vary by industry. In
other words, whether to construct the first-stage relationship of the
2SLS as:

\begin{equation}
X_{it} = \alpha_i + \beta\sum_{j=1}^{k}S_{\*j\*}G_{\*j} + \epsilon_{it}
\end{equation} \begin{equation}
X_{it} = \alpha_i + \sum_{j=1}^{k}\beta_{j}S_{\star j \star}G_{\star jt} + \epsilon_{it}
\end{equation}

```{r, include = FALSE}

# getspanel
# isatpanel(data = mines_cz,
#           formula = "Total_Educ_Total_Exp_pp ~ gdp_total_pc + gdp_priv_ind_pc + gdp_o_g_mining_quarr_21_pc + gdp_govt_pc",
#           index = c("unit","year"),
#          iis = TRUE,
#          t.pval =0.01
#          # SIS = TRUE
#           )

```

### Shift-share Instrument

Finally, using data from the US Bureau of Labor Statistics' Quarterly
Census of Employment and Wages (QCEW), I have constructed a total of six
shift-share Bartik instruments across two metrics (employment and wage)
for three sectoral aggregations (oil and gas, coal, and all extractive
industries) at the county-level. Equation
\ref{bartik} demonstrates the Bartik instrument as outlined in
@ferri2022 and @goldsmith-pinkham2020 and defined in @bartik1991 .
$\Delta X_{t,s}$ represents national-level changes in either employment
or wages in industry $s$ in time $t$ and
$\frac{N_{i,\tau, s}}{N_{i,\tau}}$ represents the \`\`sensitivity'' of a
county to these national shocks proxied by an initial share of local
employment or wage in industry $s$ in a baseline time period $\tau$. The
product of these two values defines the shift-share indicator
$\tilde{X}_{i,t,s}$. In Figure XX, I demonstrate the
wage- and employment-based shift-share instruments (each line represents
a separate county) using a base period $\tau$ = 2001. [^8]

[^8]: We explore the sensitivity of results to the choice of base period
    $\tau$ by constructing the instrument for various base periods as
    well as a rolling window.

\begin{equation}
    \tilde{X}_{i,t,s} = \Delta X_{t,s} *  \frac{N_{i,\tau, s}}{N_{i,\tau}}
\label{bartik}
\end{equation}

<!-- ![Coal Shift-share -->
<!-- Instruments](../output/bartik_graphs.jpg){#fig-bartik_graph} -->

```{r, fig.height = 10, fig.width = 8, echo = FALSE, warning = FALSE, message = FALSE}

# Draw the functions needed to calculate the SS instruments from the following script.
source(here("data/raw/QCEW/industry_shares_cleaning.R"))

```

```{r compute_ss, echo = FALSE, cache = TRUE}

ss_shares <- compute_shares(source = "QCEW", base_year = 2004, unit_id = "cz_id", flat = FALSE) 

coverage <- ss_shares %>% select(coverage_2digit_naics, coverage_3digit_naics) %>% distinct
  
ss_temp <- compute_ss(ss_shares)

```

```{r, echo = FALSE, cache = FALSE}

coverage %>% 
  ggplot() +
  geom_histogram(aes(x = coverage_2digit_naics, fill = "2-digit NAICS Codes"), alpha = 0.8) +
  geom_histogram(aes(x = coverage_3digit_naics, fill = "3-digit NAICS Codes"), alpha = 0.8) +
  labs(x = "% Coverage of County's Employed by NAICS sub-categorisation", 
       y = "No. Counties",
       title = "Data Coverage of Industry-level Employment as Share of Total Reported Employed",
       fill = "NAICS Specificity") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  scale_fill_brewer(palette = "Pastel1")

```

```{r, echo = FALSE, cache = TRUE}

temp_new <- shares_flat_filled %>%
  select(!contains("share")) %>%
  filter(year <= 2005) %>%
  group_by(unit) %>%
  # Creates a mean employment level across 2001-2005 - 5 year mean to deal with missing data
  summarise(across(!year, ~mean(., na.rm = TRUE))) %>%
  ungroup  %>%
  rename(fips = unit)

  if(unit_id == "cz_id"){
    czs_new <- czs %>%
      #select(-old_fips) %>%
      rename(old_fips = fips) %>%
      mutate(fips = case_when(!is.na(getfips[old_fips]) ~ getfips[old_fips],
                              TRUE ~ old_fips),
             cz_id = as.character(cz_id))

    missing_fips <- czs_new %>%
      pull(fips) %>%
      unique %>%
      setdiff(unique(temp_new$fips), .)

    if (length(missing_fips) > 0) {
      message("Warning: Some FIPS codes are missing from czs.")
    }

    temp <- temp_new %>%
      select(fips, matches("^annual_avg_emplvl_\\d{2}$"), annual_avg_emplvl_10_filled) %>%
      left_join(., czs_new, by = "fips", multiple = "first") %>%
      rename("unit" = cz_id) %>%
      select(-c(fips, old_fips, cz_population, cz_id_1990)) %>%
      relocate(unit)

  }else if(unit_id == "fips"){
    temp <- temp %>%
      rename(unit = fips)
  }


shift_share_filled <- temp %>%
    group_by(unit) %>%
    summarise(across(everything(), ~sum(., na.rm = TRUE))) %>%  # , total_annual_wages)
              #annual_avg_wkly_wage = mean(annual_avg_wkly_wage, na.rm = TRUE)) %>%
    ungroup %>%
  mutate(across(contains("avg_emplvl"), ~./annual_avg_emplvl_10_filled, .names = "share_{.col}")) %>%
  select(unit, contains('share')) %>%
  mutate(year = 2001) %>%
  complete(unit, year = 2001:2022) %>%
  group_by(unit) %>%
  fill(everything(), .direction = "updown") %>%
  ungroup

assert_that(nrow(shift_share_filled) == n_distinct(shift_share_filled$unit) * n_distinct(shift_share_filled$year))

ss_temp_fill <- shift_share_filled %>%
      select(unit, year, contains("share")) %>%
      left_join(., natl_rates, by = "year", relationship = "many-to-one") %>%
      rename(!!unit_id := unit) %>%
      select(-ends_with("10"))

ss_temp_filled <- compute_ss(ss_temp_fill)

ss_temp_old <- ss_temp
ss_temp <- ss_temp_filled


```

Fundamental to our identification strategy is the relationship between wages and property prices as well as real value added and property prices. We provide evidence of this relationship below. 

#### Industry-level Wages

The two regression models show that wage levels and wage growth both play important roles in influencing house prices, but in different ways. The log-level model indicates that increases in average wages have a strong and persistent effect on the level of house prices, with significant positive effects extending up to four quarters. In contrast, the growth rate model suggests that house price growth responds primarily to contemporaneous wage growth, with little evidence of lagged effects. Together, these findings imply that while higher wages steadily raise housing values over time, short-term changes in wage growth influence house price dynamics more immediately but less durably. Together, they suggest that housing markets are more responsive to trends than to transitory shocks in wages.

```{r, echo = FALSE, cache = TRUE, results = 'asis'}

feols(log_hpi ~ l(log_weighted_annual_avg_wkly_wage, 0:7) + log_real_gdp_priv_ind_pc + log_pop_total | unit + year, panel.id = c('year', 'unit'), mines_cz) %>% etable(tex = TRUE)

feols(gr_hpi ~ l(gr_weighted_annual_avg_wkly_wage, 0:7) + diff_log_real_gdp_priv_ind_pc + diff_log_pop_total| unit + year, panel.id = c('year', 'unit'), mines_cz) %>% etable(tex = TRUE)

```


```{r, echo = FALSE, cache = TRUE, results = 'asis'}
p1 <- ss_temp %>%
  ggplot(aes(x = year, y = ss_2d, group = cz_id)) +
  geom_line() +
  labs(title = "Shift-Share Instrument: 2-digit NAICS",
       x = "SS Instrument: Local Employment Share * National Industry-Specific Wage Growth",
       y = "Year")

p2 <- ss_temp %>%
  ggplot(aes(x = year, y = ss_3d, group = cz_id)) +
  geom_line() +
  labs(title = "Shift-Share Instrument: 3-digit NAICS",
       x = "SS Instrument: Local Employment Share * National Industry-Specific Wage Growth",
       y = "Year")

p3 <- ss_temp %>%
  ggplot(aes(x = year, y = lev_ss_2d, group = cz_id)) +
  geom_line() +
  labs(title = "Shift-Share Instrument: 2-digit NAICS",
       x = "SS Instrument: Local Employment Share * National Industry-Specific Wage Growth",
       y = "Year")

p4 <- ss_temp %>%
  ggplot(aes(x = year, y = lev_ss_3d, group = cz_id)) +
  geom_line() +
  labs(title = "Shift-Share Instrument: 3-digit NAICS",
       x = "SS Instrument: Local Employment Share * National Industry-Specific Wage Growth",
       y = "Year")


p5 <- ss_temp %>%
  ggplot(aes(x = year, y = gdp_ss_2d, group = cz_id)) +
  geom_line() +
  labs(title = "VA Shift-Share Instrument: 2-digit NAICS",
       x = "SS Instrument: Local Employment Share * National Industry-Specific VA Growth",
       y = "Year")

p6 <- ss_temp %>%
  ggplot(aes(x = year, y = lev_ss_3d, group = cz_id)) +
  geom_line() +
  labs(title = "VA Shift-Share Instrument: 2-digit NAICS",
       x = "SS Instrument: Local Employment Share * National Industry-Specific VA Level",
       y = "Year")

p1 + p3
#p2 + p4
p5 + p6

iv_model_2d_lev_gr <- feols(as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~ l(ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_2d_gr_gr <- feols(
    as.formula("diff_log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~ l(ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_2d_lev_lev <- feols(
    as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~ l(lev_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_2d_lev_gr_tfe <- feols(as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | year + state | log_hpi ~ l(ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_2d_gr_gr_tfe <- feols(
    as.formula("diff_log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment| year + state | log_hpi ~ l(ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_2d_lev_lev_tfe <- feols(
    as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | year + state | log_hpi ~ l(lev_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_2d_share <- feols(
    as.formula("diff_log_real_Elem_Educ_Total_Exp_pp ~ log_real_gdp_priv_ind_pc | unit + year + log_Enrollment | log_hpi ~  state_share*l(ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year", "unit")), panel.id = c("unit", "year"))

iv_model_2d_levshare <- feols(
    as.formula("diff_log_real_Elem_Educ_Total_Exp_pp ~ log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~  state_share*l(lev_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year", "unit")), panel.id = c("unit", "year"))

etable(iv_model_2d_lev_lev, iv_model_2d_lev_gr, iv_model_2d_gr_gr, adjustbox = TRUE, stage = 1:2, tex = TRUE)
etable(iv_model_2d_lev_lev_tfe, iv_model_2d_lev_gr_tfe, iv_model_2d_gr_gr_tfe, adjustbox = TRUE, stage = 1:2, tex = TRUE)

```

#### Industry-level GDP

The two regression tables investigate the relationship between real GDP per capita in the private industry and house prices. In the log-level specification, real GDP (including current and lagged values up to 7 periods) is positively and significantly associated with house prices. The contemporaneous coefficient (0.2411) is large and highly significant, and several lags (especially lags 1, 2, 3, and 5) are also positive and statistically significant, indicating a cumulative and persistent effect of GDP on housing values.

In contrast, the growth-rate specification shows a much weaker relationship. Most coefficients are small and not statistically significant, with the exception of lag 5, which shows a modest but significant positive association (0.0073). The R² values further support this contrast: the log-level model explains 96% of the variation in house prices (with a within R² of 0.203), while the growth-rate model explains only about 39% overall (and just 0.005 within variation).

Together, these findings suggest that levels (rather than short-run changes) in GDP are more systematically and substantially associated with house price dynamics. The stronger fit and significant lag effects in the log-level regression underscore the longer-term influence of economic fundamentals on housing markets.

```{r, echo = FALSE, cache = TRUE, results = 'asis'}

feols(log_hpi ~ l(log_real_gdp_priv_ind_pc, 0:7) + log_pop_total | unit + year, panel.id = c('year', 'unit'), mines_cz) %>% etable(tex = TRUE)

feols(gr_hpi ~ l(diff_log_real_gdp_priv_ind_pc, 0:7) +  diff_log_pop_total| unit + year, panel.id = c('year', 'unit'), mines_cz) %>% etable(tex = TRUE)

```


```{r, echo = FALSE, cache = TRUE, results = 'asis'}

iv_model_gdp_2d_lev_gr <- feols(as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~ l(gdp_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_gdp_2d_gr_gr <- feols(
    as.formula("diff_log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~ l(gdp_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_gdp_2d_lev_lev <- feols(
    as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year | log_hpi ~ l(lev_gdp_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_gdp_2d_lev_gr_ar1 <- feols(as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + l1_log_real_Elem_Educ_Total_Exp_pp + log_Enrollment | unit + year | log_hpi ~ l(gdp_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_gdp_2d_gr_gr_ar1 <- feols(
    as.formula("diff_log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc  + l1_diff_log_real_Elem_Educ_Total_Exp_pp + log_Enrollment | unit + year | log_hpi ~ l(gdp_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))

iv_model_gdp_2d_lev_lev_ar1 <- feols(
    as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc  + l1_log_real_Elem_Educ_Total_Exp_pp + log_Enrollment | unit + year | log_hpi ~ l(lev_gdp_ss_2d,1)"),
    data = left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")), panel.id = c("unit", "year"))


etable(iv_model_gdp_2d_lev_lev, iv_model_gdp_2d_lev_gr, iv_model_gdp_2d_gr_gr, adjustbox = TRUE, stage = 1:2, tex = TRUE)
etable(iv_model_gdp_2d_lev_lev_ar1, iv_model_gdp_2d_lev_gr_ar1, iv_model_gdp_2d_gr_gr_ar1, adjustbox = TRUE, stage = 1:2, tex = TRUE)

#etable(iv_model_3d, iv_model_3d_share, stage = 1:2, tex = TRUE)

```

The instrumental variable estimates provide evidence of a robust causal
relationship between national labor demand shocks and local housing
prices. Utilizing a shift-share instrument that interacts national
industry-specific wage growth rates with local industrial exposure
(proxied through property values), the first-stage regression yields a
statistically significant and economically large coefficient. A one-unit
increase in the shift-share measure is associated with a 50.4% increase
in the local House Price Index (p \< 0.01), with an F-statistic of
217.9—well above conventional weak instrument thresholds—confirming
instrument relevance. However, the second-stage results indicate no
statistically significant effect of instrumented housing wealth on local
education expenditure per pupil. Conditional on intergovernmental
revenues and local GDP per capita, the coefficient on log HPI is small
in magnitude and not statistically distinguishable from zero. These
findings suggest that while labor demand shocks substantially affect
local housing markets, they do not necessarily translate into
commensurate changes in public investment in elementary education, at
least in the short-to-medium term.

The lack of a causal link between housing wealth and public education
spending highlights potential institutional rigidities in the fiscal
transmission mechanism linking local economic conditions to public goods
provision. In many U.S. localities, school finance systems remain only
partially responsive to contemporaneous changes in the property tax base
due to funding caps, state-level equalization formulas, or lags in
budget-setting processes. As a result, the benefits of local labor
market booms—reflected in rising property values—may be
disproportionately captured by homeowners and landlords without being
redistributed through improved public services. This creates a
disconnect between private wealth accumulation and collective
investment, with implications for the spatial distribution of
educational opportunity. In areas where property values increase due to
industry-specific wage growth, the absence of corresponding increases in
education spending may exacerbate existing inequalities, especially in
regions with historically lower fiscal capacity or institutional
autonomy.

### State-by-state estimation

```{r, echo = FALSE, fig.height = 8, fig.width = 10}

reg_states <- mines_cz %>% 
  pull(state) %>% 
  unique

mines_cz %>% 
  group_by(state) %>% 
  summarise(n_czs = n_distinct(cz_id)) %>% 
  ggplot() + 
  geom_histogram(aes(x = n_czs), bins = 20, fill = "dodgerblue3", alpha = 0.7) + 
  common_theme +
  labs(x = "Number of commuting zones in state", y = "Number of States",
       title = "Distribution of Number of Commuting Zones per State")
  

dat_temp <- left_join(mines_cz, rename(ss_temp, unit = cz_id), by = c("year","unit")) 
        
state_mods <- list()  
for(k in reg_states){
  mod <- dat_temp %>% 
    filter(state == k) %>%
    feols(as.formula("log_real_Elem_Educ_Total_Exp_pp ~ l1_log_real_Elem_Educ_Total_Exp_pp + log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage + log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment | unit + year"),
    data = ., panel.id = c("unit", "year"))
  state_mods[[as.character(k)]] <- mod
}
  
# Define variable names of interest
vars_of_interest <- c("log_weighted_annual_avg_wkly_wage",
                      "l1_log_weighted_annual_avg_wkly_wage",
                      "l2_log_weighted_annual_avg_wkly_wage")

# Function to extract relevant info from each model
extract_coef_info <- function(model, state_name) {
  # Get summary
  coef_summary <- summary(model)$coeftable
  model_r2 <- r2(model, type = "wr2")
  
  # Filter for relevant variables
  coef_df <- as.data.frame(coef_summary) %>%
    rownames_to_column("variable") %>%
    filter(variable %in% vars_of_interest) %>%
    mutate(
      state = state_name,
      significance = case_when(
        `Pr(>|t|)` < 0.001 ~ "***",
        `Pr(>|t|)` < 0.01  ~ "**",
        `Pr(>|t|)` < 0.05  ~ "*",
        `Pr(>|t|)` < 0.1   ~ ".",
        TRUE               ~ ""
      ),
      r2_within = model_r2) %>%
    select(state, variable, estimate = Estimate, significance, r2_within)
  
  return(coef_df)
}

# Apply to list of models
# Assuming: model_list is named by state names
results_df <- imap_dfr(state_mods, extract_coef_info)
results_df$state_name <- sapply(results_df$state, get_state)

results_df_cleaned <- results_df %>%
  # Add lag labels
  mutate(label = case_when(
    variable == 'log_weighted_annual_avg_wkly_wage' ~ 'l0',
    variable == 'l1_log_weighted_annual_avg_wkly_wage' ~ 'l1',
    variable == 'l2_log_weighted_annual_avg_wkly_wage' ~ 'l2'
  )) %>%
  #filter(significance != "") %>% 
  # Compute total estimate and max R²
  group_by(state_name) %>%
  mutate(
    total_estimate = sum(estimate, na.rm = TRUE),
    r2_within_max = max(r2_within, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  # Order `state` by R²
  mutate(state_name = fct_reorder(state_name, total_estimate, .desc = TRUE)) %>%
  # Create numeric lag order (to control x-axis sorting within each state)
  mutate(
    lag_order = case_when(label == "l0" ~ 1,
                          label == "l1" ~ 2,
                          label == "l2" ~ 3),
    state_label = paste(state_name, label, sep = "_")
  ) %>%
  # Reorder state_label using combined (state, lag_order)
  arrange(state_name, lag_order) %>%
  mutate(state_label = factor(state_label, levels = unique(state_label)))


# Create axis labels: state names at 'l1' only, blanks otherwise
axis_labels <- ifelse(grepl("_l0$", levels(results_df_cleaned$state_label)),
                      gsub("_l0$", "", levels(results_df_cleaned$state_label)),
                      "")

results_df_cleaned %>%   
  # Plot
  ggplot(aes(x = state_label)) +
  geom_point(aes(y = estimate, alpha = significance), shape = "x", size = 3) +
  geom_point(
    data = . %>% filter(label == "l1"),
    aes(y = total_estimate, color = r2_within_max), size = 2
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  common_theme +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) +
  labs(
    title = "Effect of 1% Increase in Wage on Education Expenditure per Pupil",
    x = "States (Each with lags: l0, l1, l2)",
    y = "Coefficient Estimate",
    caption = "Effect of contemporaneous, l1, l2 lag of avg weekly wage on PP Education\n Expenditure controlling for enrollment, GDP, and intergovernmental transfers. \n Each state has 3 lagged wage coefficients (l0, l1, l2).\nBlue dots show total effect across lags, plotted at the l1 position.\n Within R^2 of state-level estimation reflected in color of point).",
    color = "Within R2 of State-level Estimation",
    alpha = "Statistical Significance"
  ) + 
    scale_x_discrete(labels = axis_labels) +
  scale_color_distiller(direction = 1, palette = "Purples") 


covs <- mines_cz %>% 
  select(cz_id, state, log_real_gdp_priv_ind_pc, 
         log_real_gdp_priv_ind, 
         log_real_Elem_Educ_Total_Exp_pp,
         log_Enrollment, share_state, 
         log_weighted_annual_avg_wkly_wage, 
        #gr_weighted_annual_avg_wkly_wage,
         #diff_log_real_gdp_priv_ind, diff_log_real_gdp_priv_ind_pc, 
        log_real_Property_Tax_pp) %>% 
  group_by(cz_id, state) %>% 
  summarise(across(everything(), ~mean(., na.rm = TRUE))) %>% 
  ungroup %>% 
  pivot_longer(!c(cz_id, state), names_to = 'covariate') 

covs$state_name <- sapply(covs$state, get_state)

facet_labels <- c(
  'log_real_gdp_priv_ind_pc' = "(log) Private Industry GDP per capita", 
  # 'diff_log_real_gdp_priv_ind_pc' = '(GR) Private Industry GDP per capita',
  # 'diff_log_real_gdp_priv_ind' = '(GR) Private Industry GDP',
  'log_real_gdp_priv_ind' = "(log) Private Industry GDP",
  'log_weighted_annual_avg_wkly_wage' = '(log) Annual Avg Weekly Wage',
  #'gr_weighted_annual_avg_wkly_wage' = '(GR) Annual Avg Weekly Wage',
  'log_Enrollment' = '(log) Enrollment', 
  'share_state' = 'Share of Exp from State IG Transfers', 
  'log_real_Property_Tax_pp' = "(log) Property Taxes Collected per pupil"
)


covs %>% 
  # use ordering from coefficient estimate above
  mutate(state_name = factor(state_name, levels = levels(results_df_cleaned$state_name))) %>% 
  ggplot(aes(x = state_name, y = value)) +
  geom_violin(trim = FALSE, fill = "magenta", alpha = 0.2, color = NA) + 
  stat_summary(fun = median, geom = "point", position = position_dodge(width = 0.8), 
               shape = 21, size = 1.5, color = "black", fill = "magenta") +
  facet_wrap(~covariate, ncol = 1, scales = "free", labeller = labeller(covariate = facet_labels)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Value of Covariates in Order of State-level Coefficient Estimate",
       subtitle = "Displays the distribution of various covariates by state.",
       y = "Value",
       x = "State (in order of coefficient estimate on wage level as in previous plot)")


```

#### Shift-share

```{r, echo = FALSE, fig.height = 8, fig.width = 10}

state_mods_ss <- list() 
for(k in reg_states){
  mod <- dat_temp %>% 
    filter(state == k) %>%
    feols(as.formula("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment + l1_log_real_Elem_Educ_Total_Exp_pp | unit + state | log_hpi ~ l(gdp_ss_2d,1)"),
    data = ., panel.id = c("unit", "year"))
  state_mods_ss[[as.character(k)]] <- mod
}
  
# Define variable names of interest
vars_of_interest <- c("fit_log_hpi")

# Apply to list of models
# Assuming: model_list is named by state names
results_df <- imap_dfr(state_mods_ss, extract_coef_info)
results_df$state_name <- sapply(results_df$state, get_state)

results_df_cleaned <- results_df %>%
  # Add lag labels
  mutate(label = case_when(
    variable == 'fit_log_hpi' ~ 'HPI Effect'
  )) %>%
  #filter(significance != "") %>% 
  # Compute total estimate and max R²
filter(r2_within > 0) %>% 
  # Order `state` by R²
  mutate(state_name = fct_reorder(state_name, estimate, .desc = TRUE)) %>%
  # Reorder state_label using combined (state, lag_order)
  arrange(state_name) %>%
  mutate(state_label = factor(state_name, levels = unique(state_name)))

# Create axis labels: state names at 'l1' only, blanks otherwise
axis_labels <- levels(results_df_cleaned$state_label)

results_df_cleaned %>%   
  # Plot
  ggplot(aes(x = state_label)) +
# Fill layer (with alpha by significance)
geom_point(
  aes(y = estimate, fill = r2_within, shape = significance),
   size = 4) +
geom_point(aes(y = estimate), shape = 21, size = 4, fill = NA, color = "black", stroke = 0.25, alpha = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  common_theme +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) +
  labs(
    title = "Effect of 1% Increase in SS GDP Instrument on Education Expenditure per Pupil",
    x = "States",
    y = "Coefficient Estimate",
    caption = "Effect of l1 of GDP SS Instrument on PP Education\n Expenditure controlling for enrollment, GDP, and intergovernmental transfers and AR(1) term.\n Within R^2 of state-level estimation reflected in color of point).",
    color = "Within R2 of State-level Estimation",
    alpha = "Statistical Significance"
  ) + 
    scale_x_discrete(labels = axis_labels)  +
  scale_fill_distiller(direction = 1) +
  scale_shape_manual(
  values = c(
    "***" = 21,
    "**"  = 22,
    "*"   = 23,
    "."   = 24#,
    #""    = 25  # non-significant
  )
)


covs <- mines_cz %>% 
  select(cz_id, state, log_real_gdp_priv_ind_pc, 
         log_real_gdp_priv_ind, 
         log_real_Elem_Educ_Total_Exp_pp,
         log_Enrollment, share_state, 
         log_weighted_annual_avg_wkly_wage, 
        #gr_weighted_annual_avg_wkly_wage,
         #diff_log_real_gdp_priv_ind, diff_log_real_gdp_priv_ind_pc, 
        log_real_Property_Tax_pp) %>% 
  group_by(cz_id, state) %>% 
  summarise(across(everything(), ~mean(., na.rm = TRUE))) %>% 
  ungroup %>% 
  pivot_longer(!c(cz_id, state), names_to = 'covariate') 

covs$state_name <- sapply(covs$state, get_state)

facet_labels <- c(
  'log_real_gdp_priv_ind_pc' = "(log) Private Industry GDP per capita", 
  # 'diff_log_real_gdp_priv_ind_pc' = '(GR) Private Industry GDP per capita',
  # 'diff_log_real_gdp_priv_ind' = '(GR) Private Industry GDP',
  'log_real_gdp_priv_ind' = "(log) Private Industry GDP",
  'log_weighted_annual_avg_wkly_wage' = '(log) Annual Avg Weekly Wage',
  #'gr_weighted_annual_avg_wkly_wage' = '(GR) Annual Avg Weekly Wage',
  'log_Enrollment' = '(log) Enrollment', 
  'share_state' = 'Share of Exp from State IG Transfers', 
  'log_real_Property_Tax_pp' = "(log) Property Taxes Collected per pupil"
)


covs %>% 
  # use ordering from coefficient estimate above
  mutate(state_name = factor(state_name, levels = levels(results_df_cleaned$state_name))) %>% 
  ggplot(aes(x = state_name, y = value)) +
  geom_violin(trim = FALSE, fill = "magenta", alpha = 0.2, color = NA) + 
  stat_summary(fun = median, geom = "point", position = position_dodge(width = 0.8), 
               shape = 21, size = 1.5, color = "black", fill = "magenta") +
  facet_wrap(~covariate, ncol = 1, scales = "free", labeller = labeller(covariate = facet_labels)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  labs(title = "Value of Covariates in Order of State-level Coefficient Estimate",
       subtitle = "Displays the distribution of various covariates by state.",
       y = "Value",
       x = "State (in order of coefficient estimate on wage level as in previous plot)")


```


##### Industry by Industry

```{r, echo = FALSE, warning = FALSE, message = FALSE, fig.height = 8, fig.width = 8}

rel_inds <- c("Agriculture, Forestry, Fishing, and Hunting" = "11",
"Mining" = "21",
"Construction" = "23",
"Manufacturing" = "31_33",
"Wholesale Trade" = "42",
"Retail Trade" = "44_45",
"Transportation and Warehousing" = "48_49",
"Utilities" = "22",
"Information" = "51",
"Finance and Insurance" = "52",
"Real Estate and Rental and Leasing" = "53",
"Professional, Scientific, and Technical Services" = "54",
"Management of Companies and Enterprises" = "55",
"Administrative and waste management services" = "56",
"Educational Services" = "61",
"Health Care and Social Assistance" = "62",
"Arts, Entertainment, and Recreation" = "71",
"Accommodation and Food Services" = "72",
"Other Services, except government" = "81")

temp_vars <- ss_temp %>% select(starts_with("lev_gdp_ss") & !contains("2d")) %>% names

# Define variable names of interest
vars_of_interest <- c("fit_log_hpi")

ind_mods <- tibble()
for(k in temp_vars){
    #name <- ind_codes %>% filter(industry_code == k) %>% pull(industry_title)
    #print(name)
    # IV regression using shift-share instrument
    iv_model <- feols(
      as.formula(paste0("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment + l1_log_real_Elem_Educ_Total_Exp_pp | state + year | log_hpi ~ ", k)),
      data = dat_temp, panel.id = c("unit", "year")) %>% 
      extract_coef_info(state_name = "All")
    
    
    state_mods_ind <- list() 
    for(s in reg_states){
      tmp <- dat_temp %>% 
        filter(state == s) 
      
     if(tmp %>% 
      select(state, unit, year, !!k) %>%
        group_by(unit) %>% 
        filter(!all(is.na(get(k)))) %>% n_groups(.) == 1){

        }else{
           mod <- tmp %>%
        feols(as.formula(paste0("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc + log_Enrollment  + l1_log_real_Elem_Educ_Total_Exp_pp | unit + year | log_hpi ~", k)),
        data = ., panel.id = c("unit", "year"))
                 state_mods_ind[[s]] <- mod
        }
    }
    
    ind_mods <- imap_dfr(state_mods_ind, extract_coef_info) %>% 
      rbind(iv_model) %>% 
      mutate(industry = gsub("lev_gdp_ss_", "", k)) %>% 
      rbind(., ind_mods)
}

# Apply to list of models
# Assuming: model_list is named by state names
results_df <- ind_mods %>% 
  mutate(industry_name = names(rel_inds)[match(industry, rel_inds)])

results_df_cleaned <- results_df %>%
  # Add lag labels
  mutate(label = case_when(
    variable == 'fit_log_hpi' ~ 'HPI Effect'
  )) %>%
  group_by(industry_name) %>% 
  mutate(overall_estimate = sum(estimate*(state == "All"), na.rm = TRUE)) %>% 
  ungroup %>% 
  #filter(significance != "") %>% 
  # Compute total estimate and max R²
#filter(r2_within > 0) %>% 
  # Order `state` by R²
  mutate(industry_name = fct_reorder(industry_name, overall_estimate, .desc = TRUE)) %>%
  # Reorder state_label using combined (state, lag_order)
  arrange(industry_name) %>%
  mutate(industry_label = factor(industry_name, levels = unique(industry_name))) %>% 
  filter(r2_within > 0)

plot <- results_df_cleaned %>%   
  # Plot
  ggplot(aes(x = industry_label)) +
# Fill layer (with alpha by significance)
geom_point(data = . %>% filter(state != "All" & significance != ""),
  aes(y = estimate), shape = "x",
   size = 4) +
  geom_point(data = . %>% filter(state == "All"), 
  aes(y = estimate,  color = r2_within,  alpha = significance),
   size = 6) +
# geom_point(aes(y = estimate), shape = 21, size = 4, fill = NA, color = "black", stroke = 0.25, alpha = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  common_theme +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 90, hjust = 1)
  ) +
  labs(
    title = "Effect of 1% Increase in Industry-Specific GDP on Education Expenditure per Pupil",
    x = "Industry",
    y = "Coefficient Estimate",
    caption = "Effect of 1% increase in Industry-Specific GDP (measured as shift-share instrument) on Education\n Expenditure controlling for enrollment, GDP, and intergovernmental transfers, AR(1) term, year and state-level fixed effects.\n Within R^2 of estimation reflected in color of point).",
    color = "Within R2 of State-level Estimation",
    alpha = "Statistical Significance"
  ) + 
    #scale_x_discrete(labels = axis_labels)  +
  scale_color_distiller(direction = 1) +
  scale_shape_manual(
  values = c(
    "***" = 21,
    "**"  = 22,
    "*"   = 23,
    "."   = 24#,
    #""    = 25  # non-significant
  )
) 

plot + 
  scale_color_distiller(direction = 1, palette = "Greens") +
  ylim(c(-1,1))
  

```


### Quantile Regression

```{r, echo = FALSE, fig.height = 8, fig.width = 10}

# Step 1: Remove fixed effects (Canay 2-step)
fe_model <- feols(log_real_Elem_Educ_Total_Exp_pp ~ l1_log_real_Elem_Educ_Total_Exp_pp+
                    log_weighted_annual_avg_wkly_wage +
                                      l1_log_weighted_annual_avg_wkly_wage +
                                      l2_log_weighted_annual_avg_wkly_wage +
                                      log_real_Total_IG_Revenue_pp +
                                      log_real_gdp_priv_ind_pc +
                                      log_Enrollment | unit + year, data = dat_temp) %>% 
  tidy(.) %>% 
  mutate(model = "OLS", 
         tau = "1")

# Step 3: Run quantile regression on residuals at multiple quantiles
taus <- c(0.1, 0.25, 0.5, 0.75, 0.9)
qr_models <- map(taus, ~ rq(log_real_Elem_Educ_Total_Exp_pp ~ l1_log_real_Elem_Educ_Total_Exp_pp+ log_weighted_annual_avg_wkly_wage +
                                      l1_log_weighted_annual_avg_wkly_wage +
                                      l2_log_weighted_annual_avg_wkly_wage +
                                      log_real_Total_IG_Revenue_pp +
                                      log_real_gdp_priv_ind_pc +
                                      log_Enrollment,
                            data = dat_temp, tau = .x))

# Step 4: Tidy results and add quantile info
qr_tidy <- map2_dfr(qr_models, taus, ~ tidy(.x) %>% mutate(tau = as.character(.y), model = "quantile"))

term_levels <- c("(Intercept)", "AR(1)", "L0 Wage", "L1 Wage", "L2 Wage",
                 "Enrollment", "IG Rev pp", "Ind. GDP pc")

# Step 5: Add significance stars and reshape
qr_tidy <- qr_tidy %>%
  rbind(fe_model) %>% 
  mutate(
    p.value = 2 * (1 - pnorm(abs(statistic))),
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      p.value < 0.1   ~ ".",
      TRUE ~ ""
    ),
    # Relabel variables for readability
    term = recode(term,
      "l1_log_real_Elem_Educ_Total_Exp_pp" = "AR(1)",
      "log_real_Total_IG_Revenue_pp" = "IG Rev pp",
      "log_real_gdp_priv_ind_pc" = "Ind. GDP pc",
      "log_Enrollment" = "Enrollment",
      "log_weighted_annual_avg_wkly_wage" = "L0 Wage",
      "l1_log_weighted_annual_avg_wkly_wage" = "L1 Wage",
      "l2_log_weighted_annual_avg_wkly_wage" = "L2 Wage"
    ),
    term = factor(term, levels = term_levels),
    xpos = as.numeric(term),

    # Combine shape and color label
    model_label = case_when(
      model == "OLS" ~ "OLS",
      TRUE ~ paste0("Quantile τ = ", tau)
    )
  )

# Plot with unified legend and updated shape mapping
qr_tidy %>%
  ggplot(aes(x = xpos, y = estimate, color = model_label, shape = model_label, alpha = sig)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbar(
    aes(ymin = estimate - 1.96 * std.error,
        ymax = estimate + 1.96 * std.error),
    height = 0.2,
    position = position_dodge(width = 0.6)
  ) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(
    title = "Coefficient Estimates from OLS and Quantile Regressions",
    x = "Estimate",
    y = NULL,
    color = "Model Type",
    shape = "Model Type"
  ) +
  theme(
    legend.position = "right",
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_line(linetype = "dashed"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_line(color = "slategrey")
  ) +
  scale_shape_manual(values = c(
    "OLS" = 17,  # triangle
    "Quantile τ = 0.1" = 16,
    "Quantile τ = 0.25" = 16,
    "Quantile τ = 0.5" = 16,
    "Quantile τ = 0.75" = 16,
    "Quantile τ = 0.9" = 16
  )) +
  scale_x_continuous(
    breaks = 1:length(term_levels),
    labels = term_levels
  )


```

### Declining vs. Growing Regions

I identify declining vs. growing regions by estimating commuting-zone
growth rates conditional on state and national level growth rates. Using
this distinction (on both a per capita and total gdp bases and a lenient
vs. stringent magnitude threshold), I rerun the key regressions
identified in steps 2 and 3 on the subgroups (declining and growing
regions).

What would be great is to be able to econometrically test when a
commuting zone is "declining." In the first step, it would be good to
identify when a commuting zone is declining overall (GDP, poverty, etc)
but ideally eventually apply this to the education outcome. My hope is
that being able to identify counties that are "declining" we can either
use this variable as a covariate or as a central point of analysis. The
below analysis looks at state-level variables as a first step (mainly to
aid in visual comparison and plotting). Ideally, once a method is
decided on this would be applied to commuting zone-level data which
would need to be summarise/collated in some way for plotting.

```{r, include = FALSE, cache = TRUE}

state_lev <- mines_cz %>%
  group_by(state, year) %>%
  summarise(across(c(real_gdp_priv_ind, real_gdp_total, pop_total), ~sum(., na.rm = TRUE))) %>%
  ungroup %>%
  mutate(real_gdp_priv_ind_pc = real_gdp_priv_ind/pop_total,
         real_gdp_total_pc = real_gdp_total/pop_total,
         across(!c(state, year), ~log(. + 1), .names = "log_{.col}")) %>%
  group_by(state) %>%
  mutate(across(contains("log"), ~.- dplyr::lag(., 1), .names = "diff_{.col}")) %>%
  ungroup

natl_lev <- mines_cz %>%
  group_by(year) %>%
  summarise(across(c(real_gdp_priv_ind, real_gdp_total, pop_total), ~sum(., na.rm = TRUE))) %>%
  ungroup %>%
  mutate(real_gdp_priv_ind_pc = real_gdp_priv_ind/pop_total,
         real_gdp_total_pc = real_gdp_total/pop_total,
         across(!c(year), ~log(. + 1), .names = "log_{.col}"),
         across(contains("log"), ~.- dplyr::lag(., 1), .names = "diff_{.col}"))

growth_rates <- mines_cz %>%
  select(unit, names(state_lev), ) %>%
  left_join(., state_lev, by = c("state", "year"), suffix = c("", "_state")) %>%
  left_join(., natl_lev, by = c("year"), suffix = c("", "_natl"))

```

### CZ GDP growth conditional on state and national level

```{r, echo = FALSE, fig.height = 10, fig.width = 10, warning = FALSE, cache = TRUE}

trend_forms <- c("trend" = "diff_log_real_gdp_priv_ind ~ diff_log_real_gdp_priv_ind_state + diff_log_real_gdp_priv_ind_natl",
                "trend_pc" = "diff_log_real_gdp_priv_ind_pc ~ diff_log_real_gdp_priv_ind_pc_state + diff_log_real_gdp_priv_ind_pc_natl")
                # "trend_pc" = "diff_log_real_gdp_priv_ind_pc ~ diff_log_real_gdp_priv_ind_pc_state")
                 #"trend_pc" = "diff_log_real_gdp_priv_ind_pc ~ diff_log_real_gdp_priv_ind_pc_natl")

for(k in names(trend_forms)){
  # runs model of growth rate in gdp_total and gdp_total_pc conditional on state and national growth rates
   cz_trends = growth_rates %>%
    group_by(unit) %>%
    do(model = lm(as.formula(trend_forms[k]), data = .)) %>%
    ungroup %>%
    mutate(!!k := map_dbl(model, ~ pluck(coef(.x), "(Intercept)")))

   assign(k, cz_trends)
}

cz_trends <- trend %>%
  left_join(., trend_pc, by = "unit")

percentiles <- cz_trends %>%
  summarize(
    p25trend = quantile(trend, 0.25, na.rm = TRUE),
    p75trend = quantile(trend, 0.75, na.rm = TRUE),
    p25trend_pc = quantile(trend_pc, 0.25, na.rm = TRUE),
    p75trend_pc = quantile(trend_pc, 0.75, na.rm = TRUE)
  )

cz_trends <- cz_trends %>%
  mutate(declining = trend < 0,
         declining_extreme = trend < percentiles$p25trend,
         growing_extreme = trend > percentiles$p75trend,
         declining_pc = trend_pc < 0,
         declining_pc_extreme = trend_pc < percentiles$p25trend_pc,
         growing_pc_extreme = trend_pc > percentiles$p75trend_pc)

trend_hist <- cz_trends %>%
  ggplot() +
  geom_histogram(aes(x = trend), bins = 75, fill = "blue", alpha = 0.5) +
  geom_vline(aes(xintercept = percentiles$p25trend),
             color = "darkgreen", linetype = "dashed") +
  geom_vline(aes(xintercept = percentiles$p75trend),
             color = "purple", linetype = "dashed") +
  labs(x = "CZ Trend Coefficient", title = "Distribution of CZ GDP Trend Coefficients", y = "Frequency") +
  annotate("text", x = percentiles$p25trend, y = 90, label = "25th Percentile", color = "darkgreen", angle = 90, vjust = -0.5) +
  annotate("text", x = percentiles$p75trend, y = 90, label = "75th Percentile", color = "purple", angle = 90, vjust = 1.5) +
  theme_minimal()


trend_pc_hist <- cz_trends %>%
  ggplot() +
  geom_histogram(aes(x = trend_pc), bins = 75, fill = "magenta", alpha = 0.5) +
  geom_vline(aes(xintercept = percentiles$p25trend_pc),
             color = "coral", linetype = "dashed") +
  geom_vline(aes(xintercept = percentiles$p75trend_pc),
             color = "slateblue", linetype = "dashed") +
  labs(x = "CZ Trend Coefficient", title = "Distribution of CZ GDP Trend Coefficients (pc)", y = "Frequency") +
  annotate("text", x = percentiles$p25trend_pc, y = 65, label = "25th Percentile", color = "coral", angle = 90, vjust = -0.5) +
  annotate("text", x = percentiles$p75trend_pc, y = 65, label = "75th Percentile", color = "slateblue", angle = 90, vjust = 1.5) +
  theme_minimal()

trend_hist / trend_pc_hist

# Declining is inconsistent in 21.5% of cases
cz_trends %>%
  filter(declining != declining_pc) %>% nrow()/nrow(cz_trends)

# Declining extreme is inconsistent in 15.4% of cases
cz_trends %>%
  filter(declining_extreme != declining_pc_extreme) %>% nrow()/nrow(cz_trends)

# Growing extreme is inconsistent in 11.9% of cases
cz_trends %>%
  filter(growing_extreme != growing_pc_extreme) %>% nrow()/nrow(cz_trends)

theme_update(plot.background = element_rect(fill = NA, color = NA),
             panel.background = element_rect(fill = NA, color = NA),
             panel.border = element_rect(fill = NA, color = NA),
             panel.grid.major.x = element_blank(),
             panel.grid.minor = element_blank(),
             axis.text.x = element_blank(),
             axis.text.y = element_text(size = 10),
             axis.ticks = element_blank(),
             axis.title.y = element_text(size = 13,
                                         margin = margin(r = 10)),
             legend.title = element_text(size = 9),
             plot.margin = margin(10, 25, 10, 25))

cz_trends_plot <- cz_trends %>%
  left_join(., distinct(select(mines_cz, unit, state)), by = "unit")

cz_trends_plot$state = sapply(cz_trends_plot$state, get_state)
cz_trends_plot$region = sapply(cz_trends_plot$state, get_region)

cz_states <- cz_trends_plot %>%
  group_by(state) %>%
  arrange(state, trend_pc) %>%
  mutate(midpt = ifelse(row_number() == round((max(row_number()) - min(row_number()))/2),1, NA)) %>%
  ungroup %>%
  mutate(midpt = row_number() * midpt) %>%
  mutate(unit = factor(unit, levels = unit)) %>%
  ggplot(aes(unit, trend_pc)) +
  geom_segment(aes(x=unit ,xend=unit, y=0, yend=trend_pc, color = state)) +
  geom_point( color="darkblue", size=0.5) +
  geom_label(aes(midpt, 0.15,
                   label = state,
                   color = state),
               fill = NA,
               #family = "Special Elite",
               fontface = "bold",
               label.padding = unit(.2, "lines"),
               label.r = unit(.25, "lines"),
               label.size = .05,
             size = 3) +
  coord_flip() +
  labs(x = "Commuting Zones by State", y = "GDPpc Trend Coefficient") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")

cz_trend <- cz_trends_plot %>%
  arrange(trend_pc) %>%
  mutate(unit = factor(unit, levels = unit)) %>%
  ggplot(aes(unit, trend_pc)) +
  geom_segment(aes(x=unit ,xend=unit, y=0, yend=trend_pc, color = unit)) +
  geom_point( color="darkblue", size=0.5) +
  coord_flip() +
  labs(x = "Commuting Zones", y = "GDPpc Trend Coefficient") +
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none")

cz_regions <- cz_trends_plot %>%
  group_by(region) %>%
  arrange(region, trend_pc) %>%
  mutate(midpt = ifelse(row_number() == round((max(row_number()) - min(row_number()))/2),1, NA)) %>%
  ungroup %>%
  mutate(midpt = row_number() * midpt) %>%
  mutate(unit = factor(unit, levels = unit)) %>%
  ggplot(aes(unit, trend_pc)) +
  geom_segment(aes(x=unit ,xend=unit, y=0, yend=trend_pc, color = region)) +
  geom_point( color="darkblue", size=0.5) +
  geom_label(aes(midpt, 0.08,
                   label = region,
                   color = region),
               fill = NA,
               #family = "Special Elite",
               fontface = "bold",
               label.padding = unit(.2, "lines"),
               label.r = unit(.25, "lines"),
               label.size = .05,
             size = 4) +
  coord_flip() +
  labs(x = "Commuting Zones by Region", y = "GDPpc Trend Coefficient") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        legend.position = "none") +
 scale_color_brewer(palette = "Set2")


cz_regions + cz_states + cz_trend + plot_annotation(title = "Commuting Zone GDP pc Growth Rate Controlling for National and State Level Trends", subtitle = "Calculated as mean of annual growth rate per commuting zone controlling for national and state level growth in GDP pc.",
                                                                   theme = theme(
      plot.title = element_text(size = 18),
      plot.subtitle = element_text(size = 16)
    ))

```

### CZ Wage Growth

```{r, echo = FALSE, fig.height = 10, fig.width = 10, warning = FALSE, cache = TRUE}

wage_trend_forms <- c("wage_growth" = "gr_weighted_annual_avg_wkly_wage ~ gr_natl_annual_avg_wkly_wage_10")
wage_trends <- mines_cz %>%
  select(year, unit, gr_weighted_annual_avg_wkly_wage) %>%
  left_join(., select(natl_rates, year, gr_natl_annual_avg_wkly_wage_10), by = "year")

# runs model of growth rate in gdp_total and gdp_total_pc conditional on state and national growth rates
cz_wage_trends = wage_trends %>%
    group_by(unit) %>%
    do(model = lm(as.formula(wage_trend_forms), data = .)) %>%
    ungroup %>%
    mutate(wage_trend := map_dbl(model, ~ pluck(coef(.x), "(Intercept)")))

percentiles <- cz_wage_trends %>%
  summarize(
    p25trend = quantile(wage_trend, 0.25, na.rm = TRUE),
    p75trend = quantile(wage_trend, 0.75, na.rm = TRUE)
  )

cz_wage_trends <- cz_wage_trends %>%
  mutate(declining = wage_trend < 0,
         declining_extreme = wage_trend < percentiles$p25trend,
         growing_extreme = wage_trend > percentiles$p75trend)

trend_hist <- cz_wage_trends %>%
  ggplot() +
  geom_histogram(aes(x = wage_trend), bins = 75, fill = "blue", alpha = 0.5) +
  geom_vline(aes(xintercept = percentiles$p25trend),
             color = "darkgreen", linetype = "dashed") +
  geom_vline(aes(xintercept = percentiles$p75trend),
             color = "purple", linetype = "dashed") +
  labs(x = "CZ Wage Trend Coefficient", title = "Distribution of CZ Wage Trend Coefficients", y = "Frequency") +
  annotate("text", x = percentiles$p25trend, y = 90, label = "25th Percentile", color = "darkgreen", angle = 90, vjust = -0.5) +
  annotate("text", x = percentiles$p75trend, y = 90, label = "75th Percentile", color = "purple", angle = 90, vjust = 1.5) +
  theme_minimal()

trend_hist

theme_update(plot.background = element_rect(fill = NA),
             panel.background = element_rect(fill = NA, color = NA),
             panel.border = element_rect(fill = NA, color = NA),
             panel.grid.major.x = element_blank(),
             panel.grid.minor = element_blank(),
             axis.text.x = element_blank(),
             axis.text.y = element_text(size = 10),
             axis.ticks = element_blank(),
             axis.title.y = element_text(size = 13,
                                         margin = margin(r = 10)),
             legend.title = element_text(size = 9),
             plot.margin = margin(10, 25, 10, 25))

cz_wage_trends_plot <- cz_wage_trends %>%
  left_join(., distinct(select(mines_cz, unit, state)), by = "unit")

cz_wage_trends_plot$state = sapply(cz_wage_trends_plot$state, get_state)
cz_wage_trends_plot$region = sapply(cz_wage_trends_plot$state, get_region)

cz_wage_states <- cz_wage_trends_plot %>%
  group_by(state) %>%
  arrange(state,wage_trend) %>%
  mutate(midpt = ifelse(row_number() == round((max(row_number()) - min(row_number()))/2),1, NA)) %>%
  ungroup %>%
  mutate(midpt = row_number() * midpt) %>%
  mutate(unit = factor(unit, levels = unit)) %>%
  ggplot(aes(unit, wage_trend)) +
  geom_segment(aes(x=unit ,xend=unit, y=0, yend=wage_trend, color = state)) +
  geom_point( color="darkblue", size=0.5) +
  geom_label(aes(midpt, 0.09,
                   label = state,
                   color = state),
               fill = NA,
               #family = "Special Elite",
               fontface = "bold",
               label.padding = unit(.2, "lines"),
               label.r = unit(.25, "lines"),
               label.size = .05,
             size = 3) +
  coord_flip() +
  labs(x = "Commuting Zones by State", y = "Wage Trend Coefficient") +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 14),
        legend.position = "none")

cz_wage_trend <- cz_wage_trends_plot %>%
  arrange(wage_trend) %>%
  mutate(unit = factor(unit, levels = unit)) %>%
  ggplot(aes(unit, wage_trend)) +
  geom_segment(aes(x=unit ,xend=unit, y=0, yend=wage_trend, color = unit)) +
  geom_point( color="darkblue", size=0.05) +
  coord_flip() +
  labs(x = "Commuting Zones", y = "Wage Trend Coefficient") +
    theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 14),
        legend.position = "none")

cz_wage_regions <- cz_wage_trends_plot %>%
  group_by(region) %>%
  arrange(region, wage_trend) %>%
  mutate(midpt = ifelse(row_number() == round((max(row_number()) - min(row_number()))/2),1, NA)) %>%
  ungroup %>%
  mutate(midpt = row_number() * midpt) %>%
  mutate(unit = factor(unit, levels = unit)) %>%
  ggplot(aes(unit, wage_trend)) +
  geom_segment(aes(x=unit ,xend=unit, y=0, yend=wage_trend, color = region)) +
  geom_point( color="darkblue", size=0.5) +
  geom_label(aes(midpt, 0.05,
                   label = region,
                   color = region),
               fill = NA,
               #family = "Special Elite",
               fontface = "bold",
               label.padding = unit(.2, "lines"),
               label.size = .09,
             size = 4) +
  coord_flip() +
  labs(x = "Commuting Zones by Region", y = "Wage Trend Coefficient") +
  theme(axis.text.y=element_blank(),
        #axis.ticks.y=element_blank(),
        axis.text.x = element_text(size = 14),
        legend.position = "none") +
 scale_color_brewer(palette = "Set2")


cz_wage_regions + cz_wage_states + cz_wage_trend + plot_annotation(title = "Commuting Zone Wage Growth Rate Controlling for National and State Level Trends", subtitle = "Calculated as mean of annual growth rate per commuting zone controlling for national and state level wage growth.",
                                                                   theme = theme(
      plot.title = element_text(size = 18),
      plot.subtitle = element_text(size = 16)
    ))


cz_trends %>% left_join(cz_wage_trends, by = "unit") %>% ggplot(., aes(x = trend, y = wage_trend)) +
     geom_point(alpha = 0.2) +
     geom_smooth(method = "lm", se = TRUE, color = "maroon", fill = "magenta") +
     labs(
         x = "Real GDP Trend",
         y = "Wage Trend",
         title = "Relationship Between GDP and Wage Trends (per CZ)"
     ) +
     theme_minimal() +
  theme(axis.text.x = element_text(size = 14),
        axis.text.y = element_text(size = 14),
        title = element_text(size = 18))


map_data <- mines_cz %>% 
  select(year, unit, log_weighted_annual_avg_wkly_wage, gr_weighted_annual_avg_wkly_wage) %>% 
  group_by(unit) %>% 
  mutate(gr_weighted_annual_avg_wkly_wage = mean(gr_weighted_annual_avg_wkly_wage, na.rm = TRUE)) %>%
  filter(year == 2021) %>% 
  select(unit, year, log_weighted_annual_avg_wkly_wage, gr_weighted_annual_avg_wkly_wage) %>% 
  rename(cz_id = unit)

cz_wage_trends %>% rename(cz_id = unit) %>% left_join(czs,., by = "cz_id") %>% 
  mutate(corrected_trend = 1-wage_trend) %>% 
  plot_usmap(data = ., values = "corrected_trend", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
  scale_fill_gradient2(midpoint = 1, na.value = "palegreen", low = "mediumvioletred", high = "midnightblue") +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
  theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold")) +
  labs(title = "Growth Rate of Weekly Wage in Relation to Median (CZ)",
       fill = "Growth Rate in Weekly Wage (2001-2021)") 
  

map_gr_wage <- gen_czs_years(map_data) %>% 
  left_join(., map_data, by = "cz_id") %>% 
  plot_usmap(data = ., values = "gr_weighted_annual_avg_wkly_wage", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
  scale_fill_gradient2(midpoint = 0.03, na.value = "palegreen", low = "mediumvioletred", high = "midnightblue") +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
  theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold")) +
  labs(title = "Growth Rate of Weekly Wage in Relation to Median (CZ)",
       fill = "Growth Rate in Weekly Wage (2001-2021)") 

map_lev_wage <- gen_czs_years(map_data) %>% 
  left_join(., map_data, by = "cz_id") %>% 
  plot_usmap(data = ., values = "log_weighted_annual_avg_wkly_wage", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
  scale_fill_gradient2(midpoint = 6.8, na.value = "palegreen", low = "mediumvioletred", high = "midnightblue" )+
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
  theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold")) +
  labs(title = "Weekly Wage Level (2021) in Relation to Median (CZ)",
       fill = "(log) Annual Average Weekly Wage") 

map_gr_wage
map_lev_wage

```

```{r, echo = FALSE, cache = TRUE}

mines_cz_decl <- cz_trends %>%
  select(unit, contains("trend"), contains("growing"), contains("declining")) %>%
  left_join(mines_cz, ., by = "unit")

mines_cz_wage_decl <- cz_wage_trends %>% 
  select(unit, contains('trend'), contains('growing'), contains('declining')) %>% 
  left_join(mines_cz, ., by = "unit")

# print("Declining")
# run_model(selected_iv_models, filter(mines_cz_decl, declining_pc)) %>% etable(headers = c("INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod","INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod"))
# print("Growing")
# run_model(selected_iv_models, filter(mines_cz_decl, !declining_pc)) %>% etable(headers = c("INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod","INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod"))
#
# print("Extreme Decline (25th percentile of growth rates)")
# run_model(selected_iv_models, filter(mines_cz_decl, declining_pc_extreme)) %>% etable(headers = c("INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod","INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod"))
#
# print("Extreme Growth (75th percentile of growth rates)")
# run_model(selected_iv_models, filter(mines_cz_decl, growing_pc_extreme)) %>% etable(headers = c("INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod","INSTR: Total Active Prod", "INSTR: (l1) Total Active Prod"), title = "Hello")

```

### Running base models on declining vs. growing sub-groups

```{r, echo = FALSE, results = 'asis'}

form <- "log_real_Elem_Educ_Total_Exp_pp ~ sw(log_real_gdp_priv_ind_pc + l1_log_real_gdp_priv_ind_pc + l2_log_real_gdp_priv_ind_pc, log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage, log_hpi + l1_log_hpi + l2_log_hpi + l3_log_hpi + l4_log_hpi) + log_real_Total_IG_Revenue_pp | unit + year"

gr_form <- "diff_log_real_Elem_Educ_Total_Exp_pp ~ sw(diff_log_real_gdp_priv_ind_pc + l1_diff_log_real_gdp_priv_ind_pc + l2_diff_log_real_gdp_priv_ind_pc, gr_weighted_annual_avg_wkly_wage + l1_gr_weighted_annual_avg_wkly_wage + l2_gr_weighted_annual_avg_wkly_wage, gr_hpi + l1_gr_hpi + l2_gr_hpi + l3_gr_hpi + l4_gr_hpi) + diff_log_real_Total_IG_Revenue_pp | unit + year"

prop_vals_form <- "log_hpi ~ l(log_weighted_annual_avg_wkly_wage, 0:7) | unit + year"
gr_prop_vals_form <- "gr_hpi ~ l(gr_weighted_annual_avg_wkly_wage, 0:7) | unit + year"

for(k in list(form, gr_form, prop_vals_form, gr_prop_vals_form)){
  run_model(k, filter(mines_cz_decl, declining_pc)) %>% etable(tex = TRUE, title = "Declining", adjustbox = TRUE) %>% print(.)
  run_model(k, filter(mines_cz_decl, !declining_pc)) %>% etable(tex = TRUE, title = "Growing", adjustbox = TRUE) %>% print(.)
  
  run_model(k, filter(mines_cz_decl, declining_pc_extreme)) %>% etable(tex = TRUE, title = "Hyper-Declining", adjustbox = TRUE) %>% print(.)
  run_model(k, filter(mines_cz_decl, growing_pc_extreme)) %>% etable(tex = TRUE, title = "Hyper-Growing", adjustbox = TRUE) %>% print(.)
}

```

```{r, echo = FALSE}

decl <- filter(mines_cz_decl, declining_pc)
grow <- filter(mines_cz_decl, !declining_pc)

decl_extr <- filter(mines_cz_decl, declining_pc_extreme)
grow_extr <- filter(mines_cz_decl, growing_pc_extreme)

dfs <- list("Declining" = decl, "Hyper-Declining" = decl_extr, "Growing" = grow, "Hyper-Growing" = grow_extr)


decl_wage <- filter(mines_cz_wage_decl, declining)
grow_wage <- filter(mines_cz_wage_decl, !declining)

decl_extr_wage <- filter(mines_cz_wage_decl, declining_extreme)
grow_extr_wage <- filter(mines_cz_wage_decl, growing_extreme)

dfs_wage <- list("Declining" = decl_wage, "Hyper-Declining" = decl_extr_wage, "Growing" = grow_wage, "Hyper-Growing" = grow_extr_wage)


```

```{r, echo = FALSE}

mods <- list()
for(ss in c("lev_ss_2d")){
  for(df in names(dfs)){
    # IV regression using shift-share instrument
    iv_model <- feols(
   as.formula(paste0("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc | unit + year | log_hpi ~ ", ss)),
   data = left_join(dfs[[df]], rename(ss_temp, unit = cz_id), by = c("year", "unit"))
    )
    mods[[ss]][[df]] <- iv_model
  }
}

mods_wage <- list()
for(ss in c("lev_ss_2d")){
  for(df_wage in names(dfs_wage)){
    # IV regression using shift-share instrument
    iv_model <- feols(
   as.formula(paste0("log_real_Elem_Educ_Total_Exp_pp ~ log_real_Total_IG_Revenue_pp + log_real_gdp_priv_ind_pc | unit + year | log_hpi ~ ", ss)),
   data = left_join(dfs_wage[[df_wage]], rename(ss_temp, unit = cz_id), by = c("year", "unit"))
    )
    mods_wage[[ss]][[df_wage]] <- iv_model
  }
}

```

### Running IV models on declining vs. growing sub-groups

The following implements an employment based Bartik instrument for
various industries available from the Quarterly Census of Employment and
Wages.

```{r, echo = FALSE, results = 'asis'}

for(k in names(mods)){
  etable(mods[[k]], title = k, tex = TRUE, stage = 1:2, headers = rep(names(mods[[k]]), each = 2), adjustbox = TRUE) %>% print(.)
}

for(k in names(mods_wage)){
  etable(mods_wage[[k]], title = k, tex = TRUE, stage = 1:2, headers = rep(names(mods_wage[[k]]), each = 2), adjustbox = TRUE) %>% print(.)
}

```

```{r appendix_ss, echo = FALSE, include = FALSE, results = 'asis', cache = TRUE}

append_to_appendix("## Industry Specific Shift-Share Instrument",
                   "Industry-level shift-share_instrument")

# compute_ss takes the source data "QCEW" or "QWI", a base year for the Bartik, whether to compute at CZ or fips level. This si then passed to plot_ss() which returns the instrumetn for a specific industry code and a plot.
ss_temp <- compute_ss_old(source = "QCEW", base_year = 2004, unit_id = "cz_id")


ind_codes %>% filter(nchar(industry_code) == 2 & industry_code != 99)  %>% select(-keep, -industry_code) %>% xtable() %>% print(include.rownames = FALSE)

ind_mods <- list()
inds <- ind_codes %>% filter(nchar(industry_code) == 2 & industry_code != 99) %>% pull(industry_code) %>% unique
for(k in inds){
    ss_inst <- ss_temp %>%
      plot_ss_old(., k, "gr_natl_annual_avg_wkly_wage_", "cz_id", ind_codes)
    name <- ind_codes %>% filter(industry_code == k) %>% pull(industry_title)

    mods <- list()
    for(df in names(dfs)){
      # IV regression using shift-share instrument
      iv_model <- feols(
        as.formula(paste0("log_real_Elem_Educ_Total_Exp_pp ~ log_real_gdp_priv_ind_pc + log_real_Total_IG_Revenue_pp | unit + year | log_hpi ~ ", paste0("ss_", k))),
        data = left_join(dfs[[df]], rename(ss_inst, unit = cz_id), by = c("year", "unit"))
      )
      mods[[df]] <- iv_model
    }
    ind_mods[[k]] <- mods
}

for(k in names(ind_mods)){
  etable(ind_mods[[k]], title = k, tex = TRUE, headers = names(ind_mods[[k]]),  adjustbox = TRUE) %>% print()
}

# Filter the list of models for significant log_real_gdp_priv_ind_pc coefficients
significant_mods <- mods[sapply(mods, function(m) {
  coefs <- summary(m)$coeftable
  if ("fit_log_hpi" %in% rownames(coefs)) {
    return(coefs["fit_log_hpi", "Pr(>|t|)"] < 0.05)
  } else {
    return(FALSE)
  }
})]

```

### Removing outliers - really high-income commuting zones!

As you can see in the scatterplot below, there is a somewhat non-linear
relationship between property taxes and elementary expenditure as
property taxes collected rise. This happens largely as a result of very
high-income commuting zones. Therefore, I exclude any commuting zone
that spends more than 28k per pupil to avoid any distorting effects.
This removes 12 counties (\~2% of the sample) This could benefit from
more robust outlier detection. This outlier exclusion weakens our
results (and the validity of our instrument choice) in the
production-based IV regression. Worth noting and thinking about!!

```{r, echo = FALSE, cache = TRUE}

mines_no_outliers <- mines_cz %>%
  group_by(unit) %>%
  filter(!any(log_real_Elem_Educ_Total_Exp_pp > 10.25)) %>%
  ungroup

mines_cz %>%
  ggplot(aes(x = log_real_Elem_Educ_Total_Exp_pp, y = log_real_Property_Tax_pp, colour = share_own_discrete)) +
  geom_jitter() +
    #geom_smooth(method=lm, colour = "black", linetype = "dashed") +
  labs(title = "Elem Education Expenditure pp vs Property Tax pp") +
  geom_abline(slope = 1, intercept = 0, linetype = 'dashed') +
  geom_vline(xintercept = 10.25, linetype = 'dashed') +
  theme_bw() +
  theme(title = element_text(size = 9)) +
  labs(colour = "Reliance on Local Sources")

# run_model(iv_prod_tax, mines_no_outliers) %>% etable(tex = TRUE)
# run_model(iv_prod_tax, mines_cz) %>% etable(tex = TRUE)
# run_model(iv_coaln_tax, mines_no_outliers) %>% etable(tex = TRUE)

```

## Panel VAR Specification

$$
Y_{it} = \alpha_i +  \sum_{k = 1}^{4} \gamma_{k}A_{i,t-k} + \beta X_{it} + \varepsilon_{it}
$$

Where we approach a level and per capita value expression of the
relationship between total educaiton expenditure, intergovernmental
revenue, house prices conditioned on GDP and wage levels.

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Total Educ. Exp.})_{it} \\
\log(\text{real Total IG Revenue})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP})_{it} \\
\log(\text{wage})_{it}
\end{bmatrix}
$$

-   $A_1, A_2, A_3, A_4$ are $3 \times 3$ coefficient matrices
-   $\beta$ is a $3 \times 2$ matrix of coefficients on the exogenous
    variables
-   $\alpha_i$ is a vector of unit fixed effects
-   $\varepsilon_{it}$ is the error term

Where

$$
Y_{it} =
\begin{bmatrix}
\log(\text{real Own Source Rev. per person})_{it} \\
\log(\text{real IG Revenue per person})_{it} \\
\log(\text{wage})_{it} \\
\log(\text{HPI})_{it}
\end{bmatrix},
\quad
X_{it} =
\begin{bmatrix}
\log(\text{real GDP per capita})_{it}
\end{bmatrix}
$$

-   $A_1, A_2, A_3, A_4$ are $4 \times 4$ coefficient matrices
-   $B$ is a $4 \times 1$ coefficient matrix
-   $\alpha_i$ unit fixed effects
-   $\varepsilon_{it}$ error term

```{r, echo = FALSE, cache = TRUE}

pvar_data <- mines_cz %>%
  #select(unit, year, contains("real_Total_Educ_Total_Exp"), contains("real_gdp_total")) %>%
  drop_na() %>%
  data.frame()

# Estimate PVAR with 1 lag
pvar_model_levels <- pvarfeols(
  dependent_vars = c("log_real_Total_Educ_Total_Exp",
                     #"log_real_Total_Rev_Own_Sources",
                     #"log_real_gdp_priv_ind",
                     #"log_real_Property_Tax",
                     "log_real_Total_IG_Revenue", #"log_real_Total_Fed_IG_Revenue_pp",
                     #"log_weighted_annual_avg_wkly_wage",
                     "log_hpi"),
  exog_vars = c("log_real_gdp_priv_ind", "log_weighted_annual_avg_wkly_wage"), #"share_own",
  lags = 4,
  #transformation = "demean",   # alternative: "firstdiff"
  data = pvar_data,
  panel_identifier = c("unit", "year")
)

pvar_model_levels_pc <- pvarfeols(
  dependent_vars = c(#"log_real_Total_Educ_Total_Exp_pp", #"log_real_gdp_priv_ind_pc",
                     "log_real_Total_Rev_Own_Sources_pp",
                     #"log_real_Property_Tax_pp",
                     "log_real_Total_IG_Revenue_pp",
                     "log_weighted_annual_avg_wkly_wage", #"log_real_Total_State_IG_Revenue_pp", "log_real_Total_Fed_IG_Revenue_pp",
                     "log_hpi"),
  exog_vars = c("log_real_gdp_priv_ind_pc"), #"share_own",
  lags = 4,
  #transformation = "demean",   # alternative: "firstdiff"
  data = pvar_data,
  panel_identifier = c("unit", "year")
)

for(mod in list(pvar_model_levels, pvar_model_levels_pc)){
  #summary(mod) %>% print

  # Impulse response function
  irf_result <- mod %>% girf(n.ahead = 10, ma_approx_steps= 10)

  # irf_result_bootstrapped <- bootstrap_irf(mod, typeof_irf = c("GIRF"),
  #           n.ahead = 10,
  #           nof_Nstar_draws = 3,
  #           confidence.band = 0.95,
  #           mc.cores = 5)
  # Plot IRFs
  plot(irf_result, scales = "free") %>% print()
}


# # Get shock variables (names of shock sources)
# shock_vars <- names(irf_result)
# irf_df <- map_dfr(shock_vars, function(shock) {
#   lower_matrix <- irf_result_bootstrapped$Lower[[shock]]
#   upper_matrix <- irf_result_bootstrapped$Upper[[shock]]
#
#   # Calculate mean pointwise IRF from bootstrap draws
#   mean_matrix <- (lower_matrix + upper_matrix) / 2
#
#   expand.grid(
#     horizon = seq_len(nrow(mean_matrix)),
#     response = colnames(mean_matrix)
#   ) %>%
#     mutate(
#       shock = shock,
#       irf = as.vector(mean_matrix),
#       lower = as.vector(lower_matrix),
#       upper = as.vector(upper_matrix)
#     )
# })
# ggplot(irf_df, aes(x = horizon, y = irf)) +
#   geom_line(color = "steelblue") +
#   geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.2, fill = "steelblue") +
#   facet_grid(response ~ shock, scales = "free_y") +
#   labs(
#     title = "Impulse Response Functions with 95% Confidence Intervals",
#     x = "Horizon",
#     y = "Response"
#   ) +
#   theme_minimal()

```

# Property Prices

```{r, echo = FALSE, results = 'asis', cache = TRUE}
# is <- isatpanel(
#       data = data.frame(mines_cz),
#       formula = as.formula("log_hpi ~ log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage"),
#       index = c("unit", "year"),
#       effect = "twoways",
#       #iis = TRUE,
#       fesis = TRUE,
#       #ar = 1,
#       t.pval = 0.01,
#       max.block.size = 5
#     )

fun_list <- c(log_hpi ~ log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage + log_real_gdp_priv_ind | unit + year,

gr_hpi ~ gr_weighted_annual_avg_wkly_wage + l1_gr_weighted_annual_avg_wkly_wage + l2_gr_weighted_annual_avg_wkly_wage | unit + year,

log_real_Elem_Educ_Total_Exp ~ log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage + log_real_gdp_priv_ind + l1_log_real_gdp_priv_ind + l2_log_real_gdp_priv_ind | unit + year,

diff_log_real_Elem_Educ_Total_Exp ~ gr_weighted_annual_avg_wkly_wage + l1_gr_weighted_annual_avg_wkly_wage + l2_gr_weighted_annual_avg_wkly_wage + diff_log_real_gdp_priv_ind + l1_diff_log_real_gdp_priv_ind + l2_diff_log_real_gdp_priv_ind | unit + year,

log_real_Elem_Educ_Total_Exp_pp ~ log_weighted_annual_avg_wkly_wage + l1_log_weighted_annual_avg_wkly_wage + l2_log_weighted_annual_avg_wkly_wage + log_real_gdp_priv_ind_pc + l1_log_real_gdp_priv_ind_pc + l2_log_real_gdp_priv_ind_pc | unit + year,

diff_log_real_Elem_Educ_Total_Exp_pp ~ gr_weighted_annual_avg_wkly_wage + l1_gr_weighted_annual_avg_wkly_wage + l2_gr_weighted_annual_avg_wkly_wage + diff_log_real_gdp_priv_ind_pc + l1_diff_log_real_gdp_priv_ind_pc + l2_diff_log_real_gdp_priv_ind_pc | unit + year)

run_model(fun_list, mines_no_outliers) %>% etable(tex = TRUE, adjustbox = TRUE)

```

```{r, echo = FALSE, results = 'asis', cache = TRUE}
# General-to-Specific model selection using fixest
gets_fixest <- function(dep_var, indep_vars, data, fe = NULL, pval_cutoff = 0.1) {
  remaining_vars <- indep_vars

  repeat {
    # Build formula
    rhs <- paste(remaining_vars, collapse = " + ")
    formula_str <- as.formula(paste(dep_var, "~", rhs, if (!is.null(fe)) paste("|", fe) else ""))

    # Estimate model
    model <- feols(formula_str, data = data)
    coefs <- summary(model)$coeftable

    # Filter out only the coefficients from the explanatory variables
    pvals <- coefs[, "Pr(>|t|)"]
    pvals <- pvals[names(pvals) %in% remaining_vars]

    # Stop if all variables are below the cutoff or no variables remain
    if (length(pvals) == 0 || all(pvals < pval_cutoff)) break

    # Remove the variable with the highest p-value
    worst_var <- names(which.max(pvals))
    remaining_vars <- setdiff(remaining_vars, worst_var)
  }

  # Return the final reduced model
  final_formula <- paste(dep_var, "~", paste(remaining_vars, collapse = " + "), if (!is.null(fe)) paste("|", fe) else "")
  return(final_formula)
}

# Example usage
# Replace 'your_data' with your actual data frame name
final_model <- gets_fixest(
  dep_var = "log_real_Elem_Educ_Total_Exp_pp",
  indep_vars = c("log_weighted_annual_avg_wkly_wage",
                 "l1_log_weighted_annual_avg_wkly_wage",
                 "l2_log_weighted_annual_avg_wkly_wage",
                 "log_real_gdp_priv_ind_pc",
                 "l1_log_real_gdp_priv_ind_pc",
                 "l2_log_real_gdp_priv_ind_pc",
                 "log_real_Property_Tax_pp",
                 #"log_real_Total_State_IG_Revenue_pp",
                 #"log_real_Total_Fed_IG_Revenue_pp",
                 "log_hpi"),  # replace with your actual variable names
  data = mines_cz,
  fe = "unit + year",        # optional fixed effects
  pval_cutoff = 0.1         # significance threshold
)

# View results
# At a 10% significance level, we retain a 2-timeperiod lag of real_gdp_pc, contemporaneous wage (not likely), property_tax_pp, state_revenue_pp, hpi
feols(as.formula(final_model), data = mines_cz) %>% etable(adjustbox = TRUE, tex = TRUE)


# I want to allow the coefficient to vary on the weekly wage treatment variable
feols(as.formula(gsub("log_weighted_annual_avg_wkly_wage", "i(share_own_discrete, log_weighted_annual_avg_wkly_wage)", final_model)), data = mines_cz) %>% 
  etable(adjustbox = TRUE, tex = TRUE)

```

```{r, cache = TRUE}

test <- readRDS(here("output/prop_price_breaks_csis.RDS"))

```

# Results

# Discussion

# Conclusion

The determinants of inequality in public education delivery in the US
are multiple and complex. Significant evidence exists of the role of
historically discriminatory policies related to congressional
districting, under-investment in low-income areas of color. Though this
work does not directly inform this debate, further work could explore
the extent to which wage growth interacts with such structural policies.

```{r, child = "appendix_generated.qmd"}

```
