---
title: "Wealth"
author: "Ebba Mark"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Wealth Data
Source: https://www.openicpsr.org/openicpsr/project/192306/version/V4/view?path=/openicpsr/192306/fcr:versions/V4/cz_wealth_inequality.csv&type=file
Citation: Suss, Joel, Connor, Dylan, and Kemeny, Tom. GEOWEALTH-US: Spatial wealth inequality data for the United States, 1960-2020: cz_wealth_inequality.csv. Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributor], 2024-02-22. https://doi.org/10.3886/E192306V4-156741

```{r cars}

wealth_cz <- read.csv(here("data/raw/geo_wealth_inequality/cz_wealth_inequality.csv")) %>% tibble  %>% mutate(cz_id_1990 = as.character(czone))
wealth_puma <- read.csv(here("data/raw/geo_wealth_inequality/puma_wealth_inequality.csv")) %>% tibble 
cz_1990_trends_dat <- readRDS(here('data/temp/cz_1990_trends_dat.rds')) %>% tibble

wealth_cz %>% pull(cz_id_1990) %>% unique
cz_1990_trends_dat %>% pull(cz_id_1990) %>% unique

setdiff(unique(cz_1990_trends_dat$cz_id_1990), unique(wealth_cz$cz_id_1990))
setdiff(unique(wealth_cz$cz_id_1990), unique(cz_1990_trends_dat$cz_id_1990))

cz_1990_trends_dat %>% left_join(., wealth_cz, by = c("year", 'cz_id_1990')) 

```


```{r pressure, echo=FALSE}

wealth_cz %>% 
  ggplot(aes(x = year, y = wealth_median, group = cz_id_1990)) +
  geom_line()

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
