---
title: "Measuring Decline"
format: 
  html:
    embed-resources: true
    theme: cosmo
    toc: true
    toc_float: true
    toc-expand: 3
    toc-depth: 4
    toc-location: left
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r setup, warning = FALSE, include = FALSE}

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```

# Introduction

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(sf)
library(stargazer)
library(zoo)
library(tidycensus)
library(here)
library(viridis)
library(datasets)
library(broom)
library(fixest)
library(readxl)
library(plm)
library(getspanel)
library(gridExtra)
library(kableExtra)
library(lattice)
library(lubridate)
library(stringr)
library(xts)
library(urca)
library(dynlm)
library(usmap)
library(tseries)
library(quantmod)
library(lmtest)
library(sandwich)
library(vars)
library(ggridges)
library(grid)
library(ggplotify)
library(lubridate)
library(conflicted)
conflict_prefer_all("dplyr", quiet = TRUE)
source(here('code/source_code/useful_functions.R'))
source(here('code/source_code/dicts.R'))
source(here('code/source_code/cz_cleaning.R'))

```

# Data

Data on Property Taxes, Elementary Education Expenditure, Private Industry and Total GDP at the county, commuting zone, state, and national levels.

```{r, echo = FALSE, message = FALSE, cache = TRUE}

mines_df <- readRDS(here("data/out/regression_data_complete_fips.RDS"))

```

```{r}
# Dataframe to easily match state names and CZ ids if necessary}
# To be resolved: when a CZ is present in multiple states
code_match <- mines_df %>% 
  select(fips, state_name, cz_id, cz_id_1990) %>% 
  distinct

# Relevant vars for trend analysis below
# Should be central place to add more vars if necessary
vars_grab <- c("real_gdp_priv_ind",
               "real_gdp_total",
               "real_Elem_Educ_Total_Exp", 
               "real_Total_Educ_Total_Exp",
               "real_Property_Tax", 
               "Enrollment", 
               "pop_total")

vars_grab_restr <- c("real_gdp_priv_ind",
                     "real_gdp_priv_ind_pc",
               "real_gdp_total",
               "real_gdp_total_pc",
               "real_Elem_Educ_Total_Exp",
               "real_Elem_Educ_Total_Exp_pp",
               "real_Total_Educ_Total_Exp",
               "real_Total_Educ_Total_Exp_pp",
               "real_Property_Tax", 
               "real_Property_Tax_pp", 
               "Enrollment", 
               "pop_total")

mines_restr <- mines_df %>% 
  group_by(fips) %>% 
  arrange(fips, year) %>% 
  mutate(across(c(vars_grab_restr, paste0("log_", vars_grab_restr)), ~. - dplyr::lag(., 1), .names = "diff_{.col}")) %>% 
  group_by(year) %>% 
  mutate(natl_pop = sum(pop_total, na.rm = TRUE)) %>% 
  ungroup
  
## Summarise to Commuting Zone Level
mines_cz <- mines_restr %>%
  select(cz_id, year, fips, vars_grab) %>% 
  group_by(cz_id, year) %>%
  summarise(across(vars_grab, ~sum(., na.rm = TRUE)),
            n_fips = n_distinct(fips)) %>% 
  ungroup %>% 
  mutate(real_Elem_Educ_Total_Exp_pp = real_Elem_Educ_Total_Exp/Enrollment,
         real_Total_Educ_Total_Exp_pp = real_Total_Educ_Total_Exp/Enrollment,
         real_gdp_priv_ind_pc = real_gdp_priv_ind/pop_total,
         real_gdp_total_pc = real_gdp_total/pop_total,
         real_Property_Tax_pp = real_Property_Tax/Enrollment,
         across(!c(cz_id, year, n_fips), ~log(.), .names = "log_{.col}"),
         across(!c(cz_id, year, n_fips), ~. - dplyr::lag(., 1), .names = "diff_{.col}"),
         time = year - 2000)

## Summarise to Commuting Zone Level
mines_cz_1990 <- mines_restr %>%
  select(cz_id_1990, year, fips, vars_grab) %>% 
  group_by(cz_id_1990, year) %>%
  summarise(across(vars_grab, ~sum(., na.rm = TRUE)),
            n_fips = n_distinct(fips)) %>% 
  ungroup %>% 
  mutate(real_Elem_Educ_Total_Exp_pp = real_Elem_Educ_Total_Exp/Enrollment,
         real_Total_Educ_Total_Exp_pp = real_Total_Educ_Total_Exp/Enrollment,
         real_gdp_priv_ind_pc = real_gdp_priv_ind/pop_total,
         real_gdp_total_pc = real_gdp_total/pop_total,
         real_Property_Tax_pp = real_Property_Tax/Enrollment,
         across(!c(cz_id_1990, year, n_fips), ~log(.), .names = "log_{.col}"),
         across(!c(cz_id_1990, year, n_fips), ~. - dplyr::lag(., 1), .names = "diff_{.col}"),
         time = year - 2000)

## Summarise to State Level
mines_state <- mines_restr %>%
  select(state_name, fips, cz_id, year, fips, vars_grab) %>% 
  group_by(state_name, year) %>%
  summarise(across(vars_grab, ~sum(., na.rm = TRUE)),
            n_fips = n_distinct(fips),
            n_cz = n_distinct(cz_id)) %>% 
  ungroup %>% 
  mutate(real_Elem_Educ_Total_Exp_pp = real_Elem_Educ_Total_Exp/Enrollment,
         real_Total_Educ_Total_Exp_pp = real_Total_Educ_Total_Exp/Enrollment,
         real_gdp_priv_ind_pc = real_gdp_priv_ind/pop_total,
         real_gdp_total_pc = real_gdp_total/pop_total,
         real_Property_Tax_pp = real_Property_Tax/Enrollment,
         across(!c(state_name, year, n_fips, n_cz), ~log(.), .names = "log_{.col}"),
         across(!c(state_name, year, n_fips, n_cz), ~. - dplyr::lag(., 1), .names = "diff_{.col}"),
         time = year - 2000)
      
## Summarise to natl level
mines_natl <- mines_restr %>%
  select(state_name, fips, cz_id, year, fips, vars_grab) %>% 
  group_by(year) %>%
  summarise(across(vars_grab, ~sum(., na.rm = TRUE)),
            n_fips = n_distinct(fips),
            n_cz = n_distinct(cz_id),
            n_state = n_distinct(state_name)) %>% 
  ungroup %>% 
  mutate(real_Elem_Educ_Total_Exp_pp = real_Elem_Educ_Total_Exp/Enrollment,
         real_Total_Educ_Total_Exp_pp = real_Total_Educ_Total_Exp/Enrollment,
         real_gdp_priv_ind_pc = real_gdp_priv_ind/pop_total,
         real_gdp_total_pc = real_gdp_total/pop_total,
         real_Property_Tax_pp = real_Property_Tax/Enrollment,
         across(!c(year, n_fips, n_cz, n_state), ~log(.), .names = "log_{.col}"),
         across(!c(year, n_fips, n_cz, n_state,), ~. - dplyr::lag(., 1), .names = "diff_{.col}"),
         time = year - 2000,
         id = "USA")


```

```{r, echo = FALSE, cache = TRUE}
# Data Transformation Function

# Pass a dataframe with distinct ids (ie. state_name, CZ, or fips)
# Dataframe should only have variables that should be transformed in line with all calls in the following function

data_transform <- function(df, vars, unit_id, time_id){
  df %>% 
    select(c(unit_id, time_id, vars)) %>%
    mutate(across(!c(unit_id, time_id),
                  list(t0 = ~(ifelse(!!sym(time_id) == 2001, 1, NA))*.,
                       t2 = ~(ifelse(!!sym(time_id) == 2003, 1, NA))*.),
                  .names = "{.fn}_{.col}")) %>% 
    group_by(!!sym(unit_id)) %>% 
    fill(matches("^t[[:digit:]]"),
      .direction = "down") %>% 
    mutate(across(!c(!!sym(time_id), matches("^t[[:digit:]]")), list(
                                                           ma3 = ~rollmean(., 3, align = 'right', fill = NA)),
                  .names = "{.fn}_{.col}")) %>% 
    ungroup %>% 
    mutate(across(vars,
                  list(delta = ~ . - get(paste0("t0_", cur_column())),
                       delta_ma3 = ~ get(paste0("ma3_", cur_column())) - get(paste0("t2_", cur_column()))),
                       .names = "{.fn}_{.col}")) %>% return(.)
}

rel_vars <- c("real_Elem_Educ_Total_Exp", "real_Elem_Educ_Total_Exp_pp",
              "real_Property_Tax","real_Property_Tax_pp", 
              "real_gdp_total", "real_gdp_total_pc", 
              "real_Total_Educ_Total_Exp", "real_Total_Educ_Total_Exp_pp", 
              "log_real_Elem_Educ_Total_Exp", "log_real_Elem_Educ_Total_Exp_pp",
              "log_real_Property_Tax","log_real_Property_Tax_pp",
              "log_real_Total_Educ_Total_Exp", "log_real_Total_Educ_Total_Exp_pp", 
              "log_real_gdp_total", "log_real_gdp_total_pc")

fips_trends_dat <- data_transform(mines_restr, rel_vars, "fips", "year") %>% 
  left_join(., code_match, by = "fips")

cz_trends_dat <- data_transform(mines_cz, rel_vars, "cz_id", "year") %>% 
  left_join(., distinct(select(code_match, -fips)), multiple = "first", by = "cz_id")

cz_1990_trends_dat <- data_transform(mines_cz_1990, rel_vars, "cz_id_1990", "year") %>% 
  left_join(., distinct(select(code_match, -fips)), multiple = "first", by = "cz_id_1990")

#cz_1990_trends_dat %>% saveRDS(here('data/temp/cz_1990_trends_dat.rds'))

state_trends_dat <- data_transform(mines_state, rel_vars, "state_name", "year")

natl_trends_dat <- data_transform(mines_natl, rel_vars, "id", "year")

```

```{r, echo = FALSE, cache = TRUE}

# National macro data

# # Read in macro data
uer <- read.csv(here('data/raw/macro_vars/UNRATE.csv')) %>%
    mutate(DATE = ymd(DATE),
           FD_UNRATE = UNRATE - lag(UNRATE))

gdp <- read.csv(here('data/raw/macro_vars/GDP.csv')) %>%
  mutate(DATE = ymd(DATE),
         FD_GDP = GDP - lag(GDP))

gdpc1 <- read.csv(here('data/raw/macro_vars/GDPC1.csv')) %>%
  mutate(DATE = ymd(DATE),
         FD_GDPC1 = GDPC1 - lag(GDPC1))

# NOTE: Consider replacing with total population values from official source - this is just a sum over the dataset we have - missing ~100M people!
pop_natl <- mines_natl %>% select(year, pop_total)
gdp_annual <- read.csv(here('data/raw/macro_vars/GDPC1_annual.csv')) %>% 
  mutate(year = as.numeric(substr(DATE, 1, 4)),
         gdp_natl = GDPC1) %>% 
  left_join(., pop_natl, by = "year") %>% 
  mutate(gdp_natl_pc = gdp_natl/pop_total,
         diff_gdp_natl = gdp_natl - lag(gdp_natl, 1),
         diff_gdp_natl_pc = gdp_natl_pc - lag(gdp_natl_pc, 1),
         log_gdp_natl = log(gdp_natl),
         log_gdp_natl_pc = log(gdp_natl_pc),
         diff_log_gdp_natl = log_gdp_natl - lag(log_gdp_natl, 1),
         diff_log_gdp_natl_pc = log_gdp_natl_pc - lag(log_gdp_natl_pc, 1)) 

# gdp_annual %>% 
#   select(!c(DATE, GDPC1)) %>% 
#   pivot_longer(!year) %>% 
#   ggplot(aes(x = year, y = value)) + 
#   geom_line() + 
#   facet_wrap(~name, scales = "free", ncol = 1)

#row.names(gdp_annual) = gdp_annual[1]

macro_natl <- left_join(uer, gdp, by = "DATE") %>%
  left_join(., gdpc1, by = "DATE") %>%
  tibble

#plot(decompose(gdp_annu type = 'additive'))

```

### Relationships Between the Variables

::: panel-tabset
#### Elem. Ed. Exp. and Priv. Ind. GDP

::: panel-tabset
##### Levels

At a state level, we see that elementary ed expenditure does somewhat follow the shape of private industry GDP, with significant exceptions (ex. Michigan).

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(mines_state,
       aes(x= year))+

  geom_line(aes(y = real_Elem_Educ_Total_Exp, colour = 'red'))+
  geom_line(aes(y= real_gdp_priv_ind/20, colour = 'blue')) +
  facet_wrap(~state_name, scales = "free") +
  
  scale_color_discrete(name="",
                         breaks=c("red", "blue"),
                        labels = c("Elem. Educ. Exp.", "Priv. Ind. GDP")) +

  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +

  labs(title = "Plot of Elementary Education Expenditure and Private Industry GDP",
       subtitle = "in real USD by state",
       y = NULL)

```

##### Growth Rates

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(filter(mines_state, year > 2001),
       aes(x= year))+

  geom_line(aes(y = diff_log_real_Elem_Educ_Total_Exp, colour = 'red'))+
  geom_line(aes(y= diff_log_real_gdp_priv_ind, colour = 'blue')) +
  facet_wrap(~state_name, scales = "free") +
  
  scale_color_discrete(name="",
                         breaks=c("red", "blue"),
                        labels = c("Elem. Educ. Exp.", "Priv. Ind. GDP")) +

  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +

  labs(title = "Plot of Elementary Education Expenditure and Private Industry GDP",
       subtitle = "in growth rates real USD by state",
       y = NULL)

```
:::

#### Elem. Ed. Exp. and Property Tax

::: panel-tabset
##### Levels

Elementary ed expenditure follows the behaviour of property taxes in most states, with very few exceptions...

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(mines_state,
       aes(x= year))+
  geom_line(aes(y = real_Property_Tax, colour = 'red'))+
  geom_line(aes(y= real_Elem_Educ_Total_Exp/2, colour = 'blue')) +
  facet_wrap(~state_name, scales = "free") +
  
  scale_color_discrete(name="",
                         breaks=c("red", "blue"),
                        labels = c("Property Taxes", "Elem. Educ. Exp.")) +

  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +

  labs(title = "Plot of Elementary Education and Property Taxes",
       subtitle = "in real USD by state",
       y = NULL)

```

##### FDs

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(filter(mines_state, year > 2001),
       aes(x= year))+

  geom_line(aes(y = diff_log_real_Property_Tax, colour = 'red'))+
  geom_line(aes(y = diff_log_real_Elem_Educ_Total_Exp, colour = 'blue')) +
  facet_wrap(~state_name, scales = "free") +
  
  scale_color_discrete(name="",
                         breaks=c("red", "blue"),
                        labels = c("(fd) Property Taxes", "(fd) Elem. Educ. Exp.")) +

  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +

  labs(title = "Plot of Elementary Education and Property Taxes",
       subtitle = "in growth rates real USD by state",
       y = NULL)

```
:::

#### Property Tax and Priv. Ind. GDP

::: panel-tabset
##### Levels

Property taxes collected do also seem to grow in line with private industry GDP (all in real values) with a few notable exceptions: Delaware, Idaho, Indiana, Kansas, Minnesota, North Dakota.

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(mines_state,
       aes(x= year))+

  geom_line(aes(y = real_Property_Tax, colour = 'red'))+
  geom_line(aes(y= real_gdp_priv_ind/100, colour = 'blue')) +
  facet_wrap(~state_name, scales = "free") +
  
  scale_color_discrete(name="",
                         breaks=c("red", "blue"),
                        labels = c("Property Taxes", "Priv. Ind. GDP")) +

  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +

  labs(title = "Plot of Property Tax and Private Industry GDP",
       subtitle = "in real USD by state",
       y = NULL)

```

##### FDs

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(filter(mines_state, year > 2001),
       aes(x= year))+

  geom_line(aes(y = diff_log_real_Property_Tax, colour = 'red'))+
  geom_line(aes(y= diff_log_real_gdp_priv_ind, colour = 'blue')) +
  facet_wrap(~state_name, scales = "free") +
  
  scale_color_discrete(name="",
                         breaks=c("red", "blue"),
                        labels = c("Property Taxes", "Priv. Ind. GDP")) +

  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "bottom") +

  labs(title = "Plot of Property Tax and Private Industry GDP",
       subtitle = "in growth rates real USD by state",
       y = NULL)

```
:::
:::

# Defining Decline

What would be great is to be able to econometrically test when a county is "declining." In the first step, it would be good to identify when a county is declining overall (GDP, poverty, etc) but ideally eventually apply this to the education outcome. My hope is that being able to identify counties that are "declining" we can either use this variable as a covariate or as a central point of analysis. The below analysis looks at state-level variables as a first step (mainly to aid in visual comparison and plotting). This has now been expanded to county and commuting zones levels.

Some current options:

-   Simple time trends

<!-- -   Autocorrelation (X) -->

<!-- -   Trend indicator saturation - negative trend breaks? (\_) -->

-   Exhibit national or state-level GDP growth as global common factor; counties with negative factor loadings might arguably exhibit decline (\_)

-   Jennie's suggestion:

    -   plot county- , CZ, or state-level deviations from a national trend

::: panel-tabset
## Education

### State Time Trends

```{r, echo = FALSE, cache = TRUE}

trends <- feols(log_real_gdp_priv_ind ~ time:as.factor(state_name) | state_name + year, data = mines_state, se = "cluster", cluster = "state_name", panel.id = ~state_name+year)

#trends %>% etable(., se.below = TRUE)

trends$coefficients %>%

  tibble %>%

  rename(coef = '.') %>%

  ggplot() +

  geom_histogram(aes(x = coef), bins = 15) +

  theme_bw() +

  labs(title = "Histogram of State-level Trend Coefficient: (log) Real Private Industry GDP")+

  geom_vline(aes(xintercept = 0), colour = "red", linetype = "dotted")

```

::: panel-tabset
### Deviation from T0

One way to look at this descriptively is to look at the national, state and county-level trends. Here the "trend" is simply the difference between expenditure per pupil at the relevant level and expenditure per pupil in the first available observation as represented here: https://www.princeton.edu/\~erossi/DTNLC.pdf. (Note: this measure is imperfect for many reasons).

::: panel-tabset
#### States

```{r, fig.height = 10, echo = FALSE}
natl_pivoted <- natl_trends_dat %>% 
  select(year, contains('delta') & contains('log_real_Elem_Educ_Total_Exp_pp')) %>% 
  pivot_longer(!year) %>% 
  mutate(value = ifelse(grepl("ma3", name) & year == 2002, 0, value))

## State level trends with national trend reference line
state_trends_dat %>% 
  filter(state_name != "ME") %>% 
  select(state_name, year, contains('delta') & contains('log_real_Elem_Educ_Total_Exp_pp')) %>% 
  pivot_longer(!c(state_name, year)) %>% 
  ggplot(aes(x = year, y = value, colour = state_name)) +
      geom_line() +
  geom_line(data = natl_pivoted, aes(x = year, y = value), colour = "black") +
          facet_wrap(~name, scales = "free") +
  theme(legend.position = 'none') + 
  labs(title = "Difference from t0 in levels", 
       subtitle = "States")

```

```{r, fig.height = 10, echo = FALSE}

filter(state_trends_dat, year == 2018) %>%
  rename(state = state_name) %>% 
  plot_usmap(data = ., values = "real_Elem_Educ_Total_Exp_pp", regions = "states", exclude = c("AK", "HI")) +
  scale_fill_gradient(low = "thistle1", high = "purple4", na.value = "grey98") +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank(), legend.title = element_text("hello")) +
    labs(title = "State Level Elem. Ed. Exp. pp (2018)")

```

```{r, fig.height = 10, echo = FALSE}

filter(state_trends_dat, year == 2018) %>%
  rename(state = state_name) %>% 
  plot_usmap(data = ., values = "real_Total_Educ_Total_Exp_pp", regions = "states", exclude = c("AK", "HI")) +
  scale_fill_gradient(low = "thistle1", high = "purple4", na.value = "grey98") +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank(), legend.title = element_text("hello")) +
    labs(title = "State Level Elem. Ed. Exp. pp (2018)")

```

#### CZs

::: panel-tabset
```{r, echo = FALSE, results = 'asis', fig.height = 12}

temp_cz <- cz_trends_dat %>% 
  filter(year > 2001 & state_name != "ME") %>% 
  mutate(delta_ma3_log_real_Elem_Educ_Total_Exp_pp = ifelse(year == 2002, 0, delta_ma3_log_real_Elem_Educ_Total_Exp_pp))

temp_natl <- natl_trends_dat %>% 
  filter(year > 2001) %>% 
  mutate(delta_ma3_log_real_Elem_Educ_Total_Exp_pp = ifelse(year == 2002, 0, delta_ma3_log_real_Elem_Educ_Total_Exp_pp))

temp_state <- state_trends_dat %>% 
  filter(year > 2001 & state_name != "ME") %>% 
  mutate(delta_ma3_log_real_Elem_Educ_Total_Exp_pp = ifelse(year == 2002, 0, delta_ma3_log_real_Elem_Educ_Total_Exp_pp))

## CZ level trends by state with national trend reference line
for(k in names(select(temp_cz, contains('delta') & contains('log_real_Elem_Educ_Total_Exp_pp')))){
  cat('##### ', k, ' \n')
    p <- temp_cz %>% 
      filter(state_name != "ME") %>% 
      select(state_name, cz_id, year, k) %>% 
      pivot_longer(!c(state_name, year, cz_id)) %>% 
      ggplot(aes(x = year, y = value, colour = cz_id)) +
      geom_line() +
      geom_line(data = temp_natl, aes(x = year, y = get(k)), colour = "black", linetype = "dashed") +
      geom_line(data = temp_state, aes(x = year, y = get(k)), colour = "red", linetype = "dashed") +
      theme_minimal() +
      theme(legend.position = 'none', strip.text = element_text(size = 12)) +
      scale_colour_viridis(discrete = TRUE) + 
      facet_wrap(~state_name, ncol = 7, scales = "free") +
      labs(title = paste0("Trends: ", k), 
           subtitle = "Commuting Zones",
           x = "Year",
           y = k) 
    
    print(p)
              cat('\n')
              cat('\n')
              }

```
:::

#### Counties

::: panel-tabset
```{r, echo = FALSE, results = 'asis', fig.height = 12}

temp_fips <- fips_trends_dat %>% 
  filter(year > 2001 & state_name != "ME") %>% 
  mutate(delta_ma3_log_real_Elem_Educ_Total_Exp_pp = ifelse(year == 2002, 0, delta_ma3_log_real_Elem_Educ_Total_Exp_pp))


for(k in names(select(fips_trends_dat, contains('delta') & contains('log_real_Elem_Educ_Total_Exp_pp')))){
  cat('##### ', k, ' \n')
  p <- temp_fips %>% 
    filter(state_name != "ME") %>% 
    select(state_name, fips, year, k) %>% 
    pivot_longer(!c(state_name, year, fips)) %>% 
    ggplot(aes(x = year, y = value, colour = fips)) +
    geom_line() +
    geom_line(data = temp_natl, aes(x = year, y = get(k)), colour = "black", linetype = "dashed") +
    geom_line(data = filter(temp_state, state_name != "ME"), aes(x = year, y = get(k)), colour = "red", linetype = "dashed") +
    theme_minimal() +
    theme(legend.position = 'none', strip.text = element_text(size = 12)) +
    scale_colour_viridis(discrete = TRUE) + 
    facet_wrap(~state_name, ncol = 7, scales = "free") +
    theme(legend.position = 'none') + 
    labs(title = paste0("Trends: ", k), 
         subtitle = "Counties",
         x = "Year",
           y = k) 
  
  print(p)
              cat('\n')
              cat('\n')
              }

```
:::
:::

### Trends

Trend in log Elementary Education Expenditure (growth rate). The below plots show the coefficient estimate on the time-step variable in the trend equation. In the first panel I include no additional controls but in the latter two panels I include the state and national level growth rates as controls.

```{r, echo = FALSE, cache = TRUE, fig.height = 11}

# Create state-outlines
df <- usmap::us_map(regions = "states") %>% filter(!(abbr %in% c("HI", "AK")))

dat <- do.call('rbind', split(df, paste(df$abbr, df$piece)) |>
  lapply(function(x) st_sf(state = x$abbr[1],
                           st_sfc(st_polygon(list(cbind(x$x, x$y)))), 
                           crs = usmap::usmap_crs())))

natl_trend_pc <- mines_natl %>% 
  select(year, diff_log_real_Elem_Educ_Total_Exp_pp) %>% 
  distinct %>% 
  lm(diff_log_real_Elem_Educ_Total_Exp_pp ~ 1, data = .) %>% 
  .$coefficients %>% pluck(1)
  
county_ed_trends = mines_restr %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_Elem_Educ_Total_Exp_pp ~ 1, data = .)) %>% 
  ungroup 

county_ed_trends_controlled_natl = mines_restr %>% 
  left_join(., mines_natl, by = "year") %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_Elem_Educ_Total_Exp_pp.x ~ 1 +  diff_log_real_Elem_Educ_Total_Exp_pp.y, data = .)) %>% 
  ungroup 

county_ed_trends_controlled_state = mines_restr %>% 
   left_join(., mines_state, by = c("year", "state_name")) %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_Elem_Educ_Total_Exp_pp.x ~ 1 + diff_log_real_Elem_Educ_Total_Exp_pp.y, data = .)) %>% 
  ungroup 

county_ed_trends <- county_ed_trends %>% 
  rename("trend" = model) %>% 
  left_join(., county_ed_trends_controlled_natl, by = c('fips', 'state_name')) %>% 
  rename("trend_cont_natl" = model) %>% 
  left_join(., county_ed_trends_controlled_state, by = c('fips', 'state_name')) %>% 
  rename("trend_cont_state" = model)

county_ed_coefs <- tibble()
county_ed_sensitivities <- tibble()
for(k in 1:nrow(county_ed_trends)){
  s <- county_ed_trends %>% slice(k) %>% pull(fips)
  st <- county_ed_trends %>% slice(k) %>% pull(state_name)
  t <- county_ed_trends %>% slice(k) %>% 
    pull(trend) %>% first %>% .$coefficients %>% pluck(1) 
  t_n <- county_ed_trends %>% slice(k) %>% 
    pull(trend_cont_natl) %>% first %>% .$coefficients %>% pluck(1) 
  t_s <- county_ed_trends %>% slice(k) %>% 
    pull(trend_cont_state) %>% first %>% .$coefficients %>% pluck(1)
  

  s_n <- county_ed_trends %>% slice(k) %>% 
    pull(trend_cont_natl) %>% first %>% .$coefficients %>% pluck(2) 
  s_s <- county_ed_trends %>% slice(k) %>% 
    pull(trend_cont_state) %>% first %>% .$coefficients %>% pluck(2)
  
  county_ed_coefs <- rbind(county_ed_coefs, c(s,st, t, t_n, t_s))
  county_ed_sensitivities <- rbind(county_ed_sensitivities, c(s, st, s_n, s_s))
}

colnames(county_ed_coefs) <- names(county_ed_trends)
colnames(county_ed_sensitivities) <- c("fips", "state_name", "sensitivity_state", "sensitivity_natl")
county_ed_coefs <- county_ed_coefs %>% mutate(across(c(3:5), ~as.numeric(.)))
county_ed_sensitivities <- county_ed_sensitivities %>% mutate(across(c(3:4), ~as.numeric(.)))

county_ed_coefs %>% 
  pivot_longer(cols = contains("trend")) %>% 
  mutate(name = factor(name, levels=c("trend", "trend_cont_state", "trend_cont_natl"), labels = c("Trend in Elem. Ed. Exp. pp", "Trend in Elem. Ed. Exp. pp Controlled for State Trends", "Trend in Elem. Ed. Exp. pp Controlled for National Trends"))) %>% 
  ggplot() +
  geom_histogram(aes(x = value, fill = name), bins = 100, alpha = 0.9) +
  geom_vline(aes(xintercept = 0), colour = "darkgreen", linetype = "dotted") +
  geom_vline(aes(xintercept = natl_trend_pc), colour = "red", linetype = "dashed") +
  theme_bw() +
  facet_wrap(name ~ ., ncol = 1) +
  theme(legend.position = "none", strip.text = element_text(size = 14)) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "No. of Counties")


county_ed_coefs %>% 
  pivot_longer(cols = contains("trend")) %>% 
  mutate(name = factor(name, levels=c("trend", "trend_cont_state", "trend_cont_natl"), labels = c("Trend in Elem. Ed. Exp. pp", "Trend in Elem. Ed. Exp. pp Controlled for State Trends", "Trend in Elem. Ed. Exp. pp Controlled for National Trends"))) %>% 
  ggplot(aes(x = value, fill = name)) +
  geom_density(bins = 100, alpha = 0.5, position = "identity") +
  geom_vline(aes(xintercept = 0), colour = "darkgreen", linetype = "dotted") +
  geom_vline(aes(xintercept = natl_trend_pc), colour = "magenta", linetype = "dashed") +
  theme_bw() +
  theme(legend.position = "bottom", legend.title = element_blank()) +
  scale_fill_brewer(palette = "Set2") +
  labs(x = NULL, y = "No. of Counties")

# Education Expenditure
trend_ed_plot_pc <- plot_usmap(data = mute_outliers(county_ed_coefs, "trend"), values = "var_muted", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
  scale_fill_gradient2(midpoint = 0, na.value = "grey98") +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank(), legend.title = element_blank()) +
    labs(title = "County-level Trend in Elem. Ed. Exp. pp") +
  ggplot2::geom_sf(data = dat, colour = "blue", fill = NA, linewidth = 0.1)

# Education Expenditure
trend_ed_natl_plot_pc <- plot_usmap(data = mute_outliers(county_ed_coefs, "trend_cont_natl"), values = "var_muted", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
scale_fill_gradient2(na.value = "grey98", midpoint = 0) +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank(), legend.position = "none") +
    labs(title = "County-level Trend in Elem. Ed. Exp. pp",
       subtitle = "Controlling for National Level Elem. Ed. Exp. pp") +
  ggplot2::geom_sf(data = dat, colour = "blue", fill = NA, linewidth = 0.1)

# Education Expenditure
trend_ed_state_plot_pc <- plot_usmap(data = mute_outliers(county_ed_coefs, "trend_cont_state"), values = "var_muted", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
 scale_fill_gradient2(na.value = "grey98", midpoint = 0) +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank(), legend.position = "none") +
    labs(title = "County-level Trend in Elem. Ed. Exp. pp",
       subtitle = "Controlling for State Level Elem. Ed. Exp. pp") +
  ggplot2::geom_sf(data = dat, colour = "blue", fill = NA, linewidth = 0.1)

grid.arrange(trend_ed_plot_pc, trend_ed_state_plot_pc, trend_ed_natl_plot_pc, ncol = 3) %>% suppressWarnings()

rm(natl_trend_pc)
rm(county_ed_trends)
rm(county_ed_trends_controlled_natl)
rm(county_ed_trends_controlled_state)

```

```{r}

# Coefficient on state and national level trends
county_ed_sensitivities %>% 
  pivot_longer(cols = contains("sensitivity")) %>% 
  ggplot() +
  geom_histogram(aes(x = value, fill = name), bins = 100, position = "identity") +
  scale_fill_brewer(palette = "Set2")

plot_usmap(data = mute_outliers(county_ed_sensitivities, "sensitivity_natl"), values = "var_muted", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
  scale_fill_gradient2(na.value = "grey98", midpoint = 0) +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
   theme(panel.background = element_rect(color = "white", fill = "white"),
       plot.title = element_text(face = "bold"), legend.background=element_blank()) +
    labs(title = "County-level Sensitivity to National Trend in Elem. Ed. Exp. pp") +
  ggplot2::geom_sf(data = dat, colour = "blue", fill = NA, linewidth = 0.1)


plot_usmap(data = mute_outliers(county_ed_sensitivities, "sensitivity_state"), values = "var_muted", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
 scale_fill_gradient2(na.value = "grey98", midpoint = 0) +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank()) +
    labs(title = "County-level Sensitivity to State Trend in Elem. Ed. Exp. pp") +
   ggplot2::geom_sf(data = dat, colour = "blue", fill = NA, linewidth = 0.1)


```

### Growth Rates

::: panel-tabset
#### States

```{r, echo = FALSE}

mines_state %>% 
  filter(year != 2001) %>% 
  ggplot() +
  geom_line(aes(x = year, y = diff_log_real_Elem_Educ_Total_Exp_pp, color = state_name))
  

```

#### Commuting Zones

```{r, echo = FALSE}

# Outliers

mines_cz %>% 
  filter(year != 2001) %>% 
  ggplot() +
  geom_line(aes(x = year, y = diff_log_real_Elem_Educ_Total_Exp_pp, color =  cz_id)) +
  theme(legend.position = "none")

mines_cz %>% 
  group_by(cz_id) %>%
  mutate(across(diff_log_real_Elem_Educ_Total_Exp_pp, list(ma3 = ~rollmean(., 3, align = 'right', fill = NA),
                                                           ma2 = ~rollmean(., 2, align = 'right', fill = NA)), .names = "{.fn}_{.col}")) %>% 
  ungroup %>% 
  left_join(., distinct(select(code_match, -fips)), multiple = "first", by = "cz_id") %>% 
  filter(year > 2003) %>% 
  ggplot() +
  geom_line(aes(x = year, y = ma3_diff_log_real_Elem_Educ_Total_Exp_pp, color =  cz_id)) +
  theme(legend.position = "none") +
  facet_wrap(~state_name)
  

```

#### Counties

```{r, echo = FALSE}

# Outliers

pos_coefs <- mute_outliers(county_ed_coefs, "trend_cont_natl") %>% 
  filter(var_muted != trend_cont_natl & var_muted > 0) %>% 
  pull(fips) %>% unique

mines_df %>%
  filter(fips %in% pos_coefs) %>%
  select(fips, year, log_real_Property_Tax_pp, log_real_Elem_Educ_Total_Exp_pp) %>% 
  pivot_longer(!c(fips, year)) %>% 
  ggplot(aes(x = year, y = value, color = name)) +
  geom_line() +
  facet_wrap(~fips, scales = "free")

```
:::

### Outliers

::: panel-tabset
#### Negative

```{r, echo = FALSE}

# Outliers
neg_coefs <- mute_outliers(county_ed_coefs, "trend_cont_natl") %>% 
  filter(var_muted != trend_cont_natl & var_muted < 0) %>% 
  pull(fips) %>% unique

mines_df %>%
  filter(fips %in% neg_coefs) %>%
  select(fips, year, log_real_Property_Tax_pp, log_real_Elem_Educ_Total_Exp_pp) %>% 
  pivot_longer(!c(fips, year)) %>% 
  ggplot(aes(x = year, y = value, color = name)) +
  geom_line() +
  facet_wrap(~fips, scales = "free")

```

#### Positive

```{r, echo = FALSE}

# Outliers

pos_coefs <- mute_outliers(county_ed_coefs, "trend_cont_natl") %>% 
  filter(var_muted != trend_cont_natl & var_muted > 0) %>% 
  pull(fips) %>% unique

mines_df %>%
  filter(fips %in% pos_coefs) %>%
  select(fips, year, log_real_Property_Tax_pp, log_real_Elem_Educ_Total_Exp_pp) %>% 
  pivot_longer(!c(fips, year)) %>% 
  ggplot(aes(x = year, y = value, color = name)) +
  geom_line() +
  facet_wrap(~fips, scales = "free")

```
:::
:::

## GDPPC

::: panel-tabset
### Trends

Trend in log Real GDP (growth rate). The below plots show the coefficient estimate on the time-step variable in the trend equation. In the first panel I include no additional controls but in the latter two panels I include the state and national level growth rates as controls.

```{r, echo = FALSE, cache = TRUE}

natl_trend <- gdp_annual %>% 
  select(year, diff_log_gdp_natl) %>% 
  distinct %>% 
  lm(diff_log_gdp_natl ~ 1, data = .) %>% 
  .$coefficients %>% pluck(1)
  
county_trends = mines_restr %>% group_by(fips, state_name) %>% do(model = lm(diff_log_real_gdp_total ~ 1, data = .)) %>% ungroup 

county_trends_controlled_natl = mines_restr %>% 
  left_join(., gdp_annual, by = "year") %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_gdp_total ~ 1 + diff_log_gdp_natl, data = .)) %>% 
  ungroup 


county_trends_controlled_state = mines_restr %>% 
  left_join(., mines_state, by = c("year", "state_name")) %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_gdp_total.x ~ 1 + diff_log_real_gdp_total.y, data = .)) %>% 
  ungroup 

state_trends = mines_state %>% group_by(state_name) %>% do(model = lm(diff_log_real_gdp_total ~ 1, data = .)) %>% ungroup

state_trends_controlled_natl = mines_state %>% 
  left_join(., gdp_annual, by = "year") %>% 
  group_by(state_name) %>% 
  do(model = lm(diff_log_real_gdp_total ~ 1 + diff_log_gdp_natl, data = .)) %>% 
  ungroup

natl_trend = gdp_annual %>% do(model = lm(diff_log_gdp_natl ~ 1, data = .)) %>% pull(model) %>% first %>% .$coefficients %>% pluck(1)

county_trends <- county_trends %>% 
  rename("trend" = model) %>% 
  left_join(., county_trends_controlled_natl, by = c('fips', 'state_name')) %>% 
  rename("trend_cont_natl" = model) %>% 
  left_join(., county_trends_controlled_state, by = c('fips', 'state_name')) %>% 
  rename("trend_cont_state" = model)


state_trends <- state_trends %>%
  rename("trend" = model) %>%
  left_join(., state_trends_controlled_natl, by = c('state_name')) %>%
  rename("trend_cont_natl" = model)


county_coefs <- tibble()
for(k in 1:nrow(county_trends)){
  s <- county_trends %>% slice(k) %>% pull(fips)
  st <- county_trends %>% slice(k) %>% pull(state_name)
  t <- county_trends %>% slice(k) %>% 
    pull(trend) %>% first %>% .$coefficients %>% pluck(1) 
  t_n <- county_trends %>% slice(k) %>% 
    pull(trend_cont_natl) %>% first %>% .$coefficients %>% pluck(1) 
  t_s <- county_trends %>% slice(k) %>% 
    pull(trend_cont_state) %>% first %>% .$coefficients %>% pluck(1)
  
  county_coefs <- rbind(county_coefs, c(s,st, t, t_n, t_s))
}

colnames(county_coefs) <- names(county_trends)
county_coefs <- county_coefs %>% mutate(across(c(3:5), ~as.numeric(.)))

rm(county_trends)
rm(county_trends_controlled_natl)
rm(county_trends_controlled_state)


```

```{r, echo = FALSE, cache = TRUE}

county_trends_pc = mines_restr %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_gdp_total_pc ~ 1, data = .)) %>% ungroup

county_trends_controlled_natl_pc = mines_restr %>% 
  left_join(., gdp_annual, by = "year") %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_gdp_total_pc ~ 1 + diff_log_gdp_natl_pc, data = .)) %>% 
  ungroup 

county_trends_controlled_state_pc = mines_restr %>% 
  left_join(., mines_state, by = c("year", "state_name")) %>% 
  group_by(fips, state_name) %>% 
  do(model = lm(diff_log_real_gdp_total_pc.x ~ 1 + diff_log_real_gdp_total_pc.y, data = .)) %>% 
  ungroup 

state_trends_pc = mines_state %>% 
  group_by(state_name) %>% 
  do(model = lm(diff_log_real_gdp_total_pc ~ 1, data = .)) %>% 
  ungroup

state_trends_controlled_natl_pc = mines_state %>%
  left_join(., gdp_annual, by = "year") %>% 
  group_by(state_name) %>% 
  do(model = lm(diff_log_real_gdp_total_pc ~ 1 + diff_log_gdp_natl_pc, data = .)) %>% 
  ungroup

natl_trend_pc = gdp_annual %>% 
  select(year, diff_log_gdp_natl_pc) %>% 
  distinct %>% do(model = lm(diff_log_gdp_natl_pc ~ 1, data = .)) %>% 
  pull(model) %>% first %>% .$coefficients %>% pluck(1)

county_trends_pc <- county_trends_pc %>% 
  rename("trend_pc" = model) %>% 
  left_join(., county_trends_controlled_natl_pc, by = c('fips', 'state_name')) %>% 
  rename("trend_cont_natl_pc" = model) %>% 
  left_join(., county_trends_controlled_state_pc, by = c('fips', 'state_name')) %>% 
  rename("trend_cont_state_pc" = model)

county_coefs_pc <- tibble()
for(k in 1:nrow(county_trends_pc)){
  s <- county_trends_pc %>% slice(k) %>% pull(fips)
  st <- county_trends_pc %>% slice(k) %>% pull(state_name)
  t <- county_trends_pc %>% slice(k) %>% 
    pull(trend_pc) %>% first %>% .$coefficients %>% pluck(1) 
  t_n <- county_trends_pc %>% slice(k) %>% 
    pull(trend_cont_natl_pc) %>% first %>% .$coefficients %>% pluck(1) 
  t_s <- county_trends_pc %>% slice(k) %>% 
    pull(trend_cont_state_pc) %>% first %>% .$coefficients %>% pluck(1)
  
  county_coefs_pc <- rbind(county_coefs_pc, c(s,st, t, t_n, t_s))
}

colnames(county_coefs_pc) <- names(county_trends_pc)
county_coefs_pc <- county_coefs_pc %>% mutate(across(c(3:5), ~as.numeric(.)))

county_coefs_all <- county_coefs %>% 
  left_join(county_coefs_pc, by = c("fips", "state_name"))


state_trends_pc <- state_trends_pc %>%
  rename("trend_pc" = model) %>%
  left_join(., state_trends_controlled_natl_pc, by = c('state_name')) %>%
  rename("trend_cont_natl_pc" = model)


rm(county_trends_pc)
rm(county_trends_controlled_natl_pc)
rm(county_trends_controlled_state_pc)
rm(state_trends_controlled_natl_pc)


```

```{r, echo = FALSE, cache = TRUE, fig.height = 10}

# tplot <- county_coefs_all %>% 
#   select(-contains("pc")) %>% 
#   pivot_longer(cols = contains("trend")) %>% 
#   ggplot() +
#   geom_histogram(aes(x = value), bins = 100, fill = "darkgreen", alpha = 0.7) +
#   geom_vline(aes(xintercept = 0), colour = "grey", linetype = "dotted") +
#   geom_vline(aes(xintercept = natl_trend), colour = "red", linetype = "dashed") +
#   theme_bw() +
#   facet_wrap(factor(name,levels=c("trend", "trend_cont_state", "trend_cont_natl"), labels = c("Trend in GDP", "Trend in GDP Controlled for State Trends", "Trend in GDP Controlled for National Trends")) ~ ., ncol = 1) +
#     theme(legend.position = "none") +
#   labs(x = "Linear trend in Real GDP", y = "No. of Counties")

county_coefs_all %>% 
  pivot_longer(cols = contains("pc")) %>% 
  ggplot() +
  geom_histogram(aes(x = value), bins = 100, fill = "purple") +
  geom_vline(aes(xintercept = 0), colour = "grey", linetype = "dotted") +
  geom_vline(aes(xintercept = natl_trend_pc), colour = "red", linetype = "dashed") +
  theme_bw() +
  facet_wrap(factor(name,levels=c("trend_pc", "trend_cont_state_pc", "trend_cont_natl_pc"), labels = c("Trend in PC GDP", "Trend in PC GDP Controlled for State Trends", "Trend in PC GDP Controlled for National Trends")) ~ ., ncol = 1) +
  theme(legend.position = "none") +
  labs(x = "Linear trend in Real GDP PC", y = "No. of Counties")


```

```{r, echo = FALSE, cache = TRUE, fig.height = 14}
library(ggbrain)

# Education Expenditure
trend_plot_pc <- plot_usmap(data = county_coefs_all, values = "trend_pc", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
scale_fill_gradient2(na.value = "grey98", midpoint = 0) +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank()) +
    labs(title = "County-level Trend in Real GDP PC")

# Education Expenditure
trend_natl_plot_pc <- plot_usmap(data = county_coefs_all, values = "trend_cont_natl_pc", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
scale_fill_gradient2(na.value = "grey98", midpoint = 0) +
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank()) +
    labs(title = "County-level Trend in Real GDP PC",
       subtitle = "Controlling for National Level GDP PC")

# Education Expenditure
trend_state_plot_pc <- plot_usmap(data = county_coefs_all, values = "trend_cont_state_pc", regions = "counties", col = "gray90", linewidth = 0.01, exclude = c("AK", "HI")) +
scale_fill_gradient2(na.value = "grey98", midpoint = 0) + 
  #scale_fill_distiller(colours=c(bl,"white", re), name = "", direction = 1, na.value = "gray90", limits = c(-1, 1)) +
    theme(panel.background = element_rect(color = "white", fill = "white"),
        plot.title = element_text(face = "bold"), legend.background=element_blank()) +
    labs(title = "County-level Trend in Real GDP PC",
       subtitle = "Controlling for State Level GDP PC")

grid.arrange(trend_plot_pc, trend_state_plot_pc,  trend_natl_plot_pc, ncol = 1) %>% suppressWarnings()

```

```{r, echo = FALSE, cache = TRUE}

state_coefs <- tibble()

for(k in 1:nrow(state_trends)){
  s <- state_trends %>% slice(k) %>% pull(state_name)

  t <- state_trends %>% slice(k) %>%
    pull(trend) %>% first %>% .$coefficients %>% pluck(1)

  t_n <- state_trends %>% slice(k) %>%
    pull(trend_cont_natl) %>% first %>% .$coefficients %>% pluck(1)

  state_coefs <- rbind(state_coefs, c(s,t, t_n))
}

colnames(state_coefs) <- names(state_trends)
state_coefs %>%
  mutate(across(contains("trend"), as.numeric)) %>% 
  tibble %>% 
  pivot_longer(!state_name) %>% 
  ggplot()+
  geom_histogram(aes(x = value, fill = name), bins = 15, position = "identity", alpha  = 0.7) +
  geom_vline(aes(xintercept = 0), color = "magenta", linetype = "dotted") +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "Histogram of State-level Trend Coefficient: (log) Real GDP")

```

### Outliers

::: panel-tabset
#### Negative

```{r, echo = FALSE}

# Outliers
neg_coefs <- mute_outliers(county_ed_coefs, "trend_cont_natl") %>% 
  filter(var_muted != trend_cont_natl & var_muted < 0) %>% 
  pull(fips) %>% unique

mines_restr %>%
  filter(fips %in% neg_coefs) %>%
  select(fips, year, log_real_gdp_priv_ind_pc, log_real_Property_Tax_pp) %>% 
  pivot_longer(!c(fips, year)) %>% 
  ggplot(aes(x = year, y = value, color = name)) +
  geom_line() +
  facet_wrap(~fips, scales = "free")

```

#### Positive

```{r, echo = FALSE}

# Outliers

pos_coefs <- mute_outliers(county_coefs_all, "trend_cont_natl") %>% 
  filter(var_muted != trend_cont_natl & var_muted > 0) %>% 
  pull(fips) %>% unique

mines_restr %>%
  filter(fips %in% pos_coefs) %>%
  select(fips, year, log_real_gdp_priv_ind_pc, log_real_Property_Tax_pp) %>% 
  pivot_longer(!c(fips, year)) %>% 
  ggplot(aes(x = year, y = value, color = name)) +
  geom_line() +
  facet_wrap(~fips, scales = "free")


```
:::
:::
:::

# Progress on Jennie's Suggestion

1.  National data on GDP, UER, (and poverty)

```{r, echo = FALSE, cache = TRUE}

macro_natl %>% 
  select(-FD_UNRATE, -contains("GDPC1")) %>%
  mutate(GDP = GDP/10,
         UNRATE = UNRATE*100) %>%
  pivot_longer(cols = c(UNRATE, GDP, FD_GDP)) %>%
  filter(!is.na(value)) %>% 
  ggplot() +
  geom_line(aes(x = DATE, y = value)) +
  facet_wrap(~name, ncol = 1, scales = "free") +
  theme_minimal() +
  scale_x_date(limits = as.Date(c('2000-01-01','2022-01-01')))
  
```

```{r, echo = FALSE}

# library(covFactorModel)
# iX <- mines_restr %>% 
#   select(fips, year, real_gdp_total) %>% 
#   pivot_wider(names_from = "fips", values_from = "real_gdp_total") %>% 
#   mutate(year=as_date(year))
# 
# z <- read.zoo(iX, format = "%Y/%m/%d")
# x <- as.xts(z)
# 
# X <- xts(iX, order.by = as_date(iX$year))
# natl_gdp <- xts(gdp_annual[,-1], order.by=as_date(gdp_annual[,1]))
# 
# factor_model <- factorModel(xts(iX, order.by = as_date(iX$year)), type = "M", econ_fact = natl_gdp)
# cbind(alpha = factor_model$alpha, beta = factor_model$beta)

```

2.  State data on GDP, UER, and poverty

3.  Extract trends of state and national data

```{r, echo = FALSE, cache = TRUE}

county_gdp <- readRDS(here("data/temp/gdp_controls.RDS")) %>%
  mutate(across(contains("mining"), ~ifelse(is.na(.), 0, .))) %>%
  select(year, fips, real_gdp_total, real_gdp_priv_ind)

# county_gdp %>%
#   group_by(fips) %>%
#   mutate(across(c(real_gdp_total, real_gdp_priv_ind), list(ma5 = ~rollmean(., 5, align = 'right', fill = NA),
#                                                            ma3 = ~rollmean(., 3, align = 'right', fill = NA),
#                                                            ma2 = ~rollmean(., 2, align = 'right', fill = NA)),
#          .names = "{.fn}_{.col}"))


```

4.  Deviations at county level from national and state-level trends

5.  Plot deviations

6.  Saturation to find structural breaks in these deviations

Result: UE, GDP, and Poverty gaps

7.  PCA on those three gaps

8.  Retain index from PCA

9.  Step break in index

10. Line up with any notable events?

11. Place "gaps" in a panel structure for break detection

12. Autocorrelation plot to test no residual autocorrelation in model above with PCA indices.

```{r, echo = FALSE, cache = TRUE}

county_gdp %>% 
  ggplot() 

county_gdp %>% 
  ggplot() 

county_gdp %>% 
  ggplot() 

county_gdp %>% 
  ggplot() 

```

## Supplementary Materials

### Cointegration testing

Suggestions from Jennie:

Engle-Granger Test - bivariate scenario

Johanson Test - multivariate setting \[GDP, Property Taxes, Ed. Spending\]

### Autocorrelation

First-differences

```{r, echo = FALSE, cache = TRUE}

state_list <- unique(mines_state$state_name)

acf_list <- list()

for(s in 1:length(state_list)){

  ps <- mines_state %>%
    filter(state_name == state_list[s]) %>%
    pull(diff_real_gdp_priv_ind) %>%
    acf(., na.action = na.pass, plot = FALSE)
  acf_list[[s]] <- ps

}

acf_list[[1]] %>% plot
acf_list[[10]] %>% plot
acf_list[[15]] %>% plot
acf_list[[30]] %>% plot

```

-   Autocorrelation (Education only) (X)

-   Correlation plots (X)

-   Cross-correlation plot of Education and GDP (X)

-   Are GDP and Education Expenditure cointegrating (or do they exhibit a decoupling?) (\_)

### Autocorrelation

First-differences

```{r, echo = FALSE, cache = TRUE}

state_list <- unique(mines_state$state_name)

acf_list <- list()

for(s in 1:length(state_list)){
  ps <- mines_state %>%
    filter(state_name == state_list[s]) %>%
    pull(diff_real_Elem_Educ_Total_Exp) %>%
    acf(., na.action = na.pass, plot = FALSE)
  acf_list[[s]] <- ps

}

acf_list[[1]] %>% plot
acf_list[[10]] %>% plot
acf_list[[15]] %>% plot
acf_list[[30]] %>% plot

```

### Correlation Plots (Education and GDP)

```{r, fig.height = 10, echo = FALSE, cache = TRUE}

ggplot(mines_state,
       aes(x= real_gdp_priv_ind,
           y= real_Elem_Educ_Total_Exp))+
  geom_point()+
  facet_wrap(~state_name, scales = "free") +
  geom_smooth(fill="blue", alpha=0.1)+
  geom_smooth(method='lm', formula= y~x, se=FALSE, col="red",lty=2)+
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())+
    labs(title = "Correlation Plot of Elementary Ed. Expenditure and Private Industry GDP",
       subtitle = "in real USD by state",
       y = "Elem. Ed. Exp",
       x = "Priv. Ind. GDP")

```

### Cross-correlation of Educ Exp and Priv. Industry GDP

```{r, echo = FALSE, cache = TRUE}

for(s in 1:length(state_list[1:6])){
  gdp_var <- mines_state %>%
    filter(year != 2001) %>% 
    filter(state_name == state_list[s]) %>%
    pull(diff_real_gdp_priv_ind)
  
  elem_ed_var <- mines_state %>%
    filter(year != 2001) %>% 
    filter(state_name == state_list[s]) %>%
    pull(diff_real_Elem_Educ_Total_Exp)
  
  cor(gdp_var, elem_ed_var)

  ccf(as.numeric(gdp_var), as.numeric(elem_ed_var)) %>% print
}

#aTSA::coint.test(Deathts, VAXts, d = 1, nlag = 2, output = TRUE)

```
